### 2014-12-11 Report/project_report.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.
\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{threeparttable}

% Some handy commands
\newcommand{\norm}[1]{\left\Vert#1\right\Vert}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\Real}{\mathbb R}
\newcommand{\Complex}{\mathbb C}
\newcommand{\eps}{\varepsilon}
\newcommand{\To}{\longrightarrow}
\newcommand{\Ker}{\textup{Ker}}
\newcommand{\Img}{\textup{Img}}
\newcommand{\diag}{\textup{diag}}
\newcommand{\circulant}{\textup{circ}}
\newcommand\T{\rule{0pt}{2.6ex}}        % Top strut
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut

% Custom math defs
\def\Vec#1{\mathbf{#1}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}

\title{\LARGE \bf
The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper
}

\author{Lee Clement$^{1}$ and Valentin Peretroukhin$^{1}$% <-this % stops a space
%\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Institute for Aerospace Studies,
        University of Toronto, Toronto, ON, Canada
        {\tt \{lee.clement, valentin.peretroukhin\} @robotics.utias.utoronto.ca}
        }%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Monkey Style Chinese Kung Fu

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS} \label{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section*{ACKNOWLEDGMENT}
%\vspace{8pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\url#1{} % Get rid of url in citations -- also had to modify IEEEtran.bst
\bibliographystyle{IEEEtran}
%\bibliography{msckf_bib}

\end{document}

### 2014-12-11 Report/project_report.tex
Commit in case my computer explodes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{INTRODUCTION}
+\section{INTRODUCTION} \label{sec:introduction}
+\cite{Li:2012:ICRA} \cite{Li:2013:IJRR}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{MAPLESS VISION-AIDED INERTIAL NAVIGATION}
+The combination of visual and inertial sensors is a powerful tool for autonomous navigation.
+Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
+Since an IMU directly measures accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
+However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
+The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is stationary.
+On the other hand, cameras are not robust to motion blur induced by large accelerations.
+In these cases, the strength of the IMU's signal far exceeds its baseline noise and can be relied upon more heavily in estimating pose changes.
+
+The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
+A complicating factor in this problem is the absence of a known map of landmarks from which the camera can generate measurements.
+Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
+What follows is a discussion of two common techniques, the Extended Kalman Filter (EKF) and the sliding window filter (SWF), and a third hybrid approach, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The Extended Kalman Filter} \label{subsec:ekf}
+In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
+This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of landmarks by recursively updating the state as new measurements become available.
+
+[EKF EQUATIONS HERE?]
+
+Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
+This behavior is due to the fact that the length of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
+Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
+
+Another limitation of EKF-SLAM is that it is forgetful.
+Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later landmark measurements ought to constrain them.
+By locking in past poses, the EKF-SLAM formulation condemns itself to sub-optimally estimating both vehicle motion and landmark positions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sliding Window Bundle Adjustment} \label{subsec:slidingwindow}
+Another solution to vision-aided inertial navigation is to optimize a sliding window of vehicle poses and landmark positions to obtain locally optimal estimates of both.
+The optimization problem in the sliding window formulation is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
+
+[BATCH MATH GOES HERE?]
+
+An important advantage of the sliding window solution is that its computational cost depends only on the number of landmarks in the current window, and not the number of landmarks in the entire map.
+By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary spatial or temporal extent.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The Multi-State Constraint Kalman Filter} \label{subsec:msckf}
+The MSCKF \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of sliding-window bundle adjustment and traditional EKF-SLAM.
+The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose using batch-optimized estimates of landmarks that are visible across the entire window.
+
+[MSCKF MATH SOMEWHERE]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{EXPERIMENTS} \label{sec:experiments}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \bibliographystyle{IEEEtran}
-%\bibliography{msckf_bib}
+\bibliography{project_bib.bib}
 

### 2014-12-11 Report/project_report.tex
Some more words. Also pictures.

@@
 \section{MAPLESS VISION-AIDED INERTIAL NAVIGATION}
-The combination of visual and inertial sensors is a powerful tool for autonomous navigation.
+The combination of visual and inertial sensors is a powerful tool for autonomous navigation in unknown environments.
 Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
-What follows is a discussion of two common techniques, the Extended Kalman Filter (EKF) and the sliding window filter (SWF), and a third hybrid approach, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+What follows is a discussion of two common techniques, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid approach, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Extended Kalman Filter} \label{subsec:ekf}
+\subsection{The Extended Kalman Filter (EKF)} \label{subsec:ekf}
 In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sliding Window Bundle Adjustment} \label{subsec:slidingwindow}
-Another solution to vision-aided inertial navigation is to optimize a sliding window of vehicle poses and landmark positions to obtain locally optimal estimates of both.
-The optimization problem in the sliding window formulation is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
+\subsection{The Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
+In contrast to EKF-SLAM, the aim of the SWF is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
+The optimization problem in the Sliding Window Filter (SWF) is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
 
@@
 
-An important advantage of the sliding window solution is that its computational cost depends only on the number of landmarks in the current window, and not the number of landmarks in the entire map.
+An important advantage of the SWF is that its computational cost depends on the number of landmarks in the current window rather than the number of landmarks in the entire map.
 By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary spatial or temporal extent.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Multi-State Constraint Kalman Filter} \label{subsec:msckf}
-The MSCKF \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of sliding-window bundle adjustment and traditional EKF-SLAM.
-The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose using batch-optimized estimates of landmarks that are visible across the entire window.
+\subsection{The Multi-State Constraint Kalman Filter (MSCKF)} \label{subsec:msckf}
+The MSCKF \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
+The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of landmarks that are visible across the entire window.
+This update step typically occurs when a tracked landmark goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.
 
@@
 \section{EXPERIMENTS} \label{sec:experiments}
+We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
+The dataset consists of a rigidly attached stereo camera and IMU observing a set of 20 landmarks as they move along an arbitrary 3D path (Figure \ref{fig:sensorhead}).
+The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
+Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figs/sensorhead}
+    \caption{The sensor head used in our experiments.
+The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point landmarks.
+In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
+    \label{fig:sensorhead}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
+    \caption{Vicon ground truth for sensor head motion (blue) and landmark positions (red) in the ``Starry Night'' dataset.}
+    \label{fig:trajectory_groundtruth}
+\end{figure}
+
+Things quickly got out of hand when Matlab achieved sentience. 
+We were forced to abandon our lab and have been surviving in the harsh environment of the MarsDome for weeks.
+If this ever gets published, please send help!
 

### 2014-12-12 Report/project_report.tex
Added some MSCKF math

@@
 \section{INTRODUCTION} \label{sec:introduction}
-\cite{Li:2012:ICRA} \cite{Li:2013:IJRR}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{MAPLESS VISION-AIDED INERTIAL NAVIGATION}
 The combination of visual and inertial sensors is a powerful tool for autonomous navigation in unknown environments.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Extended Kalman Filter (EKF)} \label{subsec:ekf}
+\section{EXTENDED KALMAN FILTER} \label{sec:ekf}
 In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
 This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of landmarks by recursively updating the state as new measurements become available.
+%Landmarks are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
+%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
-In contrast to EKF-SLAM, the aim of the SWF is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
+\section{SLIDING WINDOW FILTER} \label{sec:slidingwindow}
+In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
 The optimization problem in the Sliding Window Filter (SWF) is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The Multi-State Constraint Kalman Filter (MSCKF)} \label{subsec:msckf}
-The MSCKF \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
+\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:msckf}
+The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
 The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of landmarks that are visible across the entire window.
@@
 
-[MSCKF MATH SOMEWHERE]
+\subsection{State Parametrization}
+As described in Section \ref{sec:experiments}, we evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations.
+In order to accommodate this alternative state parametrization, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We parametrize the IMU state at time $k$ as the 13-dimensional vector
+\begin{align}
+    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{g,k} & \Vec{b}_{v,k} & \Vec{p}_{G,k}^{IG} \ebm^T
+\end{align}
+where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{g,k}$ is the bias on the gyro measurements, $\Vec{b}_{v,k}$ is the bias on the velocity measurements, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).
+
+At time $k$, the state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
+\begin{align}
+    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k} & \hat{\Vec{q}}_{C_1 G} & \hat{\Vec{p}}_G^{C_1 G} & \hdots & \hat{\Vec{q}}_{C_N G} & \hat{\Vec{p}}_G^{C_N G} \ebm ^T
+\end{align}.
+
+\subsection{IMU State Estimate Propagation}
+The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by the following differential equations:
+\begin{align}
+    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
+    \dot{\hat{\Vec{b}}}_g  &= \Vec{0}_{3\times1} \\
+    \dot{\hat{\Vec{b}}}_v &= \Vec{0}_{3\times1} \\
+    \dot{\hat{\Vec{p}}}^G_I &= \hat{\Vec{C}}_{I G}^T\left(\hat{\Vec{v}}^G_I - \hat{\Vec{b}}_v \right)
+\end{align}
+where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
+\begin{align*}
+    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
+                                                                -\hat{\boldsymbol{\omega}}^T & 0
+                                                            \ebm 
+\end{align*}
+with
+\begin{align*}
+    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_g
+\end{align*}
+and
+\begin{align*}
+    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
+                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
+                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
+                                            \ebm .
+\end{align*}
+
+In our implementation we use a simple forward Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{CONCLUSIONS} \label{sec:conclusions}

### 2014-12-12 Report/project_report.tex
Minor notation change on the bias terms

@@
 The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
-A complicating factor in this problem is the absence of a known map of landmarks from which the camera can generate measurements.
+A complicating factor in the general form of this problem is the absence of a known map of landmarks from which the camera can generate measurements.
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
-What follows is a discussion of two common techniques, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid approach, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+What follows is a discussion of two common solutions, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
 
@@
 Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
-This behavior is due to the fact that the length of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
+This behavior is due to the fact that the dimension of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
 Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
@@
 An important advantage of the SWF is that its computational cost depends on the number of landmarks in the current window rather than the number of landmarks in the entire map.
-By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary spatial or temporal extent.
+By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
 
@@
 
-We parametrize the IMU state at time $k$ as the 13-dimensional vector
+In our implementation, we parametrize the IMU state at time $k$ as the 13-dimensional vector
 \begin{align}
-    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{g,k} & \Vec{b}_{v,k} & \Vec{p}_{G,k}^{IG} \ebm^T
+    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{\boldsymbol{\omega},k} & \Vec{b}_{\Vec{v},k} & \Vec{p}_{G,k}^{IG} \ebm^T
 \end{align}
-where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{g,k}$ is the bias on the gyro measurements, $\Vec{b}_{v,k}$ is the bias on the velocity measurements, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).
+where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{\boldsymbol{\omega},k}$ is the bias on the gyro measurements $\boldsymbol{\omega}_m$, $\Vec{b}_{\Vec{v},k}$ is the bias on the velocity measurements $\Vec{v}_m$, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).
 
-At time $k$, the state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
+At time $k$, the full state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
 \begin{align}
@@
     \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
-    \dot{\hat{\Vec{b}}}_g  &= \Vec{0}_{3\times1} \\
-    \dot{\hat{\Vec{b}}}_v &= \Vec{0}_{3\times1} \\
-    \dot{\hat{\Vec{p}}}^G_I &= \hat{\Vec{C}}_{I G}^T\left(\hat{\Vec{v}}^G_I - \hat{\Vec{b}}_v \right)
+    \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \\
+    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
+    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T\left(\Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}} \right)
 \end{align}
@@
 \begin{align*}
-    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_g
+    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}}
 \end{align*}
@@
 We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
-The dataset consists of a rigidly attached stereo camera and IMU observing a set of 20 landmarks as they move along an arbitrary 3D path (Figure \ref{fig:sensorhead}).
+The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 landmarks while moving along an arbitrary 3D path.
 The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).

### 2014-12-12 Report/project_report.tex
Added more stuff to state parametrization and augmentation

@@
 
-\subsection{State Parametrization}
-As described in Section \ref{sec:experiments}, we evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations.
+\subsection{MSCKF State Parametrization}
+We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations(see Section \ref{sec:experiments}).
 In order to accommodate this alternative state parametrization, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
@@
 \begin{align}
-    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{\boldsymbol{\omega},k} & \Vec{b}_{\Vec{v},k} & \Vec{p}_{G,k}^{IG} \ebm^T
+    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k}^T & \Vec{b}_{\boldsymbol{\omega},k}^T & \Vec{b}_{\Vec{v},k}^T & \Vec{p}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
 \end{align}
@@
 \begin{align}
-    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k} & \hat{\Vec{q}}_{C_1 G} & \hat{\Vec{p}}_G^{C_1 G} & \hdots & \hat{\Vec{q}}_{C_N G} & \hat{\Vec{p}}_G^{C_N G} \ebm ^T
-\end{align}.
+    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k}^T & \hat{\Vec{q}}_{C_1 G}^T & \hat{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \hat{\Vec{q}}_{C_N G}^T & \hat{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
+\end{align}
+
+We can also define the MSCKF ``error state'' at time $k$:
+\begin{align}
+    \tilde{\Vec{x}}_k &= \bbm \tilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \tilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \tilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
+\end{align}
+where
+\begin{align}
+    \tilde{\Vec{x}}_{I,k} &= \bbm \boldsymbol{\delta\theta}_I^T & \tilde{\Vec{b}}_{\boldsymbol{\omega},k}^T & \tilde{\Vec{b}}_{\Vec{v},k}^T & \tilde{\Vec{p}}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
+\end{align}
+is the 12-dimensional IMU error state.
+In the above, $\tilde{x}$ denotes the difference between the true value and the estimated value of the quantity $x$.
+The rotational errors $\boldsymbol{\delta\theta}$ are defined according to
+\begin{align*}
+    \boldsymbol{\delta}\Vec{q} &= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T.
+\end{align*}
+
+Accordingly, the MSCKF state covariance $\hat{\Vec{P}}_k$ is a $(12+6N)\times(12+6N)$ matrix that may be partitioned as
+\begin{align}
+    \hat{\Vec{P}}_k &= \bbm \hat{\Vec{P}}_{II,k} & \hat{\Vec{P}}_{IC,k} \\ \hat{\Vec{P}}_{IC,k}^T & \hat{\Vec{P}}_{CC,k} \ebm
+\end{align}
+where $\hat{\Vec{P}}_{II,k}$ is the $12\times12$ covariance matrix of the current IMU state, $\hat{\Vec{P}}_{CC,k}$ is the $6N\times6N$ covariance matrix of the camera poses, and $\hat{\Vec{P}}_{IC,k}$ is the $12\times6N$ cross-correlation between the current IMU state and the camera poses.
+
+\subsection{MSCKF State Augmentation}
+When a new camera image becomes available, the MSCKF state must be augmented with the current camera state.
+We obtain the camera state by applying the known transformation $\left(\Vec{q}_{CI}, \Vec{p}_I^{CI}\right)$ to a copy of the current IMU state:
+\begin{align}
+    \hat{\Vec{q}}_{C_{N+1} G} &= \Vec{q}_{CI} \otimes \hat{\Vec{q}}_{IG,k} \\
+    \hat{\Vec{p}}_G^{C_{N+1}G} &= \hat{\Vec{p}}_G^{IG} + \hat{\Vec{C}}_{IG,k} \hat{\Vec{p}}_I^{CI}
+\end{align}
+where $\hat{\Vec{C}}_{IG,k}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG,k}$ and $\otimes$ denotes quaternion multiplication.
+
+Assuming the MSCKF state has already been augmented by $N$ camera poses, we add the $\left(N+1\right)^\text{th}$ camera pose to the state as follows:
+\begin{align}
+    \hat{\Vec{x}}_k &\leftarrow \bbm \hat{\Vec{x}}_k^T & \hat{\Vec{q}}_{C_{N+1} G}^T & \left(\hat{\Vec{p}}_G^{C_{N+1}G}\right)^T \ebm^T
+\end{align}
+
+We must also augment the MSCKF state covariance:
+\begin{align}
+    \hat{\Vec{P}}_k &\leftarrow \bbm \Vec{1}_{12+6N} \\ \Vec{J}_k \ebm \hat{\Vec{P}}_k \bbm \Vec{1}_{12+6N} \\ \Vec{J}_k \ebm^T
+\end{align}
+where the Jacobian $\Vec{J}_k$ is given by
+\begin{align}
+\Vec{J}_k &= \bbm \hat{\Vec{C}}_{C I, k} & \Vec{0}_{3\times6} & \Vec{0}_{3\times3} & \Vec{0}_{3\times6N} \\
+                            \left(\hat{\Vec{C}}_{I G, k}^T \Vec{p}_{I,k}^{CI}\right)^\times & \Vec{0}_{3\times6} & \Vec{1}_3 & \Vec{0}_{3\times6N}
+                \ebm
+\end{align}
 
@@
 In our implementation we use a simple forward Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
+
+\subsection{MSCKF State Covariance Propagation}
 

### 2014-12-12 Report/project_report.tex
Added the rest of the prediction equations

@@
 \subsection{MSCKF State Parametrization}
-We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations(see Section \ref{sec:experiments}).
+We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
 In order to accommodate this alternative state parametrization, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
@@
 \begin{align}
-    \tilde{\Vec{x}}_k &= \bbm \tilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \tilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \tilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
+    \widetilde{\Vec{x}}_k &= \bbm \widetilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \widetilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \widetilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
 \end{align}
@@
 \begin{align}
-    \tilde{\Vec{x}}_{I,k} &= \bbm \boldsymbol{\delta\theta}_I^T & \tilde{\Vec{b}}_{\boldsymbol{\omega},k}^T & \tilde{\Vec{b}}_{\Vec{v},k}^T & \tilde{\Vec{p}}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
+    \widetilde{\Vec{x}}_{I,k} &= \bbm \boldsymbol{\delta\theta}_I^T & \widetilde{\Vec{b}}_{\boldsymbol{\omega},k}^T & \widetilde{\Vec{b}}_{\Vec{v},k}^T & \widetilde{\Vec{p}}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
 \end{align}
 is the 12-dimensional IMU error state.
-In the above, $\tilde{x}$ denotes the difference between the true value and the estimated value of the quantity $x$.
+In the above, $\widetilde{x}$ denotes the difference between the true value and the estimated value of the quantity $x$.
 The rotational errors $\boldsymbol{\delta\theta}$ are defined according to
-\begin{align*}
+\begin{align}
     \boldsymbol{\delta}\Vec{q} &= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T.
-\end{align*}
+\end{align}
 
 Accordingly, the MSCKF state covariance $\hat{\Vec{P}}_k$ is a $(12+6N)\times(12+6N)$ matrix that may be partitioned as
-\begin{align}
+\begin{align} \label{eq:covariance}
     \hat{\Vec{P}}_k &= \bbm \hat{\Vec{P}}_{II,k} & \hat{\Vec{P}}_{IC,k} \\ \hat{\Vec{P}}_{IC,k}^T & \hat{\Vec{P}}_{CC,k} \ebm
@@
 \subsection{MSCKF State Covariance Propagation}
+With reference to the partitions defined in Equation \ref{eq:covariance}, we compute the predicted camera-camera and IMU-camera state covariances as follows:
+\begin{align}
+    \hat{\Vec{P}}^-_{CC,k+1} &= \hat{\Vec{P}}_{CC,k} \\
+    \hat{\Vec{P}}^-_{IC,k+1} &= \boldsymbol{\Phi}\left(t_k + T, t_k \right) \hat{\Vec{P}}_{IC,k}
+\end{align}
+where $T$ is the IMU sampling period, and the state transition matrix $\boldsymbol{\Phi}\left(t_k + T, t_k \right)$ is obtained by integrating
+\begin{align} \label{eq:statetransition}
+    \dot{\boldsymbol{\Phi}}\left(t_k + \tau, t_k \right) &= \Vec{F} \boldsymbol{\Phi}\left(t_k + \tau, t_k \right), \tau\in\left[0,T\right]
+\end{align}
+with the initial condition $\boldsymbol{\Phi}\left(t_k, t_k \right) = \Vec{1}_{12}$.
+
+We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
+\begin{align} \label{eq:imucovar}
+    \dot{\hat{\Vec{P}}}_{II}\left(t_k, t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \notag \\ 
+                & + \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \Vec{F}^T \notag \\ 
+                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
+\end{align}
+with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$. 
+$\Vec{Q}_I$ is the covariance of the IMU measurement noise.
+
+The Jacobians $\Vec{F}$, $\Vec{G}$ in Equations \ref{eq:statetransition} and \ref{eq:imucovar} are given by
+\begin{align}
+\Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+            \ebm
+\end{align}
+\begin{align}
+\Vec{G} &=  \bbm    -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                                \Vec{0}_{3\times3} & \Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
+                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3}
+            \ebm .
+\end{align}
+
+\subsection{MSCKF State Correction Equations}
+[NULL SPACE TRICK, KALMAN GAIN, ETC.]
 

### 2014-12-12 Report/project_report.tex
My world is transposed.
Minor fixes and some more words in the SWF section

@@
 However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
-The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is stationary.
+The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is moving slowly.
 On the other hand, cameras are not robust to motion blur induced by large accelerations.
@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
-What follows is a discussion of two common solutions, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+What follows is a discussion of two common solutions, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), as well as a third hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
 
@@
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
-The optimization problem in the Sliding Window Filter (SWF) is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
+The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
 
@@
 By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
+
+However, the hard cut-off of the SWF may result in only some measurements of a particular feature contributing to the optimization.
+As a result, the filter may not maximally constrain some vehicle poses, and hence localization may be less accurate than we could expect from the full batch solution.
 
@@
 
-At time $k$, the full state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
+At time $k$, the full state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ 7-dimensional past camera poses in which active feature tracks were visible:
 \begin{align}
@@
 \end{align}
-where $\hat{\Vec{P}}_{II,k}$ is the $12\times12$ covariance matrix of the current IMU state, $\hat{\Vec{P}}_{CC,k}$ is the $6N\times6N$ covariance matrix of the camera poses, and $\hat{\Vec{P}}_{IC,k}$ is the $12\times6N$ cross-correlation between the current IMU state and the camera poses.
+where $\hat{\Vec{P}}_{II,k}$ is the $12\times12$ covariance matrix of the current IMU state, $\hat{\Vec{P}}_{CC,k}$ is the $6N\times6N$ covariance matrix of the camera poses, and $\hat{\Vec{P}}_{IC,k}$ is the $12\times6N$ cross-correlation between the current IMU state and the past camera poses.
 
 \subsection{MSCKF State Augmentation}
-When a new camera image becomes available, the MSCKF state must be augmented with the current camera state.
-We obtain the camera state by applying the known transformation $\left(\Vec{q}_{CI}, \Vec{p}_I^{CI}\right)$ to a copy of the current IMU state:
+When a new camera image becomes available, the MSCKF state must be augmented with the current camera pose.
+We obtain the camera pose by applying the known transformation $\left(\Vec{q}_{CI}, \Vec{p}_I^{CI}\right)$ to a copy of the current IMU pose:
 \begin{align}
     \hat{\Vec{q}}_{C_{N+1} G} &= \Vec{q}_{CI} \otimes \hat{\Vec{q}}_{IG,k} \\
-    \hat{\Vec{p}}_G^{C_{N+1}G} &= \hat{\Vec{p}}_G^{IG} + \hat{\Vec{C}}_{IG,k} \hat{\Vec{p}}_I^{CI}
+    \hat{\Vec{p}}_G^{C_{N+1}G} &= \hat{\Vec{p}}_G^{IG} + \hat{\Vec{C}}_{IG,k}^T \hat{\Vec{p}}_I^{CI}
 \end{align}
@@
 \begin{align}
-    \hat{\Vec{x}}_k &\leftarrow \bbm \hat{\Vec{x}}_k^T & \hat{\Vec{q}}_{C_{N+1} G}^T & \left(\hat{\Vec{p}}_G^{C_{N+1}G}\right)^T \ebm^T
+    \hat{\Vec{x}}_k &\leftarrow \bbm \hat{\Vec{x}}_k^T & \hat{\Vec{q}}_{C_{N+1} G}^T & \hat{\Vec{p}}_G^{C_{N+1} G \hspace{0.3em} T} \ebm^T .
 \end{align}
@@
                             \left(\hat{\Vec{C}}_{I G, k}^T \Vec{p}_{I,k}^{CI}\right)^\times & \Vec{0}_{3\times6} & \Vec{1}_3 & \Vec{0}_{3\times6N}
-                \ebm
+                \ebm .
 \end{align}
@@
 
-In our implementation we use a simple forward Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
+In our implementation we use a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
 

### 2014-12-14 Matlab/MSCKF.m
Corrected F and G (I think)

@@ -307,7 +307,7 @@ plot(tPlot, p_C_G_est(1,:) - p_C_G_GT(1,:), 'LineWidth', 1.2)
 hold on
 %plot(t(k1:k2), 3*sigma_x, '--r')
 %plot(t(k1:k2), -3*sigma_x, '--r')
-% ylim([-0.5 0.5])
+ylim([-0.5 0.5])
 xlim([tPlot(1) tPlot(end)])
 title('Translational Error')
 ylabel('\delta r_x')
@@ -318,7 +318,7 @@ plot(tPlot, p_C_G_est(2,:) - p_C_G_GT(2,:), 'LineWidth', 1.2)
 hold on
 %plot(t(k1:k2), 3*sigma_y, '--r')
 %plot(t(k1:k2), -3*sigma_y, '--r')
-% ylim([-0.5 0.5])
+ylim([-0.5 0.5])
 xlim([tPlot(1) tPlot(end)])
 ylabel('\delta r_y')
 
@@ -327,7 +327,7 @@ plot(tPlot, p_C_G_est(3,:) - p_C_G_GT(3,:), 'LineWidth', 1.2)
 hold on
 %plot(t(k1:k2), 3*sigma_z, '--r')
 %plot(t(k1:k2), -3*sigma_z, '--r')
-% ylim([-0.5 0.5])
+ylim([-0.5 0.5])
 xlim([tPlot(1) tPlot(end)])
 ylabel('\delta r_z')
 xlabel('t_k')

### 2014-12-14 Matlab/propagateMsckfStateAndCovar.m

@@ -3,7 +3,7 @@ function msckfState_prop = propagateMsckfStateAndCovar(msckfState, measurements_
     % Jacobians
     Q_imu = noiseParams.Q_imu*measurements_k.dT;
     F = calcF(msckfState.imuState, measurements_k);
-    G = calcG();
+    G = calcG(msckfState.imuState);
 
     %Propagate State
     msckfState_prop.imuState = propagateImuState(msckfState.imuState, measurements_k);

### 2014-12-14 Matlab/MSCKF.m
IMU noise params were in the wrong order for G. Filter still diverges, but it's getting better.

@@ -27,7 +27,8 @@ camera.p_C_I    = rho_v_c_v;            % 3x1 Camera position in IMU frame
 %Set up the noise parameters
 noiseParams.u_var_prime = y_var(1)/camera.f_u^2;
 noiseParams.v_var_prime = y_var(2)/camera.f_v^2;
-noiseParams.Q_imu = diag([v_var; w_var; 1e-12*ones(6,1)]);
+% noiseParams.Q_imu = diag([v_var; w_var; 1e-12*ones(6,1)]);
+noiseParams.Q_imu = diag([w_var; 1e-12*ones(3,1); v_var; 1e-12*ones(3,1)]);
 noiseParams.initialIMUCovar = 0.01*eye(12);
 noiseParams.initialCamCovar = 0.01*eye(6);
 

### 2014-12-14 Matlab/propagateImuState.m

@@ -5,7 +5,7 @@ function imuState_prop = propagateImuState(imuState_k, measurements_k)
 
     % Rotation state
     psi = (measurements_k.omega - imuState_k.b_g) * measurements_k.dT;
-    imuState_prop.q_IG = imuState_k.q_IG + 0.5 * omegaMat(psi) * imuState_k.q_IG ;
+    imuState_prop.q_IG = imuState_k.q_IG + 0.5 * omegaMat(psi) * imuState_k.q_IG;
 %     diffRot = axisAngleToRotMat(psi);
 %     C_IG_prop = diffRot * C_IG;
 %     imuState_prop.q_IG = rotMatToQuat(C_IG_prop);

### 2014-12-14 Matlab/MSCKF.m
Commit before magic merge. Awaiting glorious MSCKF leader

@@ -12,8 +12,8 @@ addpath('utils');
 load('dataset3.mat')
 
 %Dataset window bounds
-kStart = 500;
-kEnd = 600;
+kStart = 1200;
+kEnd = 1700;
 
 %Set up the camera parameters
 camera.c_u      = cu;                   % Principal point [pixels]
@@ -35,7 +35,7 @@ noiseParams.imageVariance = mean([noiseParams.u_var_prime, noiseParams.v_var_pri
 
 %MSCKF parameters
 msckfParams.minTrackLength = 2;
-msckfParams.maxTrackLength = 10;
+msckfParams.maxTrackLength = 5;
 msckfParams.maxGNCost = 1;
 
 % IMU state for plotting etc. Structures indexed in a cell array
@@ -201,11 +201,11 @@ for state_k = kStart:(kEnd-1)
 %              end
             
             
-%             [p_f_G, Jcost] = calcGNPosEst(track.camStates, track.observations, noiseParams);
+             [p_f_G, Jcost] = calcGNPosEst(track.camStates, track.observations, noiseParams);
 
             % Uncomment to use ground truth map instead
-            p_f_G = groundTruthMap(:, track.featureId);
-            Jcost = 0;
+            %p_f_G = groundTruthMap(:, track.featureId);
+            %Jcost = 0;
             
             if Jcost > msckfParams.maxGNCost
                 break;

### 2014-12-16 Report/project_report.tex
Started on the residual computation math

@@
 The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
-A complicating factor in the general form of this problem is the absence of a known map of landmarks from which the camera can generate measurements.
+A complicating factor in the general form of this problem is the absence of a known map of features from which the camera can generate measurements.
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
@@
 \section{EXTENDED KALMAN FILTER} \label{sec:ekf}
-In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
-This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of landmarks by recursively updating the state as new measurements become available.
-%Landmarks are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
+In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
+This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
+%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
 %In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.
@@
 Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
-This behavior is due to the fact that the dimension of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
+This behavior is due to the fact that the dimension of the state grows linearly with the number of features, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
 Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
@@
 Another limitation of EKF-SLAM is that it is forgetful.
-Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later landmark measurements ought to constrain them.
-By locking in past poses, the EKF-SLAM formulation condemns itself to sub-optimally estimating both vehicle motion and landmark positions.
+Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.
+By locking in past poses, the EKF-SLAM formulation condemns itself to sub-optimally estimating both vehicle motion and feature positions.
 
@@
 \section{SLIDING WINDOW FILTER} \label{sec:slidingwindow}
-In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
+In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
 The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
@@
 
-An important advantage of the SWF is that its computational cost depends on the number of landmarks in the current window rather than the number of landmarks in the entire map.
+An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
 By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
@@
 The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
-The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of landmarks that are visible across the entire window.
-This update step typically occurs when a tracked landmark goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.
+The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
+This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.
 
@@
 \subsection{IMU State Estimate Propagation}
-The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by the following differential equations:
+The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:
 \begin{align}
@@
 
-In our implementation we use a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
+In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
 
@@
 \Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T \\
                     \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
                     \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3} \\
             \ebm
@@
                                 \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
-                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3}
+                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
             \ebm .
@@
 
+\subsection{Feature Position Estimation}
+When a feature $f_j$ goes out of view of the camera, the MSCKF generates a batch-optimized estimate $\hat{\Vec{p}}_G^{f_j G}$ of the feature's position in the global frame.
+
 \subsection{MSCKF State Correction Equations}
+Now that we have estimated the positions of any features that have gone out of view, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
+We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from the $i^\text{th}$ camera pose in the window $C_i$:
+\begin{align} \label{eq:residual_def}
+    \Vec{r}_i^{(j)} &:= \Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}
+\end{align}
+where
+\begin{align}
+    \hat{\Vec{z}}_i^{(j)} &= \frac{1}{\hat{Z}^{(j)}_{C_i}} \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} \ebm ^T
+\end{align}
+with 
+\begin{align}
+\bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T &= \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
+\end{align}
+
+If we linearize Equation \ref{eq:residual_def} about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
+\begin{align}
+    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}_i} + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
+\end{align}
+where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $i$ with respect to the filter state and the position of the feature, respectively.
+These are given by
+\begin{align}
+    \Vec{H}_{\Vec{x}, i}^{(j)} &=   \\
+    \Vec{H}_{f, i}^{(j)} &= 
+\end{align}
+$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \sigma^2_{\text{im}} \Vec{1}_{2}$.
+
 [NULL SPACE TRICK, KALMAN GAIN, ETC.]
@@
 We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
-The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 landmarks while moving along an arbitrary 3D path.
-The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
+The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
+The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
@@
     \caption{The sensor head used in our experiments.
-The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point landmarks.
+The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
 In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
@@
     \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
-    \caption{Vicon ground truth for sensor head motion (blue) and landmark positions (red) in the ``Starry Night'' dataset.}
+    \caption{Vicon ground truth for sensor head motion (blue) and feature positions (red) in the ``Starry Night'' dataset.}
     \label{fig:trajectory_groundtruth}

### 2014-12-17 Report/project_report.tex
Added the rest of the MSCKF math.

@@
 \begin{align}
-    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k}^T & \Vec{b}_{\boldsymbol{\omega},k}^T & \Vec{b}_{\Vec{v},k}^T & \Vec{p}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
+    \Vec{x}_{I,k} &:= \bbm \Vec{q}_{IG,k}^T & \Vec{b}_{\boldsymbol{\omega},k}^T & \Vec{b}_{\Vec{v},k}^T & \Vec{p}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
 \end{align}
@@
 \begin{align}
-    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k}^T & \hat{\Vec{q}}_{C_1 G}^T & \hat{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \hat{\Vec{q}}_{C_N G}^T & \hat{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
+    \hat{\Vec{x}}_k &:= \bbm \hat{\Vec{x}}_{I,k}^T & \hat{\Vec{q}}_{C_1 G}^T & \hat{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \hat{\Vec{q}}_{C_N G}^T & \hat{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
 \end{align}
@@
 \begin{align}
-    \widetilde{\Vec{x}}_k &= \bbm \widetilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \widetilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \widetilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
+    \widetilde{\Vec{x}}_k &:= \bbm \widetilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \widetilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \widetilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
 \end{align}
@@
 \begin{align}
-    \widetilde{\Vec{x}}_{I,k} &= \bbm \boldsymbol{\delta\theta}_I^T & \widetilde{\Vec{b}}_{\boldsymbol{\omega},k}^T & \widetilde{\Vec{b}}_{\Vec{v},k}^T & \widetilde{\Vec{p}}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
+    \widetilde{\Vec{x}}_{I,k} &:= \bbm \boldsymbol{\delta\theta}_I^T & \widetilde{\Vec{b}}_{\boldsymbol{\omega},k}^T & \widetilde{\Vec{b}}_{\Vec{v},k}^T & \widetilde{\Vec{p}}_{G,k}^{IG \hspace{0.3em} T} \ebm^T
 \end{align}
@@
 \begin{align}
-    \boldsymbol{\delta}\Vec{q} &= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T.
+    \boldsymbol{\delta}\Vec{q} &:= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T.
 \end{align}
@@
 with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$. 
-$\Vec{Q}_I$ is the covariance of the IMU measurement noise.
-
-The Jacobians $\Vec{F}$, $\Vec{G}$ in Equations \ref{eq:statetransition} and \ref{eq:imucovar} are given by
+The Jacobians $\Vec{F}$, $\Vec{G}$ are given by
 \begin{align}
@@
                                 \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
-            \ebm .
+            \ebm ,
 \end{align}
+and $\Vec{Q}_I$ is the covariance of the IMU measurement noise.
 
@@
 Now that we have estimated the positions of any features that have gone out of view, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
-We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from the $i^\text{th}$ camera pose in the window $C_i$:
+We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from the $i^\text{th}$ camera pose $C_i$ in the window:
 \begin{align} \label{eq:residual_def}
@@
 \begin{align}
-\bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T &= \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
+    \hat{\Vec{p}}_{C_i}^{f_j C_i} &= \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T = \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
 \end{align}
@@
 \begin{align}
-    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}_i} + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
+    \Vec{r}_i^{(j)} &\simeq \Vec{H}_{\Vec{x}, i}^{(j)} \widetilde{\Vec{x}}_i + \Vec{H}_{f, i}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}_i^{(j)}
 \end{align}
@@
 \begin{align}
-    \Vec{H}_{\Vec{x}, i}^{(j)} &=   \\
-    \Vec{H}_{f, i}^{(j)} &= 
+    \Vec{H}_{\Vec{x}, i}^{(j)} &= \bbm \Vec{0} & \Vec{J}^{\left(j\right)}_i \left(\hat{\Vec{p}}_{C_i}^{f_j C_i}\right)^\times & -\Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G} & \Vec{0} \ebm \\
+    \Vec{H}_{f, i}^{(j)} &= \Vec{J}^{\left(j\right)}_i \hat{\Vec{C}}_{C_i G}
+\end{align}
+where the left $\Vec{0}$ in $\Vec{H}_{\Vec{x}, i}^{(j)}$ has dimension $2\times\left(12+6\left(i-1\right)\right)$, the right $\Vec{0}$ has dimension $2\times6\left(N-i\right)$, and
+\begin{align}
+    \Vec{J}^{\left(j\right)}_i &= \frac{1}{\left(\hat{Z}^{(j)}_{C_i}\right)^2}
+    \bbm \hat{Z}^{(j)}_{C_i} & 0 & -\hat{X}^{(j)}_{C_i} \vspace{0.5em} \\
+             0 & \hat{Z}^{(j)}_{C_i} & -\hat{Y}^{(j)}_{C_i}
+    \ebm .
 \end{align}
@@
 
-[NULL SPACE TRICK, KALMAN GAIN, ETC.]
+By stacking the errors $\Vec{r}_i^{(j)}$, we arrive at an expression for the $2M_j \times 1$ measurement error vector of feature $f_j$ over the entire window of camera poses, where $M_j$ is the number of camera poses from which feature $f_j$ was observed:
+\begin{align} \label{eq:residual_stacked}
+    \Vec{r}^{(j)} &\simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
+\end{align}
+However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
+Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]
+
+In order to transform $\Vec{r}^{(j)}$ into a usable form for the EKF, we can define a semi-unitary matrix $\Vec{A}$ whose columns form the basis of the left nullspace of $\Vec{H}_{f}^{(j)}$, and project $\Vec{r}^{(j)}$ into this nullspace to obtain an error equation of the correct form:
+\begin{align}
+    \Vec{r}_o^{(j)} &:= \Vec{A}^T \left(\Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}\right) \\
+                    &\simeq \Vec{A}^T \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{0} + \Vec{A}^T \Vec{n}^{(j)} \\
+                    &=: \Vec{H}_{o}^{(j)} \widetilde{\Vec{x}} + \Vec{n}_o^{(j)}
+\end{align}
+
+Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $\left(2M_j - 3\right) \times 3$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
+Furthermore, because $\Vec{A}$ is semi-unitary, the covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o = \sigma^2_{\text{im}}\Vec{A}^T\Vec{A} = \sigma^2_{\text{im}}\Vec{1}_{2M_j - 3}$.
+
+We can now stack all the errors $\Vec{r}_o^{(j)}$ for all the features in the current batch to arrive at
+\begin{align} \label{eq:residual_nullspace_stacked}
+    \Vec{r}_o &= \Vec{H}_o \widetilde{\Vec{x}} + \Vec{n}_o .
+\end{align}
+The dimension of this vector can be quite large in practice, so we use the QR-decomposition of $\Vec{H}_o$ to reduce the computational complexity of the EKF update:
+\begin{align}
+    \Vec{H}_o = \bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm
+\end{align}
+where $\Vec{Q}_1, \Vec{Q}_2$ are unitary matrices and $\Vec{T}_\Vec{H}$ is an upper-triangular matrix.
+Substituting this result into Equation \ref{eq:residual_nullspace_stacked} and pre-multiplying by $\bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm ^T$, we obtain
+\begin{align}
+    \bbm \Vec{Q}_1^T \Vec{r}_o \\ \Vec{Q}_2^T \Vec{r}_o \ebm &= \bbm \Vec{T}_\Vec{H} \\ \Vec{0} \ebm \widetilde{\Vec{x}} + \bbm \Vec{Q}_1^T \Vec{n}_o \\ \Vec{Q}_2^T \Vec{n}_o \ebm .
+\end{align}
+Noting that the quantity $\Vec{Q}_2^T \Vec{r}_o$ is only noise, we discard it and define a new error term that we use in the EKF update:
+\begin{align}
+    \Vec{r}_n &:= \Vec{Q}_1^T \Vec{r}_o = \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{Q}_1^T \Vec{n}_o =: \Vec{T}_\Vec{H} \widetilde{\Vec{x}} + \Vec{n}_n
+\end{align} 
+The covariance matrix of $\Vec{n}_n$ is given by $\Vec{R}_n = \Vec{Q}_1^T \Vec{R}_o \Vec{Q}_1 = \sigma^2_\text{im} \Vec{1}_r$ where $r$ is the number of columns in $\Vec{Q}_1$.
+
+Finally, we can formulate the Kalman gain and correction equations to arrive at the updated estimates for the filter state and covariance:
+\begin{align}
+    \Vec{K} &= \hat{\Vec{P}}^-_{k+1}\Vec{T}_{H}^T\left(\Vec{T}_{H}\hat{\Vec{P}}^-_{k+1}\Vec{T}_{H}^T + \Vec{R}_{n} \right)^{-1} \\
+    \Delta\Vec{x}_{k+1} &= \Vec{K}\Vec{r}_{n} \\
+    \hat{\Vec{P}}_{k+1} &= \left(\Vec{1}_{12+6N} - \Vec{K}\Vec{T}_{H} \right)\hat{\Vec{P}}^-_{k+1}\left(\Vec{1}_{12+6N} - \Vec{K}\Vec{T}_{H} \right)^T \notag \\ 
+                        &+ \Vec{K}\Vec{R}_{n}\Vec{K}^T .
+\end{align}
 

### 2014-12-17 Report/project_report.tex
Minor in-text equation reference notation change

@@
 \subsection{MSCKF State Covariance Propagation}
-With reference to the partitions defined in Equation \ref{eq:covariance}, we compute the predicted camera-camera and IMU-camera state covariances as follows:
+With reference to the partitions defined in (\ref{eq:covariance}), we compute the predicted camera-camera and IMU-camera state covariances as follows:
 \begin{align}
@@
 
-If we linearize Equation \ref{eq:residual_def} about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
+If we linearize (\ref{eq:residual_def}) about the estimates for the camera pose and feature position, we obtain an estimate of the exteroceptive measurement error
 \begin{align}
@@
 where $\Vec{Q}_1, \Vec{Q}_2$ are unitary matrices and $\Vec{T}_\Vec{H}$ is an upper-triangular matrix.
-Substituting this result into Equation \ref{eq:residual_nullspace_stacked} and pre-multiplying by $\bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm ^T$, we obtain
+Substituting this result into (\ref{eq:residual_nullspace_stacked}) and pre-multiplying by $\bbm \Vec{Q}_1 & \Vec{Q}_2 \ebm ^T$, we obtain
 \begin{align}

### 2014-12-20 Report/project_report.tex
Minor changes. Still haven't found the problem yet.

@@
 
-Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $\left(2M_j - 3\right) \times 3$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
+Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $2M_j \times \left(2M_j - 3\right)$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
 Furthermore, because $\Vec{A}$ is semi-unitary, the covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o = \sigma^2_{\text{im}}\Vec{A}^T\Vec{A} = \sigma^2_{\text{im}}\Vec{1}_{2M_j - 3}$.

### 2014-12-20 Matlab/msckf/MSCKF.m

@@ -9,11 +9,11 @@ clear;
 close all;
 clc;
 addpath('utils');
-load('../dataset3_fresh.mat')
+load('../dataset3_fresh2.mat')
 
 %Dataset window bounds
-kStart = 500;
-kEnd = 1000;
+kStart = 1215;
+kEnd = 1715;
 
 %Set constant
 numLandmarks = size(y_k_j,3);
@@ -33,13 +33,12 @@ noiseParams.v_var_prime = y_var(2)/camera.f_v^2;
 noiseParams.Q_imu = diag([w_var; 1e-12*ones(3,1); v_var; 1e-12*ones(3,1)]); % [w; w bias; v; v bias]
 noiseParams.initialIMUCovar = 1e-12*eye(12); % should be small since we're initializing with ground truth
 
-noiseParams.imageVariance = 2*mean([noiseParams.u_var_prime, noiseParams.v_var_prime]);  % Slightly hacky. Used to compute the Kalman gain and corrected covariance in the EKF step
+noiseParams.imageVariance = mean([noiseParams.u_var_prime, noiseParams.v_var_prime]);  % Slightly hacky. Used to compute the Kalman gain and corrected covariance in the EKF step
 
 %MSCKF parameters
-msckfParams.maxGNCost = 1;
-msckfParams.minTrackLength = 2;     % Uncomment to dead-reckon only
-msckfParams.maxTrackLength = 5;     % Uncomment to wait for features to go out of view
-msckfParams.maxGNCost = inf;          % Uncomment to allow any triangulation, no matter how bad
+msckfParams.minTrackLength = 2;     % Set to inf to dead-reckon only
+msckfParams.maxTrackLength = 30;     % Set to inf to wait for features to go out of view
+msckfParams.maxGNCost      = inf;     % Set to inf to allow any triangulation, no matter how bad
 
 % IMU state for plotting etc. Structures indexed in a cell array
 imuStates = cell(1,numel(t));
@@ -130,7 +129,7 @@ for state_k = kStart:(kEnd-1)
     
     
     %% ==========================STATE PROPAGATION======================== %%
-
+    
     %Propagate state and covariance
     msckfState = propagateMsckfStateAndCovar(msckfState, measurements{state_k}, noiseParams);
 
@@ -151,6 +150,7 @@ for state_k = kStart:(kEnd-1)
             track = featureTracks{trackedFeatureIds == featureId};
             
             if meas_k(1,1) == -1 || size(track.observations, 2) >= msckfParams.maxTrackLength
+                                
[…]

### 2014-12-20 Matlab/msckf/calcHoj.m

@@ -7,7 +7,6 @@ function [H_o_j, A_j] = calcHoj(p_f_G, msckfState, camStateIndices)
 
 
 N = length(msckfState.camStates);
-
 M = length(camStateIndices);
 H_f_j = zeros(2*M, 3);
 H_x_j = zeros(2*M, 12 + 6*N);
@@ -35,12 +34,9 @@ for camStateIndex = camStateIndices
     c_i = c_i + 1;
 end
 
-if all(~isnan(H_f_j(:)))
-    A_j = null(H_f_j');
-    H_o_j = A_j'*H_x_j;
-else
-    a =2;
-end
+
+A_j = null(H_f_j');
+H_o_j = A_j'*H_x_j;
 
 end
 

### 2014-12-20 Matlab/msckf/propagateImuState.m

@@ -2,7 +2,7 @@ function imuState_prop = propagateImuState(imuState_k, measurements_k)
 % prop == propagated to k+1
 
     C_IG = quatToRotMat(imuState_k.q_IG);
-
+    
     % Rotation state
     psi = (measurements_k.omega - imuState_k.b_g) * measurements_k.dT;
     imuState_prop.q_IG = imuState_k.q_IG + 0.5 * omegaMat(psi) * imuState_k.q_IG;

### 2014-12-20 Matlab/msckf/propagateMsckfStateAndCovar.m

@@ -14,7 +14,7 @@ function msckfState_prop = propagateMsckfStateAndCovar(msckfState, measurements_
                                 + msckfState.imuCovar * F' ...
                                 + G * Q_imu * G' ) ...
                                         * measurements_k.dT;
-                     
+                                    
     % Camera-Camera Covariance
     msckfState_prop.camCovar = msckfState.camCovar;
     

### 2014-12-20 Matlab/msckf/updateStateHistory.m

@@ -28,7 +28,7 @@ function imuStates_up = updateStateHistory(imuStates, msckfState, camera, state_
     p_I_C = - C_CI' * camera.p_C_I;
     
     
-    for camIdx = 1:size(msckfState.camStates, 2)      
+    for camIdx = 1:size(msckfState.camStates, 2)             
         q_IG = quatLeftComp(q_IC) * msckfState.camStates{camIdx}.q_CG;
         
         C_IG = quatToRotMat(q_IG);

### 2014-12-20 Matlab/msckf/utils/buildUpdateQuat.m

@@ -14,4 +14,6 @@ function updateQuat = buildUpdateQuat(deltaTheta)
     else
         updateQuat = [deltaq; sqrt(1 - checkNorm)];
     end
+    
+    updateQuat = updateQuat / norm(updateQuat);
 end

### 2014-12-21 Matlab/msckf/MSCKF.m
Another term was missing from F

@@ -9,8 +9,8 @@ clear;
 close all;
 clc;
 addpath('utils');
-% load('../dataset3_fresh2.mat')
-load('../dataset3.mat')
+load('../dataset3_fresh2.mat')
+% load('../dataset3.mat')
 
 %Dataset window bounds
 kStart = 1215;
@@ -32,12 +32,12 @@ camera.p_C_I    = rho_v_c_v;            % 3x1 Camera position in IMU frame
 noiseParams.u_var_prime = y_var(1)/camera.f_u^2;
 noiseParams.v_var_prime = y_var(2)/camera.f_v^2;
 
-noiseParams.Q_imu = diag([w_var; 1e-12*ones(3,1); v_var; 1e-12*ones(3,1)]); % [w; w bias; v; v bias]
+noiseParams.Q_imu = diag([w_var; 1e-6*ones(3,1); v_var; 1e-6*ones(3,1)]); % [w; w bias; v; v bias]
 noiseParams.initialIMUCovar = 1e-12*eye(12); % should be small since we're initializing with ground truth
 
 %MSCKF parameters
-msckfParams.minTrackLength = 5;     % Set to inf to dead-reckon only
-msckfParams.maxTrackLength = 60;     % Set to inf to wait for features to go out of view
+msckfParams.minTrackLength = 2;     % Set to inf to dead-reckon only
+msckfParams.maxTrackLength = 80;     % Set to inf to wait for features to go out of view
 msckfParams.maxGNCost      = inf;     % Set to inf to allow any triangulation, no matter how bad
 
 % IMU state for plotting etc. Structures indexed in a cell array
@@ -164,8 +164,7 @@ for state_k = kStart:(kEnd-1)
                
                 %Remove the track
                 featureTracks = featureTracks(trackedFeatureIds ~= featureId);
-                trackedFeatureIds(trackedFeatureIds == featureId) = [];
-                
+                trackedFeatureIds(trackedFeatureIds == featureId) = [];                
             else
                 %Append observation and append id to cam states
                 featureTracks{trackedFeatureIds == featureId}.observations(:, end+1) = meas_k;
@@ -199,16 +198,9 @@ for state_k = kStart:(kEnd-1)
 
             %Estimate feature 3D location through Gauss Newton inverse depth
             %optimization
-            
-%              camStatesGT = {};
-%              for c_i_temp = 1:length(track.camStates)
[…]

### 2014-12-21 Matlab/msckf/calcF.m

@@ -5,10 +5,12 @@ function F = calcF(imuState_k, measurements_k)
     F = zeros(12,12);
     
     omegaHat = measurements_k.omega - imuState_k.b_g;
+    vHat = measurements_k.v - imuState_k.b_v;
     C_IG = quatToRotMat(imuState_k.q_IG);
     
     F(1:3,1:3) = -crossMat(omegaHat);
     F(1:3,4:6) = -eye(3);
+    F(10:12,1:3) = -C_IG' * crossMat(vHat);
     F(10:12,7:9) = -C_IG';
   
 end

### 2014-12-21 Matlab/msckf/initializeMSCKF.m

@@ -8,21 +8,9 @@ firstImuState.b_g = zeros(3,1);
 firstImuState.b_v = zeros(3,1);
 msckfState.imuState = firstImuState;
 msckfState.imuCovar = noiseParams.initialIMUCovar;
-% msckfState.camCovar = noiseParams.initialCamCovar;
-% msckfState.imuCamCovar = zeros(12, 6);
 msckfState.camCovar = [];
 msckfState.imuCamCovar = [];
-
-% Compute camera pose from current IMU pose
-% camState = {};
-% C_IG = quatToRotMat(msckfState.imuState.q_IG);
-% camState.q_CG = quatLeftComp(camera.q_CI) * msckfState.imuState.q_IG;
-% camState.p_C_G = msckfState.imuState.p_I_G + C_IG' * camera.p_C_I;
-% camState.trackedFeatureIds = [];
-% camState.state_k = state_k;
-
 msckfState.camStates = {};
-% msckfState.camStates{1} = camState;
 
 msckfState = augmentState(msckfState, camera, state_k);
 

### 2014-12-21 Matlab/msckf/propagateImuState.m

@@ -18,6 +18,7 @@ function imuState_prop = propagateImuState(imuState_k, measurements_k)
     imuState_prop.b_v = imuState_k.b_v;
     
     % Translation state
-    imuState_prop.p_I_G = C_IG' * (measurements_k.v - imuState_k.b_v) * measurements_k.dT + imuState_k.p_I_G;
+    d = (measurements_k.v - imuState_k.b_v) * measurements_k.dT;
+    imuState_prop.p_I_G = C_IG' * d + imuState_k.p_I_G;
     
 end

### 2014-12-21 Matlab/msckf/propagateMsckfStateAndCovar.m

@@ -8,18 +8,21 @@ function msckfState_prop = propagateMsckfStateAndCovar(msckfState, measurements_
     %Propagate State
     msckfState_prop.imuState = propagateImuState(msckfState.imuState, measurements_k);
 
+    % State Transition Matrix
+    Phi = eye(size(F,1)) + F * measurements_k.dT; % Leutenegger 2013
+    
     % IMU-IMU Covariance
-    msckfState_prop.imuCovar = msckfState.imuCovar + ...
-                                ( F * msckfState.imuCovar ...
-                                + msckfState.imuCovar * F' ...
-                                + G * Q_imu * G' ) ...
-                                        * measurements_k.dT;
+%     msckfState_prop.imuCovar = msckfState.imuCovar + ...
+%                                 ( F * msckfState.imuCovar ...
+%                                 + msckfState.imuCovar * F' ...
+%                                 + G * Q_imu * G' ) ...
+%                                         * measurements_k.dT;
+
+    msckfState_prop.imuCovar = Phi * msckfState.imuCovar * Phi' ...
+                                + G * Q_imu * G' * measurements_k.dT; % Leutenegger 2013
                                     
     % Camera-Camera Covariance
     msckfState_prop.camCovar = msckfState.camCovar;
-    
-    % State Transition Matrix
-    Phi = eye(size(F,1)) + F * measurements_k.dT; % Leutenegger 2013
     
     % IMU-Camera Covariance
     msckfState_prop.imuCamCovar = Phi * msckfState.imuCamCovar;

### 2014-12-21 Matlab/msckf/MSCKF.m
Minor parameter updates. Seems to work well now.

@@ -36,8 +36,8 @@ noiseParams.Q_imu = diag([w_var; 1e-6*ones(3,1); v_var; 1e-6*ones(3,1)]); % [w; 
 noiseParams.initialIMUCovar = 1e-12*eye(12); % should be small since we're initializing with ground truth
 
 %MSCKF parameters
-msckfParams.minTrackLength = 2;     % Set to inf to dead-reckon only
-msckfParams.maxTrackLength = 80;     % Set to inf to wait for features to go out of view
+msckfParams.minTrackLength = 20;     % Set to inf to dead-reckon only
+msckfParams.maxTrackLength = 100;     % Set to inf to wait for features to go out of view
 msckfParams.maxGNCost      = inf;     % Set to inf to allow any triangulation, no matter how bad
 
 % IMU state for plotting etc. Structures indexed in a cell array
@@ -198,9 +198,9 @@ for state_k = kStart:(kEnd-1)
 
             %Estimate feature 3D location through Gauss Newton inverse depth
             %optimization
-%             [p_f_G, Jcost] = calcGNPosEst(track.camStates, track.observations, noiseParams);
+            [p_f_G, Jcost] = calcGNPosEst(track.camStates, track.observations, noiseParams);
             % Uncomment to use ground truth map instead
-            p_f_G = groundTruthMap(:, track.featureId); Jcost = 0; 
+%             p_f_G = groundTruthMap(:, track.featureId); Jcost = 0; 
             
             if Jcost > msckfParams.maxGNCost
                 break;

### 2014-12-23 Report/project_report.tex
Another term was missing from F
Some minor changes relating to noise

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{EXTENDED KALMAN FILTER} \label{sec:ekf}
+\subsection{Extended Kalman Filter (EKF)} \label{subsec:ekf}
 In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{SLIDING WINDOW FILTER} \label{sec:slidingwindow}
+\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
@@
     \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
-    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T\left(\Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}} \right)
+    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
 \end{align}
-where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
+where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$,
+\begin{align*}
+    \hat{\Vec{v}} &= \Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}}, & \hat{\boldsymbol{\omega}} &= \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}},
+\end{align*}
 \begin{align*}
@@
                                                                 -\hat{\boldsymbol{\omega}}^T & 0
-                                                            \ebm 
-\end{align*}
-with
-\begin{align*}
-    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}}
-\end{align*}
-and
-\begin{align*}
+                                                            \ebm ,
+                                                            & \text{and} &&
     \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
@@
 \end{align*}
+In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
 
-In our implementation we propagate the motion model using a simple forward-Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.
+We can also examine the linearized continuous-time model of the IMU error state:
+\begin{align}
+    \dot{\widetilde{\Vec{x}}}_I = \Vec{F}\widetilde{\Vec{x}}_I + \Vec{G}\Vec{n}_I
+\end{align}
+where the Jacobians $\Vec{F}$, $\Vec{G}$ are given by
+\begin{align}
+\Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                    -\hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}^\times & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3} \\
+            \ebm
+\end{align}
+\begin{align}
+\Vec{G} &=  \bbm    -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                                \Vec{0}_{3\times3} & \Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
+                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
+                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
+            \ebm ,
+\end{align}
+and $\Vec{n}_I = \bbm \Vec{n}_{\boldsymbol{\omega}}^T & \Vec{n}_{\Vec{b}_{\boldsymbol{\omega}}}^T & \Vec{n}_{\Vec{v}}^T & \Vec{n}_{\Vec{b}_{\Vec{v}}}^T \ebm^T$ is the IMU process noise, which has covariance matrix $\Vec{Q}_I$.
 
@@
 with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$.
-The Jacobians $\Vec{F}$, $\Vec{G}$ are given by
-\begin{align}
-\Vec{F} &= \bbm -\hat{\boldsymbol{\omega}}^\times & -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                    \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3} \\
-            \ebm
-\end{align}
-\begin{align}
-\Vec{G} &=  \bbm    -\Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                                \Vec{0}_{3\times3} & \Vec{1}_3 & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} \\
-                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & \Vec{1}_3 \\
-                                \Vec{0}_{3\times3} & \Vec{0}_{3\times3} & -\hat{\Vec{C}}_{I G}^T & \Vec{0}_{3\times3}
-            \ebm ,
-\end{align}
-and $\Vec{Q}_I$ is the covariance of the IMU measurement noise.
 
@@
 \end{align}
-$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \sigma^2_{\text{im}} \Vec{1}_{2}$.
+$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.
 
@@
 \end{align}
+The noise vector $\Vec{n}^{(j)}$ has covariance matrix $\Vec{R}^{(j)} = \text{diag}\left\{\Vec{R}_1^{(j)}, \hdots, \Vec{R}_{M_j}^{(j)}\right\}$.
+
 However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
@@
 Since $\Vec{H}_{f}^{(j)}$ has full column rank, $\Vec{A}$ has dimension $2M_j \times \left(2M_j - 3\right)$ and $\Vec{r}_o^{(j)}$ has dimension $\left(2M_j - 3\right) \times 1$.
-Furthermore, because $\Vec{A}$ is semi-unitary, the covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o = \sigma^2_{\text{im}}\Vec{A}^T\Vec{A} = \sigma^2_{\text{im}}\Vec{1}_{2M_j - 3}$.
+The covariance matrix of $\Vec{n}_o^{(j)}$ is given by $\Vec{R}_o^{(j)} = \Vec{A}^T\Vec{R}^{(j)}\Vec{A}$.
 
@@
 \end{align} 
-The covariance matrix of $\Vec{n}_n$ is given by $\Vec{R}_n = \Vec{Q}_1^T \Vec{R}_o \Vec{Q}_1 = \sigma^2_\text{im} \Vec{1}_r$ where $r$ is the number of columns in $\Vec{Q}_1$.
+The covariance matrix of $\Vec{n}_n$ is given by $\Vec{R}_n = \Vec{Q}_1^T \Vec{R}_o^{(j)} \Vec{Q}_1$.
 

### 2014-12-27 Matlab/swf/SlidingWindow.m
LEE HELP

@@ -29,9 +29,9 @@ lineLambda = 1;
 useMonoCamera = true; %If true, only left camera will be used
 imuPropagationOnly = false; %Test again dead-reckoning
 
-kStart = 1214;
-kEnd = 1250; 
-kappa = 25; %Sliding window size
+kStart = 1550;
+kEnd = 1650; 
+kappa = 10; %Sliding window size
 maxOptIter = 5;
 
 k1 = kStart;
@@ -81,11 +81,6 @@ end
 stateVecHistStruct = {};
 stateSigmaHistMat = [];
 
-%Keep track of triangulated landmarks
-% rho_i_pj_i_est = nan(3, 20);
-% rho_i_pj_i_est = initializeMap(initialStateStruct, y_k_j, y_var, ...
-%     calibParams, vehicleCamTransform, k1, k2);
-
 
 for k1 = kStart:kEnd    
 tic    
@@ -104,6 +99,7 @@ for k = (k1+1):k2
     totalLandmarkObs = totalLandmarkObs + sum(y_k_j(1, k, :) > -1);
 end
 totalUniqueObservedLandmarks = sum(observedBinaryFlags);
+observedLandmarkIds = find(observedBinaryFlags);
 
 
 %To initialize G-N, we propagate all the states for the first window
@@ -123,7 +119,8 @@ else
         currentStateStruct{end+1} = propagateState(currentStateStruct{end}, imuMeasurement, deltaT);
 end
 
-% Initialize the landmark positions 
+% Initialize the landmark positions (start fresh each window)
+rho_i_pj_i_est = NaN(3, numLandmarks);
 for kIdx = 1:K
         k = kIdx + k1;
         validLmObsId = find(y_k_j(1, k, :) > -1);
@@ -136,10 +133,12 @@ for kIdx = 1:K
 
             if (isnan(rho_i_pj_i_est(1, lmId)))
                 %Use triangulation to find the position of the landmark
-                 rho_pc_c = triangulate(yMeas, calibParams);
-                 rho_pi_i = kState.C_vi'*(vehicleCamTransform.C_cv'*rho_pc_c + vehicleCamTransform.rho_cv_v) +  kState.r_vi_i;
-                 rho_i_pj_i_est(:, lmId) = rho_pi_i;
-                  %rho_i_pj_i_est(:, lmId) = rho_i_pj_i(:, lmId);
[…]

### 2014-12-27 Matlab/swf/utils/updateStateStruct.m

@@ -1,4 +1,4 @@
-function [updatedStates, rho_i_pj_i_est] = updateStateStruct( currentStates, rho_i_pj_i_est, dx)
+function [updatedStates, rho_i_pj_i_est] = updateStateStruct( currentStates, lmIds, rho_i_pj_i_est, dx)
 %UPDATESTATESTRUCT Updates states by applying a step dx
 
 updatedStates = currentStates;
@@ -13,10 +13,10 @@ for stIdx = 1:numStates
 end
 
 initialIdx = numStates*6 + 1;
-for lm_i = 1:size(rho_i_pj_i_est, 2)
+for lm_i = 1:length(lmIds)
     idx = initialIdx + (lm_i-1)*3;
     dp = dx(idx:idx+2);
-    rho_i_pj_i_est(:, lm_i) = rho_i_pj_i_est(:, lm_i) + dp;
+    rho_i_pj_i_est(:, lmIds(lm_i)) = rho_i_pj_i_est(:, lmIds(lm_i)) + dp;
 end
 
 

### 2014-12-27 Matlab/swf/utils/triangulate.m
Forgot to commit a change in the triangulate function

@@ -17,4 +17,5 @@ p_fc_c = b/(u_l - u_r)*[0.5*(u_l + u_r) - c_u; ...
                         f_u/f_v*(0.5*(v_l + v_r) - c_v); ...
                         f_u];
 
+p_fc_c(1) = p_fc_c(1) + b/2;        
 end

### 2014-12-27 Report/project_report.tex

@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
-What follows is a discussion of two common solutions, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), as well as a third hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+
+In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
 

### 2014-12-27 Report/project_report.tex
Added swf math to paper

@@
 \newcommand{\ebm}{\end{bmatrix}}
+\DeclareMathAlphabet{\mbf}{OT1}{ptm}{b}{n}
+\newcommand{\mbs}[1]{{\boldsymbol{#1}}}
+\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
 
@@
 
-[BATCH MATH GOES HERE?]
+\begin{equation}
+(\mbf H^T \mbf T ^{-1} \mbf H ) \delta \mbf{x}^*  = - \mbf{H}^T\mbf{T}^{-1}\mbf{e}(\bar{\mbf{x}})
+\end{equation}
+
+\begin{equation}
+\delta\mbf{x} := \bbm \delta \mbf x_0 \\ \vdots \\ \delta \mbf x_K \\ \delta \mbf f_1 \\ \vdots \\ \delta \mbf f_N\ebm,
+\end{equation}
+where $N$ is the number of unique features observed in the window and 
+\begin{equation}
+\mathbf{H} := \begin{bmatrix}
+ -\mbf H_{\mbf x, 1} & \mbf 1 & &  & &   \\ 
+ & -\mbf G_{\mbf x, 1} &  &  &  &   -\mbf G_{\mbf f, 1} \\ \hline
+ & -\mbf H_{\mbf x, 2} & \mbf 1 & & &    \\ 
+ & & -\mbf G_{\mbf x, 2} &  & &  -\mbf G_{\mbf f, 2}   \\ \hline
+ &  & \ddots & \ddots & & \vdots  \\  \hline
+  & & & -\mbf H_{\mbf x, K} & \mbf 1 & \\ 
+ & & & & -\mbf G_{\mbf x, K} &  -\mbf G_{\mbf f, K}   \\  
+\end{bmatrix}, 
+\end{equation},
+
+\begin{equation}
+\mbf G_{\mbf x, k}^j := \pd{\mbf g}{ \mbf p}\Bigr|_{\mbf p = \bar{\mbf p}_{c_k}^{p_j c_k}} \bbm -\mbf C_{cv} \mbf C_{v_k i} & \mbf C_{cv} (\mbf C_{v_k i} (\pmb{\rho}_i^{p_j i} - \mbf r_i^{v_k i} ) )^\times  \ebm\end{equation}
+
+\begin{equation}
+\quad \mbf G_{\mbf x, k} := \bbm \mbf G_{\mbf x, k}^1 \\ \vdots \\ \mbf G_{\mbf x, k}^M \ebm
+\end{equation}
+
+\begin{equation}
+\mbf G_{\mbf f, k}^j := \pd{\mbf g}{ \mbf p}\Bigr|_{\mbf p = \bar{\mbf p}_{c_k}^{p_j c_k}} \bbm \mbf C_{cv} \mbf C_{v_k i}\ebm\end{equation}
+
+\begin{equation}
+\quad \mbf G_{\mbf f, k} := \underbrace{\bbm \mbf G_{\mbf f, k}^1 & & \\ & & \mbf G_{\mbf f, k}^2 & \\ & & \vdots &  \\ & & & \mbf G_{\mbf f, k}^M \ebm}_{\text{$j$th block column position given by feature id $n$}} \in \Real^{4M \times 3N}
+\end{equation}
+
+and,
+\begin{equation}
+\bar{\mbf p}_{c_k}^{p_j c_k} := \bbm x \\ y \\ z \ebm = \mbf C_{cv} \left( \mbf C_{v_k i} (\pmb{\rho}_i^{p_j i} - \mbf r_i^{v_k i} )   - \pmb \rho_v^{cv} \right),
+\end{equation}
+
+\begin{equation}
+\pd{\mbf g}{ \mbf p}\Bigr|_{\mbf p = \bar{\mbf p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y  \\ \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u (x-b) \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y\ebm.
+\end{equation}
+or for monocular,
+\begin{equation}
+\pd{\mbf g}{ \mbf p}\Bigr|_{\mbf p = \bar{\mbf p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y \ebm.
+\end{equation}
 

### 2014-12-27 Matlab/swf/SlidingWindow.m
Changed the history to keep track of the 2nd state instead of the first

@@ -21,15 +21,15 @@ vehicleCamTransform.C_cv = C_c_v;
 vehicleCamTransform.rho_cv_v = rho_v_c_v;
 
 %Set up sliding window
-LMLambda = 0;
+LMLambda = 1e-3;
 lineLambda = 1;
 useMonoCamera = true; %If true, only left camera will be used
 imuPropagationOnly = false; %Test again dead-reckoning
 
 kStart = 500;
-kEnd = 600; 
-kappa = 10; %Sliding window size
-maxOptIter = 15;
+kEnd = 700; 
+kappa = 20; %Sliding window size
+maxOptIter = 10;
 
 k1 = kStart;
 k2 = k1+kappa;
@@ -269,6 +269,7 @@ for kIdx = 1:K
     T(THelperIdx:(THelperIdx + Tksize - 1), THelperIdx:(THelperIdx + Tksize - 1)) = T_k;
     THelperIdx = THelperIdx + Tksize;
 end
+        H = H(:, 7:end);
 
     %Calculate scalar objective
     Jnew = 0.5*errorVector'*(T\errorVector);
@@ -289,7 +290,6 @@ end
 
     % Solve for the optimal step size!
     if optIdx <= maxOptIter
-        H = H(:, 7:end);
         dx = (H'*(T\H) + LMLambda*eye(size(H,2)))\(-H'*(T\errorVector));
         [currentStateStruct, rho_i_pj_i_est] = updateStateStruct(currentStateStruct, observedLandmarkIds, rho_i_pj_i_est,  lineLambda*dx);
     end
@@ -320,8 +320,8 @@ stateVar = diag(stateCov);
 if ~all(stateVar > 0)
     warning('Variances not positive');
 end
-stateVecHistStruct{k1 - kStart + 1} = currentStateStruct{1};
-stateSigmaHistMat(:,k1 - kStart + 1) = sqrt(abs(stateVar(1:6)));
+stateVecHistStruct{k1 - kStart + 1} = currentStateStruct{2};
+stateSigmaHistMat(:,k1 - kStart + 1) = sqrt(abs(stateVar(7:12)));
 toc
 end %End Sliding window
 
@@ -347,8 +347,8 @@ for stIdx = 1:length(stateVecHistStruct)
     rotErrVec(:, stIdx) = [eRotMat(3,2); eRotMat(1,3); eRotMat(2,1)];
 end
 
-transLim = 0.5;
[…]

### 2014-12-27 Report/project_report.tex

@@
 \begin{equation}
-\mathbf{H} := \begin{bmatrix}
+\mathbf{H} := \left[\begin{array}{cccccc}
  -\mbf H_{\mbf x, 1} & \mbf 1 & &  & &   \\ 
@@
  & & & & -\mbf G_{\mbf x, K} &  -\mbf G_{\mbf f, K}   \\  
-\end{bmatrix}, 
+\end{array}\right], 
 \end{equation},
@@
 We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
-In order to accommodate this alternative state parametrization, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
+In order to accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
 

### 2014-12-29 Report/project_report.tex
Cleaned up MSCKF triangulation math notation

--- 2014-12-27 Report/project_report.tex
@@
 
-[BATCH MATH GOES HERE?]
+\begin{equation}
+(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
+\end{equation}
+
+\begin{equation}
+\delta\Vec{x} := \bbm \delta \Vec x_0 \\ \vdots \\ \delta \Vec x_K \\ \delta \Vec f_1 \\ \vdots \\ \delta \Vec f_N\ebm,
+\end{equation}
+where $N$ is the number of unique features observed in the window and 
+\begin{equation}
+\mathbf{H} := \left[\begin{array}{cccccc}
+ -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
+ & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
+ & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
+ & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
+ &  & \ddots & \ddots & & \vdots  \\  \hline
+  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
+ & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
+\end{array}\right], 
+\end{equation},
+
+\begin{equation}
+\Vec G_{\Vec{x}, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm -\Vec C_{cv} \Vec C_{v_k i} & \Vec C_{cv} (\Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} ) )^\times  \ebm\end{equation}
+
+\begin{equation}
+\quad \Vec G_{\Vec x, k} := \bbm \Vec G_{\Vec x, k}^1 \\ \vdots \\ \Vec G_{\Vec x, k}^M \ebm
+\end{equation}
+
+\begin{equation}
+\Vec G_{\Vec f, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm \Vec C_{cv} \Vec C_{v_k i}\ebm\end{equation}
+
+\begin{equation}
+\quad \Vec G_{\Vec f, k} := \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature id $n$}} \in \Real^{4M \times 3N}
+\end{equation}
+
+and,
+\begin{equation}
+\bar{\Vec p}_{c_k}^{p_j c_k} := \bbm x \\ y \\ z \ebm = \Vec C_{cv} \left( \Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} )   - \boldsymbol \rho_v^{cv} \right),
+\end{equation}
+
+\begin{equation}
+\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y  \\ \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u (x-b) \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y\ebm.
+\end{equation}
+or for monocular,
+\begin{equation}
+\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y \ebm.
+\end{equation}
 
@@
 \subsection{Feature Position Estimation}
-When a feature $f_j$ goes out of view of the camera, the MSCKF generates a batch-optimized estimate $\hat{\Vec{p}}_G^{f_j G}$ of the feature's position in the global frame.
+When a feature $f_j$ goes out of view of the camera, the MSCKF estimates its position $\hat{\Vec{p}}_G^{f_j G}$ using an inverse-depth least-squares Gauss-Newton optimization. 
+The procedure takes as input $N$ camera poses and $N$ sets of ``ideal'' pixel measurements, where ``ideal'' means that the pixel measurements have been corrected for the camera intrinsics:
+\begin{align}
+\hat{\Vec{z}}_i^{(j)} &= \bbm u'_i & v'_i \ebm ^T = \bbm (u_i - c_u)/f_u & (v_i - c_v)/f_v \ebm ^T
+\end{align}
+We initialize the optimization by estimating the position of feature $f_j$ in camera frame $C_1$ using a linear least squares method with measurements from the first two camera frames, $C_1$ and $C_2$:
+\begin{align}
+\hat{\Vec{p}}_{C_1}^{f_j C_1} := \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & \hat{Z}^{(j)}_{C_1} \ebm ^T = \lambda \boldsymbol{\rho}_{C_1}^{f_j}
+\end{align}
+where
+\begin{align}
+\boldsymbol{\rho}_{C_i}^{f_j} &:= \frac{1}{\sqrt{u'^2_i + v'^2_i + 1}}\bbm u'_i & v'_i & 1\ebm^T
+\end{align}
+and
+\begin{align}
+\lambda &= \left[(\Vec A^T \Vec A)^{-1} \Vec A^T \hat{\Vec{p}}_{C_1}^{C_2 C_1} \right]_1
+\end{align}
+with
+\begin{align}
+\Vec A &:= \bbm \boldsymbol{\rho}_{C_1}^{f_j} & -\boldsymbol{\rho}_{C_2}^{f_j} \ebm.
+\end{align}
+
+Next, we define a vector of three parameters:
+\begin{align}
+\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
+\end{align}
+The feature position in camera frame $C_i$ can be expressed as
+\begin{align}
+\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
+\end{align}
+With our state parameters, we can use this equation to define three functions:
+\begin{align}
+    \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
+    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
+    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
+            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
+            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
+        \ebm 
+\end{align}
+The camera measurement error can then be written as
+\begin{align}
+\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
+\end{align}
+The least squares linear system then becomes
+\begin{align}
+(\Vec  E^T \Vec W ^{-1} \Vec E ) \delta \Vec{y}^*  = - \Vec{E}^T\Vec{W}^{-1}\Vec{e}(\hat{\Vec{y}})
+\end{align}
+where
+\begin{align}
+\Vec E &= \frac{\partial \Vec e}{\partial \hat{\Vec{y}}}
+& &
+\text{and}
+&
+\Vec W &= \bbm \Vec{R}_1^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N^{(j)} \ebm 
+\end{align}
+with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.
+
+The Jacobian elements are given by
+\begin{align}
+\frac{\partial \Vec e}{\partial \alpha} &= 
+    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_1 \\[0.5em]
+                        -\frac{\partial h_2}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_2 \ebm
+\\[0.5em]
+\frac{\partial \Vec e}{\partial \beta} &= 
+    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_1 \\[0.5em] 
+                        -\frac{\partial h_2}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_2 \ebm
+\\[0.5em]
+\frac{\partial \Vec e}{\partial \gamma} &= 
+    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_1 \\[0.5em] 
+                        -\frac{\partial h_2}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_2 \ebm
+\end{align}
+with
+\begin{align}
+\frac{\partial h_1}{\partial \alpha} &= C_{11}, & \frac{\partial h_1}{\partial \beta} &= C_{12}, & \frac{\partial h_1}{\partial \gamma} &= \hat{x}_{C_i}^{C_1 C_i} \notag \\
+\frac{\partial h_2}{\partial \alpha} &= C_{21}, & \frac{\partial h_2}{\partial \beta} &= C_{22}, & \frac{\partial h_2}{\partial \gamma} &= \hat{y}_{C_i}^{C_1 C_i} \\
+\frac{\partial h_3}{\partial \alpha} &= C_{31}, & \frac{\partial h_3}{\partial \beta} &= C_{32}, & \frac{\partial h_3}{\partial \gamma} &= \hat{z}_{C_i}^{C_1 C_i} \notag .
+\end{align}
 

### 2014-12-29 Report/project_report.tex
Missed a few primes
Cleaned up the SWF notation

@@
 In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+We present an experimental comparison of the SWF and MSCKF using a dataset consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
 
@@
 %In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.
-
-[EKF EQUATIONS HERE?]
 
@@
 The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
-
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
-\delta\Vec{x} := \bbm \delta \Vec x_0 \\ \vdots \\ \delta \Vec x_K \\ \delta \Vec f_1 \\ \vdots \\ \delta \Vec f_N\ebm,
+\delta\Vec{x} := \bbm \delta \Vec x_0 & \hdots & \delta \Vec x_K & \delta \Vec f_1 & \hdots & \delta \Vec f_N\ebm ^T,
 \end{equation}
-where $N$ is the number of unique features observed in the window and 
+where $N$ is the number of unique features observed in the window.
+The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
 \begin{equation}
-\mathbf{H} := \left[\begin{array}{cccccc}
+\mathbf{H} = \left[\begin{array}{cccccc}
  -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
@@
 \end{array}\right], 
-\end{equation},
-
+\end{equation}
+where
+\begin{align}
+\Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
+    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\boldsymbol{\rho}_i^{f_j G} - \Vec r_G^{IG,k} ) )^\times  \ebm ^T
+\\[0.5em]
+\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
+\\[0.5em]
+\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
+\\[0.5em]
+\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
+\end{align}
+with
 \begin{equation}
-\Vec G_{\Vec{x}, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm -\Vec C_{cv} \Vec C_{v_k i} & \Vec C_{cv} (\Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} ) )^\times  \ebm\end{equation}
-
-\begin{equation}
-\quad \Vec G_{\Vec x, k} := \bbm \Vec G_{\Vec x, k}^1 \\ \vdots \\ \Vec G_{\Vec x, k}^M \ebm
+\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\boldsymbol{\rho}_G^{f_j G} - \Vec r_G^{IG,k} )   - \boldsymbol \rho_I^{CI} \right),
 \end{equation}
@@
 \begin{equation}
-\Vec G_{\Vec f, k}^j := \pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} \bbm \Vec C_{cv} \Vec C_{v_k i}\ebm\end{equation}
-
-\begin{equation}
-\quad \Vec G_{\Vec f, k} := \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature id $n$}} \in \Real^{4M \times 3N}
-\end{equation}
-
-and,
-\begin{equation}
-\bar{\Vec p}_{c_k}^{p_j c_k} := \bbm x \\ y \\ z \ebm = \Vec C_{cv} \left( \Vec C_{v_k i} (\boldsymbol{\rho}_i^{p_j i} - \Vec r_i^{v_k i} )   - \boldsymbol \rho_v^{cv} \right),
-\end{equation}
-
-\begin{equation}
-\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y  \\ \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u (x-b) \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y\ebm.
-\end{equation}
-or for monocular,
-\begin{equation}
-\pd{\Vec g}{ \Vec p}\Bigr|_{\Vec p = \bar{\Vec p}_{c_k}^{p_j c_k}} = \bbm \frac{f_u}{z} & 0 & -\frac{1}{z^2} f_u x \\  0 & \frac{f_v}{z} & -\frac{1}{z^2} f_v y \ebm.
+\pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} = \frac{1}{z^2} \bbm f_u z & 0 & - f_u x \\  0 & f_v z & -f_v y \ebm.
 \end{equation}
@@
 \begin{align} \label{eq:imucovar}
-    \dot{\hat{\Vec{P}}}_{II}\left(t_k, t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \notag \\ 
-                & + \hat{\Vec{P}}_{II}\left(t_k, t_k  + \tau \right) \Vec{F}^T \notag \\ 
+    \dot{\hat{\Vec{P}}}_{II}\left(t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) 
+                \hat{x}_{C_i}^{C_1 C_i} + \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) \Vec{F}^T \notag \\ 
                 & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
 \end{align}
-with the initial condition $\hat{\Vec{P}}_{II}\left(t_k, t_k\right) = \hat{\Vec{P}}_{II,k}$.
+with the initial condition $\hat{\Vec{P}}_{II}\left(t_k\right) = \hat{\Vec{P}}_{II,k}$.
 
@@
 &
-\Vec W &= \bbm \Vec{R}_1^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N^{(j)} \ebm 
+\Vec W &= \bbm \Vec{R}_1'^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N'^{(j)} \ebm 
 \end{align}
-with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.
+with $\Vec{R}_i'^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
 

### 2014-12-29 Matlab/reportPlots.m
Added units to plots

@@ -19,7 +19,7 @@ xlim([t(kStart) t(kEnd)]);
 ylim([-transLim transLim])
 legend('MSCKF', 'SWF');
 title(sprintf('Translational Error (MSCKF vs. SWF)'))
-ylabel('\delta r_x')
+ylabel('\delta r_x [m]')
 
 
 subplot(3,1,2)
@@ -28,7 +28,7 @@ hold on
 plot(t(kStart:kEnd), swf_trans_err(2,:), '-g', 'LineWidth', 1.2)
 xlim([t(kStart) t(kEnd)]);
 ylim([-transLim transLim])
-ylabel('\delta r_y')
+ylabel('\delta r_y [m]')
 
 subplot(3,1,3)
 plot(t(kStart:kEnd), msckf_trans_err(3,:), '-b', 'LineWidth', 1.2)
@@ -36,8 +36,8 @@ hold on
 plot(t(kStart:kEnd), swf_trans_err(3,:), '-g', 'LineWidth', 1.2)
 xlim([t(kStart) t(kEnd)]);
 ylim([-transLim transLim])
-ylabel('\delta r_z')
-xlabel('t_k')
+ylabel('\delta r_z [m]')
+xlabel('t_k [s]')
 %set(gca,'FontSize',12)
 %set(findall(gcf,'type','text'),'FontSize',12)
 
@@ -69,4 +69,4 @@ plot(t(kStart:kEnd), swf_rot_err(3,:), '-g', 'LineWidth', 1.2)
 xlim([t(kStart) t(kEnd)]);
 ylim([-rotLim rotLim])
 ylabel('\delta\theta_z')
-xlabel('t_k')
+xlabel('t_k [s]')

### 2014-12-29 Matlab/swf/SlidingWindow.m
Added some plots to the report. Restructured the folders

@@ -5,7 +5,7 @@ clc
 clear
 close all
 addpath('utils')
-load('../dataset3.mat')
+load('../datasets/dataset3.mat')
 
 %Set number of landmarks
 numLandmarks = size(y_k_j,3);
@@ -47,7 +47,7 @@ initialStateStruct = {};
 % Extract noise values
 Q = diag([v_var; w_var]);
 if useMonoCamera
-    R = diag(y_var(1:2));
+    R = 2*diag(y_var(1:2));
 else
     R = diag(y_var);
 end

### 2014-12-29 Matlab/nullspace_plots.m
Committing uncommitted things

@@ -14,7 +14,7 @@ figure
 subplot(3,1,1)
 plot(tPlot, msckf_trans_err(1,:), '-b', 'LineWidth', 1.2)
 hold on
-plot(tPlot, swf_trans_err(1,:), '-g', 'LineWidth', 1.2)
+plot(tPlot, swf_trans_err(1,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-transLim transLim])
 legend('With Nullspace Projection', 'Without Nullspace Projection');
@@ -25,7 +25,7 @@ ylabel('\delta r_x [m]')
 subplot(3,1,2)
 plot(tPlot, msckf_trans_err(2,:), '-b', 'LineWidth', 1.2)
 hold on
-plot(tPlot, swf_trans_err(2,:), '-g', 'LineWidth', 1.2)
+plot(tPlot, swf_trans_err(2,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-transLim transLim])
 ylabel('\delta r_y [m]')
@@ -33,7 +33,7 @@ ylabel('\delta r_y [m]')
 subplot(3,1,3)
 plot(tPlot, msckf_trans_err(3,:), '-b', 'LineWidth', 1.2)
 hold on
-plot(tPlot, swf_trans_err(3,:), '-g', 'LineWidth', 1.2)
+plot(tPlot, swf_trans_err(3,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-transLim transLim])
 ylabel('\delta r_z [m]')
@@ -45,7 +45,7 @@ figure
 subplot(3,1,1)
 plot(tPlot, msckf_rot_err(1,:), '-b', 'LineWidth', 1.2)
 hold on
-plot(tPlot, swf_rot_err(1,:), '-g', 'LineWidth', 1.2)
+plot(tPlot, swf_rot_err(1,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-rotLim rotLim])
 title(sprintf('Rotational Error'))
@@ -57,7 +57,7 @@ legend('With Nullspace Projection', 'Without Nullspace Projection');
 subplot(3,1,2)
[…]

### 2014-12-29 Matlab/simulateLandmarks/augmentDataset.m

@@ -13,7 +13,7 @@ calibParams.f_v = fv;
 calibParams.b = b;
 
 %Generate new landmarks
-newLmNum = 40;
+newLmNum = 100;
 % xRange = linspace(min(rho_i_pj_i(1,:)), max(rho_i_pj_i(1,:)), newLmNum);
 % yRange = linspace(min(rho_i_pj_i(2,:)), max(rho_i_pj_i(2,:)), newLmNum);
 % zRange = linspace(min(rho_i_pj_i(3,:)), max(rho_i_pj_i(3,:)), newLmNum);
@@ -21,7 +21,7 @@ newLmNum = 40;
 
 xRange = [min(rho_i_pj_i(1,:)) - 5, max(rho_i_pj_i(1,:)) + 5];
 yRange = [min(rho_i_pj_i(2,:)) - 5, max(rho_i_pj_i(2,:)) + 5];
-zRange = [min(rho_i_pj_i(3,:)) - 5, max(rho_i_pj_i(3,:)) - 5];
+zRange = [min(rho_i_pj_i(3,:)) - 5, max(rho_i_pj_i(3,:)) ];
 
 newLmPosX = range(xRange)*rand(1,newLmNum) + xRange(1);
 newLmPosY = range(yRange)*rand(1,newLmNum) + yRange(1);
@@ -53,4 +53,4 @@ for k = 1:length(t)
 end
 
 
-save('../dataset3_fresh2.mat');
+save('../dataset3_fresh3.mat');

### 2014-12-29 Matlab/plotting/reportPlots.m
removed personal addpath

@@ -1,7 +1,6 @@
 load('swf_est25.mat');
 load('msckf_est.mat');
-load('datasets/dataset3.mat');
-addpath('/Users/valentinp/Research/MATLAB/export_fig'); %Use Oliver Woodford's awesome export_fig package to get trimmed PDFs
+load('../datasets/dataset3.mat');
 kStart = 500;
 kEnd = 1000;
 swf_trans_err = transErrVec;

### 2014-12-29 Matlab/plotting/reportPlots.m
Added minor grid lines

@@ -22,6 +22,7 @@ title(sprintf('Translational Error (MSCKF vs. SWF)'))
 ylabel('\delta r_x [m]')
 set(gca,'FontSize',fontSize)
 grid on
+grid minor
 
 subplot(3,1,2)
 plot(t(kStart:kEnd), msckf_trans_err(2,:), '-b', 'LineWidth', 1.2)
@@ -32,6 +33,7 @@ ylim([-transLim transLim])
 ylabel('\delta r_y [m]')
 set(gca,'FontSize',fontSize)
 grid on
+grid minor
 
 
 subplot(3,1,3)
@@ -45,6 +47,7 @@ xlabel('t_k [s]')
 set(gca,'FontSize',fontSize)
 set(findall(gcf,'type','text'),'FontSize',fontSize)
 grid on
+grid minor
 
 filename = sprintf('transError.pdf');
 export_fig(gcf, filename, '-transparent');
@@ -61,6 +64,7 @@ ylabel('\delta\theta_x')
 legend('MSCKF', 'SWF');
 set(gca,'FontSize',fontSize)
 grid on
+grid minor
 
 
  
@@ -73,6 +77,7 @@ ylim([-rotLim rotLim])
 ylabel('\delta\theta_y')
 set(gca,'FontSize',fontSize)
 grid on
+grid minor
 
 subplot(3,1,3)
 plot(t(kStart:kEnd), msckf_rot_err(3,:), '-b', 'LineWidth', 1.2)
@@ -83,6 +88,7 @@ ylim([-rotLim rotLim])
 ylabel('\delta\theta_z')
 xlabel('t_k [s]')
 grid on
+grid minor
 
 set(gca,'FontSize',fontSize)
 set(findall(gcf,'type','text'),'FontSize',fontSize)

### 2014-12-29 Matlab/plotting/nullspace_plots.m
Gridlines yo

@@ -9,6 +9,7 @@ swf_rot_err = theta_CG_err_NSToff;
 
 transLim = 0.4;
 rotLim = 0.4;
+fontSize = 14;
 
 figure
 subplot(3,1,1)
@@ -20,7 +21,8 @@ ylim([-transLim transLim])
 legend('With Nullspace Projection', 'Without Nullspace Projection');
 title(sprintf('Translational Error'))
 ylabel('\delta r_x [m]')
-
+set(gca,'FontSize',fontSize)
+grid on; grid minor;
 
 subplot(3,1,2)
 plot(tPlot, msckf_trans_err(2,:), '-b', 'LineWidth', 1.2)
@@ -29,6 +31,8 @@ plot(tPlot, swf_trans_err(2,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-transLim transLim])
 ylabel('\delta r_y [m]')
+set(gca,'FontSize',fontSize)
+grid on; grid minor;
 
 subplot(3,1,3)
 plot(tPlot, msckf_trans_err(3,:), '-b', 'LineWidth', 1.2)
@@ -40,6 +44,8 @@ ylabel('\delta r_z [m]')
 xlabel('t_k [s]')
 %set(gca,'FontSize',12)
 %set(findall(gcf,'type','text'),'FontSize',12)
+set(gca,'FontSize',fontSize)
+grid on; grid minor;
 
 figure
 subplot(3,1,1)
@@ -51,7 +57,8 @@ ylim([-rotLim rotLim])
 title(sprintf('Rotational Error'))
 ylabel('\delta\theta_x')
 legend('With Nullspace Projection', 'Without Nullspace Projection');
-
+set(gca,'FontSize',fontSize)
+grid on; grid minor;
 
  
 subplot(3,1,2)
@@ -61,6 +68,8 @@ plot(tPlot, swf_rot_err(3,:), '-r', 'LineWidth', 1.2)
 xlim([tPlot(1) tPlot(end)]);
 ylim([-rotLim rotLim])
 ylabel('\delta\theta_y')
+set(gca,'FontSize',fontSize)
+grid on; grid minor;
 
 subplot(3,1,3)
 plot(tPlot, msckf_rot_err(3,:), '-b', 'LineWidth', 1.2)
@@ -70,3 +79,5 @@ xlim([tPlot(1) tPlot(end)]);
[…]

### 2014-12-29 Report/project_report.tex
Added some plots to the report. Restructured the folders
Lovely plots
Added minor grid lines
Gridlines yo
Figure cleanup

@@
 However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
-Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]
+Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly.
+Figure \ref{fig:nullspace} shows the effect of attempting to use this error directly in the EKF.
+The correlation between the measurement error on the state causes the filter to behave very poorly after a small number of time steps.
+
+\begin{figure}
+    \centering
+    \subfigure {
+        \includegraphics[width=0.45\textwidth]{figs/NST_trans}
+        \label{fig:nullspace:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.45\textwidth]{figs/NST_rot}
+        \label{fig:nullspace:rot_err}
+    }
+
+    \caption{Effect of nullspace projection on MSCKF performance}
+    \label{fig:nullspace}
+\end{figure}
 
@@
     \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/Comparison-Trans-Error}
+        \label{fig:comp_1:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/Comparison-Rot-Error}
+        \label{fig:comp_1:rot_err}
+    }
+
+    \caption{MSCKF vs. SWF}
+    \label{fig:comp_1}
+\end{figure}
+
+\begin{figure}
+    \centering
     \includegraphics[width=0.45\textwidth]{figs/sensorhead}

### 2014-12-29 Matlab/msckf/MSCKF.m
Added MSCKF and IMU data for plotting

@@ -13,8 +13,8 @@ load('../datasets/dataset3.mat')
 % load('../dataset3.mat')
 
 %Dataset window bounds
-kStart = 500;
-kEnd = 1000;
+% kStart = 500; kEnd = 1000;
+kStart = 1215; kEnd = 1715;
 
 %Set constant
 numLandmarks = size(y_k_j,3);

### 2014-12-29 Matlab/swf/SlidingWindow.m
Creating dem mats

@@ -28,7 +28,7 @@ imuPropagationOnly = false; %Test again dead-reckoning
 
 kStart = 500;
 kEnd = 1000; 
-kappa = 25; %Sliding window size
+kappa = 100; %Sliding window size
 maxOptIter = 20;
 
 k1 = kStart;
@@ -282,7 +282,7 @@ end
     end
     
     %Check for convergence
-    if norm(dx) < 1e-2
+    if norm(dx) < 0.5e-2
         disp('Converged!')
         break;
     end
@@ -350,7 +350,7 @@ end
 % Save estimates
 swf_trans_err = transErrVec;
 swf_rot_err = rotErrVec;
-save('swf_est.mat', 'transErrVec', 'rotErrVec');
+save(sprintf('swf_%d_%d_%d.mat',kStart,kEnd, kappa), 'swf_trans_err', 'swf_rot_err');
 
 
 transLim = 0.4;

### 2014-12-29 Matlab/plotting/reportPlots.m
Final commit for the night. 2 new sets of plots added to the pdf

@@ -30,8 +30,8 @@ plot(t(kStart:kEnd), swf100.swf_trans_err(1,:), '--g', 'LineWidth', 1.2)
 
 xlim([t(kStart) t(kEnd) ]);
 ylim([-transLim transLim])
-h_legend = legend('IMU Only','MSCKF 5-Inf', 'MSCKF 10-50', 'MSCKF 20-100', 'SWF 10', 'SWF 50', 'SWF 100','Location', 'southeast');
-set(h_legend,'FontSize',8);
+h_legend = legend('IMU Only','MSCKF 5-Inf', 'MSCKF 10-50', 'MSCKF 20-100', 'SWF 10', 'SWF 50', 'SWF 100','Location', 'northwest');
+set(h_legend,'FontSize',10);
 
 title(sprintf('Translational Error (Interval 1)'))
 ylabel('\delta r_x [m]')
@@ -75,7 +75,7 @@ set(findall(gcf,'type','text'),'FontSize',fontSize)
 grid on
 grid minor
 
-filename = sprintf('6-Way-Comparison-500-100-Trans.pdf');
+filename = sprintf('6-Way-Comparison-500-1000-Trans.pdf');
 export_fig(gcf, filename, '-transparent');
 
 figure
@@ -92,8 +92,8 @@ xlim([t(kStart) t(kEnd) ]);
 ylim([-rotLim rotLim])
 title(sprintf('Rotational Error (Interval 1)'))
 ylabel('\delta\theta_x')
-h_legend = legend('IMU Only','MSCKF 5-Inf', 'MSCKF 10-50', 'MSCKF 20-100', 'SWF 10', 'SWF 50', 'SWF 100','Location', 'southeast');
-set(h_legend,'FontSize',8);
+h_legend = legend('IMU Only','MSCKF 5-Inf', 'MSCKF 10-50', 'MSCKF 20-100', 'SWF 10', 'SWF 50', 'SWF 100','Location', 'northwest');
+set(h_legend,'FontSize',10);
 set(gca,'FontSize',fontSize)
 grid on
 grid minor
@@ -135,7 +135,7 @@ grid minor
 set(gca,'FontSize',fontSize)
 set(findall(gcf,'type','text'),'FontSize',fontSize)
 
-filename = sprintf('6-Way-Comparison-500-100-Rot.pdf');
+filename = sprintf('6-Way-Comparison-500-1000-Rot.pdf');
[…]

### 2014-12-29 Report/project_report.tex
Added MSCKF and IMU data for plotting
Final commit for the night. 2 new sets of plots added to the pdf
Fixed plot axes

--- 2014-12-29 Report/project_report.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Extended Kalman Filter (EKF)} \label{subsec:ekf}
+\section{Extended Kalman Filter (EKF)} \label{sec:EKF}
 In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
+\section{Sliding Window Filter (SWF)} \label{sec:SWF}
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:msckf}
+\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:MSCKF}
 The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
@@
 &
-\Vec W &= \bbm \Vec{R}_1'^{(j)} & & \\ & \ddots & \\ & & \Vec{R}_N'^{(j)} \ebm 
+\Vec W &= \text{diag}\left\{ \Vec{R}_1'^{(j)}, \hdots, \Vec{R}_N'^{(j)} \right\}
 \end{align}
@@
 \begin{align} \label{eq:residual_stacked}
-    \Vec{r}^{(j)} &\simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
+    \Vec{r}^{(j)} &= \Vec{z}^{(j)} - \hat{\Vec{z}}^{(j)} \simeq \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G} + \Vec{n}^{(j)}
 \end{align}
@@
 However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
-Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly. [TALK HERE ABOUT WHAT HAPPENS IF YOU LEAVE IT IN]
+Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly.
+Figure \ref{fig:nullspace} shows the effect of attempting to use this error directly in the EKF.
+The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.
 
-In order to transform $\Vec{r}^{(j)}$ into a usable form for the EKF, we can define a semi-unitary matrix $\Vec{A}$ whose columns form the basis of the left nullspace of $\Vec{H}_{f}^{(j)}$, and project $\Vec{r}^{(j)}$ into this nullspace to obtain an error equation of the correct form:
+\begin{figure}
+    \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/NST_trans}
+        \label{fig:nullspace:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/NST_rot}
+        \label{fig:nullspace:rot_err}
+    }
+
+    \caption{Effect of nullspace projection on MSCKF performance}
+    \label{fig:nullspace}
+\end{figure}
+
+In order to transform $\Vec{r}^{(j)}$ into a usable form for the EKF, we can define a semi-unitary matrix $\Vec{A}$ whose columns form the basis of the left nullspace of $\Vec{H}_{f}^{(j)}$, and project $\Vec{r}^{(j)}$ into this nullspace to obtain an error equation of the correct form (to first order):
 \begin{align}
-    \Vec{r}_o^{(j)} &:= \Vec{A}^T \left(\Vec{z}_i^{(j)} - \hat{\Vec{z}}_i^{(j)}\right) \\
-                    &\simeq \Vec{A}^T \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{0} + \Vec{A}^T \Vec{n}^{(j)} \\
+    \Vec{r}_o^{(j)} := \Vec{A}^T \Vec{r}^{(j)} 
+                    &\simeq \Vec{A}^T \Vec{H}_{\Vec{x}}^{(j)} \widetilde{\Vec{x}} + \Vec{0} + \Vec{A}^T \Vec{n}^{(j)} \notag \\
                     &=: \Vec{H}_{o}^{(j)} \widetilde{\Vec{x}} + \Vec{n}_o^{(j)}
@@
     \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
+        \label{fig:comp_1:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
+        \label{fig:comp_1:rot_err}
+    }
+
+    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[500,1000]$.
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
+    \label{fig:comp_1}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
+        \label{fig:comp_2:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
+        \label{fig:comp_2:rot_err}
+    }
+
+    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[1215,1715]$.
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
+    \label{fig:comp_2}
+\end{figure}
+
+\begin{figure}
+    \centering
     \includegraphics[width=0.45\textwidth]{figs/sensorhead}

### 2014-12-29 Matlab/plotting/reportPlots.m

@@ -38,6 +38,7 @@ ylabel('\delta r_x [m]')
 set(gca,'FontSize',fontSize)
 grid on
 grid minor
+box on
 
 subplot(3,1,2)
 plot(t(kStart:kEnd), imu.msckf_trans_err(2,:), '-k', 'LineWidth', 1.2)
@@ -55,7 +56,7 @@ ylabel('\delta r_y [m]')
 set(gca,'FontSize',fontSize)
 grid on
 grid minor
-
+box on
 
 subplot(3,1,3)
 plot(t(kStart:kEnd), imu.msckf_trans_err(3,:), '-k', 'LineWidth', 1.2)
@@ -69,11 +70,13 @@ plot(t(kStart:kEnd), swf100.swf_trans_err(3,:), '--g', 'LineWidth', 1.2)
 
 xlim([t(kStart) t(kEnd) ]);
 ylim([-transLim transLim])
+ylabel('\delta r_z [m]')
 xlabel('t_k [s]')
 set(gca,'FontSize',fontSize)
 set(findall(gcf,'type','text'),'FontSize',fontSize)
 grid on
 grid minor
+box on
 
 filename = sprintf('6-Way-Comparison-500-1000-Trans.pdf');
 export_fig(gcf, filename, '-transparent');
@@ -97,7 +100,7 @@ set(h_legend,'FontSize',10);
 set(gca,'FontSize',fontSize)
 grid on
 grid minor
-
+box on
 
  
 subplot(3,1,2)
@@ -115,6 +118,7 @@ ylabel('\delta\theta_y')
 set(gca,'FontSize',fontSize)
 grid on
 grid minor
+box on
 
 subplot(3,1,3)
 plot(t(kStart:kEnd), imu.msckf_rot_err(3,:), '-k', 'LineWidth', 1.2)
@@ -131,6 +135,7 @@ ylabel('\delta\theta_z')
 xlabel('t_k [s]')
 grid on
 grid minor
+box on
 
 set(gca,'FontSize',fontSize)
 set(findall(gcf,'type','text'),'FontSize',fontSize)
@@ -149,9 +154,8 @@ msckf20 = load('msckf_1215_1715_min20_max100');
 imu = load('imu_1215_1715.mat');
 
 load('../datasets/dataset3.mat');
-kStart = 500;
-kEnd = 1000;
-
+kStart = 1215;
+kEnd = 1715;
 
 transLim = 1;
 rotLim = 0.5;
[…]

### 2014-12-29 Report/project_report.tex
Added plots for number of visible features in each interval

@@
 \section{EXPERIMENTS} \label{sec:experiments}
-We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
+We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
 The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
+This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figs/sensorhead}
+    \caption{The sensor head used in our experiments.
+The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
+In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
+    \label{fig:sensorhead}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
+    \caption{Vicon ground truth for sensor head motion (blue) and feature positions (red) in the ``Starry Night'' dataset.}
+    \label{fig:trajectory_groundtruth}
+\end{figure}
+
+\subsection{Interval 1: Many visible features}
+In the first experiment, we compared the SWF and MSCKF for various parameter settings on an interval with many visible features.
+Figure \ref{fig:nVisible_int1} shows the number of visible features in the first interval.
+At least three features are visible in $64\%$ of this interval, and the number of visible features is frequently in excess of ten.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figs/nVisible_500_1000}
+    \caption{Number of visible features in the interval $k\in[500,1000]$.
+Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
+    \label{fig:nVisible_int1}
+\end{figure}
 
@@
     \label{fig:comp_1}
+\end{figure}
+
+\subsection{Interval 2: Fewer visible features}
+In the second experiment, we compared the SWF and MSCKF for various parameter settings on an interval with few visible features.
+Figure \ref{fig:nVisible_int2} shows the number of visible features in the second interval.
+At least three features are visible in only $56\%$ of this interval, and the number of visible features rarely exceeds ten.
+Moreover, the features that are visible do not remain in view of the camera for as long a time as in the first interval.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figs/nVisible_1215_1715}
+    \caption{Number of visible features in the interval $k\in[1215,1715]$.
+Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
+    \label{fig:nVisible_int2}
 \end{figure}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/sensorhead}
-    \caption{The sensor head used in our experiments.
-The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
-In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
-    \label{fig:sensorhead}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
-    \caption{Vicon ground truth for sensor head motion (blue) and feature positions (red) in the ``Starry Night'' dataset.}
-    \label{fig:trajectory_groundtruth}
-\end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{CONCLUSIONS} \label{sec:conclusions}
 Things quickly got out of hand when Matlab achieved sentience. 
@@
 If this ever gets published, please send help!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{CONCLUSIONS} \label{sec:conclusions}
 

### 2014-12-29 Report/project_report.tex
Quick commit before taking off. Added a few words to plot captions.

@@
 
-Monkey Style Chinese Kung Fu
+Monkey Style Chinese Kung Fu vs. SWFfer Wet Jet
 
@@
 
-    \caption{Effect of nullspace projection on MSCKF performance}
+    \caption{Effect of nullspace projection on MSCKF performance.
+The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.}
     \label{fig:nullspace}
@@
 
-\subsection{Interval 1: Many visible features}
+\subsection{Experiment 1: Many visible features}
 In the first experiment, we compared the SWF and MSCKF for various parameter settings on an interval with many visible features.
@@
     \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[500,1000]$.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
     \label{fig:comp_1}
@@
 
-\subsection{Interval 2: Fewer visible features}
+\subsection{Experiment 2: Fewer visible features}
 In the second experiment, we compared the SWF and MSCKF for various parameter settings on an interval with few visible features.
@@
     \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[1215,1715]$.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature tracks lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
     \label{fig:comp_2}

### 2014-12-30 Matlab/msckf/MSCKF.m
Added swf on less noisy

@@ -13,8 +13,8 @@ addpath('utils');
 % load('../datasets/dataset3.mat')
 % load('../datasets/dataset3_fresh_10noisy.mat')
 % load('../datasets/dataset3_fresh_20noisy.mat')
-load('../datasets/dataset3_fresh_40noisy.mat')
-% load('../datasets/dataset3_fresh_100noisy.mat')
+%load('../datasets/dataset3_fresh_40noisy.mat')
+ load('../datasets/dataset3_fresh_100noisy.mat')
 
 %Dataset window bounds
 kStart = 500; kEnd = 1000;

### 2014-12-30 Matlab/swf/SlidingWindow.m

@@ -5,7 +5,7 @@ clc
 clear
 close all
 addpath('utils')
-fileName = '100noisy';
+fileName = '60lessnoisy';
 load(['../datasets/dataset3_fresh_' fileName '.mat'])
 
 
@@ -49,7 +49,8 @@ initialStateStruct = {};
 % Extract noise values
 Q = diag([v_var; w_var]);
 if useMonoCamera
-    R = diag(y_var(1:2));
+    R = eye(2);
+    %R = diag(y_var(1:2));
 else
     R = diag(y_var);
 end

### 2014-12-30 Report/project_report.tex

@@
 &
-\Vec W &= \text{diag}\left\{ \Vec{R}_1'^{(j)}, \hdots, \Vec{R}_N'^{(j)} \right\}
+\Vec W &= \text{diag}\left\{ \Vec{R}_1^{(j)}, \hdots, \Vec{R}_N^{(j)} \right\}
 \end{align}
-with $\Vec{R}_i'^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
+with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
 
@@
 \end{align}
-$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_u, \sigma^2_v\right\}$.
+$\Vec{n}_i^{(j)}$ is a zero-mean Gaussian noise term with covariance matrix $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
 

### 2014-12-30 Report/project_report.tex
Fixed merge

@@
 
+\begin{figure}
+    \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-Feat-Density-1215-1715-Trans}
+        \label{fig:comp_3:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-Feat-Density-1215-1715-Rot}
+        \label{fig:comp_3:rot_err}
+    }
+
+    \caption{Comparison of MSCKF, SWF, and IMU integration for three different synthetic maps generated based on the imu data in the dataset interval $k\in[1215,1715]$.
+The numbers next to ``MSCKF'' and ``SWF'' refer to the amount of features in the dataset (20, 40 or 60).}
+    \label{fig:comp_3}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-12-30 Matlab/plotting/reportPlots.m

@@ -332,7 +332,169 @@ set(findall(gcf,'type','text'),'FontSize',fontSize)
 filename = sprintf('6-Way-Comparison-1215-1715-Rot.pdf');
 export_fig(gcf, filename, '-transparent');
 
-%% Fresh2 dataset
+%% Less Noisy Dataset
 
-msckf_1215_1715_min20_max100_fresh2
-swf_1215_1715_50_fresh2
+
+swf20 = load('swf_1215_1715_50_20lessnoisy.mat');
+swf40 = load('swf_1215_1715_50_40lessnoisy.mat');
+swf60 = load('swf_1215_1715_50_60lessnoisy.mat');
+msckf20 = load('msckf_1215_1715_min20_max100_20lessnoisy.mat');
+msckf40 = load('msckf_1215_1715_min20_max100_40lessnoisy.mat');
+msckf60 = load('msckf_1215_1715_min20_max100_60lessnoisy.mat');
+imu = load('imu_1215_1715.mat');
+
+load('../datasets/dataset3.mat');
+kStart = 1215;
+kEnd = 1715;
+plotAbsErr = true;
+transLim = [-1 1];
+rotLim = [-0.5 0.5];
+fontSize = 14;
+
+if plotAbsErr
+    imu.msckf_trans_err = abs(imu.msckf_trans_err);
+    msckf20.msckf_trans_err = abs(msckf20.msckf_trans_err);
+    msckf40.msckf_trans_err = abs(msckf40.msckf_trans_err);
+    msckf60.msckf_trans_err = abs(msckf60.msckf_trans_err);
+    swf20.swf_trans_err = abs(swf20.swf_trans_err);
+    swf40.swf_trans_err = abs(swf40.swf_trans_err);
+    swf60.swf_trans_err = abs(swf60.swf_trans_err);
+
+    imu.msckf_rot_err = abs(imu.msckf_rot_err);
+    msckf20.msckf_rot_err = abs(msckf20.msckf_rot_err);
+    msckf40.msckf_rot_err = abs(msckf40.msckf_rot_err);
+    msckf60.msckf_rot_err = abs(msckf60.msckf_rot_err);
+    swf20.swf_rot_err = abs(swf20.swf_rot_err);
[…]

### 2014-12-30 Report/project_report.tex
Added RMSE comparisons

@@
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-Feat-Density-1215-1715-Trans}
+        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
         \label{fig:comp_3:trans_err}
@@
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-Feat-Density-1215-1715-Rot}
+        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
         \label{fig:comp_3:rot_err}

### 2014-12-30 Matlab/msckf/MSCKF.m

@@ -10,14 +10,14 @@ close all;
 clc;
 addpath('utils');
 
-% load('../datasets/dataset3.mat')
+%load('../datasets/dataset3.mat')
 % load('../datasets/dataset3_fresh_10noisy.mat')
 % load('../datasets/dataset3_fresh_10lessnoisy.mat')
-% load('../datasets/dataset3_fresh_20lessnoisy.mat')
-% load('../datasets/dataset3_fresh_40lessnoisy.mat')
-load('../datasets/dataset3_fresh_60lessnoisy.mat')
-% load('../datasets/dataset3_fresh_80lessnoisy.mat')
-% load('../datasets/dataset3_fresh_100lessnoisy.mat')
+%load('../datasets/dataset3_fresh_20lessnoisy.mat')
+%load('../datasets/dataset3_fresh_40lessnoisy.mat')
+%load('../datasets/dataset3_fresh_60lessnoisy.mat')
+%load('../datasets/dataset3_fresh_80lessnoisy.mat')
+load('../datasets/dataset3_fresh_20lessnoisy.mat')
 
 %Dataset window bounds
 % kStart = 500; kEnd = 1000;
@@ -46,7 +46,7 @@ noiseParams.Q_imu = 0.1*diag([w_var', 1e-4*ones(1,3), v_var', 1e-4*ones(1,3)]);
 noiseParams.initialIMUCovar = 1e-4 * eye(12);
     
 %MSCKF parameters
-msckfParams.minTrackLength = 10;     % Set to inf to dead-reckon only
+msckfParams.minTrackLength = 5;     % Set to inf to dead-reckon only
 msckfParams.maxTrackLength = 50;     % Set to inf to wait for features to go out of view
 msckfParams.maxGNCost      = 10;     % Set to inf to allow any triangulation, no matter how bad
 msckfParams.minRCOND       = 0;

### 2014-12-30 Matlab/plotting/reportPlots.m

@@ -335,38 +335,38 @@ export_fig(gcf, filename, '-transparent');
 %% Less Noisy Dataset
 
 
-swf20 = load('swf_1215_1715_50_20lessnoisy.mat');
-swf40 = load('swf_1215_1715_50_40lessnoisy.mat');
-swf60 = load('swf_1215_1715_50_60lessnoisy.mat');
 msckf20 = load('msckf_1215_1715_min20_max100_20lessnoisy.mat');
-msckf40 = load('msckf_1215_1715_min20_max100_40lessnoisy.mat');
+%msckf40 = load('msckf_1215_1715_min20_max100_40lessnoisy.mat');
 msckf60 = load('msckf_1215_1715_min20_max100_60lessnoisy.mat');
+%msckf80 = load('msckf_1215_1715_min20_max100_60lessnoisy.mat');
+msckf100 = load('msckf_1215_1715_min20_max100_100lessnoisy.mat');
+
 imu = load('imu_1215_1715.mat');
 
 load('../datasets/dataset3.mat');
 kStart = 1215;
 kEnd = 1715;
 plotAbsErr = true;
-transLim = [-1 1];
-rotLim = [-0.5 0.5];
+transLim = [0 1];
+rotLim = [0 1];
 fontSize = 14;
 
 if plotAbsErr
     imu.msckf_trans_err = abs(imu.msckf_trans_err);
     msckf20.msckf_trans_err = abs(msckf20.msckf_trans_err);
-    msckf40.msckf_trans_err = abs(msckf40.msckf_trans_err);
+    msckf100.msckf_trans_err = abs(msckf100.msckf_trans_err);
     msckf60.msckf_trans_err = abs(msckf60.msckf_trans_err);
-    swf20.swf_trans_err = abs(swf20.swf_trans_err);
-    swf40.swf_trans_err = abs(swf40.swf_trans_err);
-    swf60.swf_trans_err = abs(swf60.swf_trans_err);
+%     swf20.swf_trans_err = abs(swf20.swf_trans_err);
+%     swf40.swf_trans_err = abs(swf40.swf_trans_err);
+%     swf60.swf_trans_err = abs(swf60.swf_trans_err);
 
[…]

### 2014-12-30 Matlab/swf/SlidingWindow.m

@@ -5,9 +5,9 @@ clc
 clear
 close all
 addpath('utils')
-%fileName = '60lessnoisy';
-%load(['../datasets/dataset3_fresh_' fileName '.mat'])
-load('../datasets/dataset3.mat');
+fileName = '40lessnoisy';
+load(['../datasets/dataset3_fresh_' fileName '.mat'])
+%load('../datasets/dataset3.mat');
 
 %Set number of landmarks
 numLandmarks = size(y_k_j,3);
@@ -28,8 +28,8 @@ lineLambda = 1;
 useMonoCamera = true; %If true, only left camera will be used
 imuPropagationOnly = false; %Test again dead-reckoning
 
-kStart = 500;
-kEnd = 1000; 
+kStart = 1215;
+kEnd = 1715; 
 kappa = 25; %Sliding window size
 maxOptIter = 25;
 
@@ -351,9 +351,9 @@ end
 
 
 % Save estimates
- swf_trans_err = transErrVec;
- swf_rot_err = rotErrVec;
- save(sprintf('swf_%d_%d_%d.mat',kStart,kEnd, kappa), 'swf_trans_err', 'swf_rot_err', 'stateSigmaHistMat');
+swf_trans_err = transErrVec;
+swf_rot_err = rotErrVec;
+save(sprintf('swf_%d_%d_%d_%s.mat',kStart,kEnd, kappa, fileName), 'swf_trans_err', 'swf_rot_err', 'stateSigmaHistMat');
 
 
 transLim = 0.5;

### 2014-12-30 Report/project_report.tex
Updated plots for new noise params and changed aspect ratios

--- 2014-12-30 Report/project_report.tex
@@
 
+\begin{figure}
+    \centering
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
+        \label{fig:comp_3:trans_err}
+    }
+
+    \subfigure {
+        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
+        \label{fig:comp_3:rot_err}
+    }
+
+    \caption{Comparison of cumulative Root Mean Squared Error (RMSE) of MSCKF, SWF, and IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
+The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (20, 40 or 60).}
+    \label{fig:comp_3}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-12-30 Matlab/msckf/MSCKF.m

@@ -10,14 +10,14 @@ close all;
 clc;
 addpath('utils');
 
-%load('../datasets/dataset3.mat')
+load('../datasets/dataset3.mat')
 % load('../datasets/dataset3_fresh_10noisy.mat')
 % load('../datasets/dataset3_fresh_10lessnoisy.mat')
 %load('../datasets/dataset3_fresh_20lessnoisy.mat')
 %load('../datasets/dataset3_fresh_40lessnoisy.mat')
 %load('../datasets/dataset3_fresh_60lessnoisy.mat')
 %load('../datasets/dataset3_fresh_80lessnoisy.mat')
-load('../datasets/dataset3_fresh_20lessnoisy.mat')
+% load('../datasets/dataset3_fresh_20lessnoisy.mat')
 
 %Dataset window bounds
 % kStart = 500; kEnd = 1000;
@@ -47,7 +47,7 @@ noiseParams.initialIMUCovar = 1e-4 * eye(12);
     
 %MSCKF parameters
 msckfParams.minTrackLength = 5;     % Set to inf to dead-reckon only
-msckfParams.maxTrackLength = 50;     % Set to inf to wait for features to go out of view
+msckfParams.maxTrackLength = Inf;     % Set to inf to wait for features to go out of view
 msckfParams.maxGNCost      = 10;     % Set to inf to allow any triangulation, no matter how bad
 msckfParams.minRCOND       = 0;
 msckfParams.doNullSpaceTrick = true;

### 2014-12-30 Matlab/plotting/reportPlots.m

@@ -504,6 +504,7 @@ fontSize = 14;
 
 
 figure
+set(gcf, 'Position', [100 100 600 200]);
 plot(t(kStart:kEnd), sqrt(sum(imu.msckf_trans_err.^2, 1)/3), '-k', 'LineWidth', 2)
 hold on
 plot(t(kStart:kEnd), sqrt(sum(msckf40.msckf_trans_err.^2, 1)/3), '-b', 'LineWidth', 2)
@@ -519,6 +520,7 @@ ylim(transLim)
 h_legend = legend('IMU Only','MSCKF 40', 'MSCKF 60', 'MSCKF 100', 'SWF 40','SWF 60','SWF 100', 'Location', 'northwest');
 set(h_legend,'FontSize',10);
 title(sprintf('Feature Density Comparison: Position RMSE'))
+xlabel('t_k [s]');
 ylabel('RMSE [m]')
 set(gca,'FontSize',fontSize)
 grid on
@@ -528,6 +530,7 @@ filename = sprintf('RMSE-Comparison-Feat-Density-1215-1715-Trans.pdf');
 export_fig(gcf, filename, '-transparent');
 
 figure
+set(gcf, 'Position', [100 100 600 200]);
 plot(t(kStart:kEnd), sqrt(sum(imu.msckf_rot_err.^2, 1)/3), '-k', 'LineWidth', 2)
 hold on
 plot(t(kStart:kEnd), sqrt(sum(msckf40.msckf_rot_err.^2, 1)/3), '-b', 'LineWidth', 2)
@@ -538,9 +541,10 @@ plot(t(kStart:kEnd), sqrt(sum(swf60.swf_rot_err.^2, 1)/3), '--g', 'LineWidth', 2
 plot(t(kStart:kEnd), sqrt(sum(swf100.swf_rot_err.^2, 1)/3), '-.g', 'LineWidth', 2)
 
 xlim([t(kStart) t(kEnd) ]);
-ylim(rotLim)
+ylim([0 0.25])
 title(sprintf('Feature Density Comparison: Rotation RMSE'))
-ylabel('RMSE')
+ylabel('RMSE [m]')
+xlabel('t_k [s]');
 h_legend = legend('IMU Only','MSCKF 40', 'MSCKF 60', 'MSCKF 100', 'SWF 40','SWF 60','SWF 100', 'Location', 'northwest');
 set(h_legend,'FontSize',10);
 set(gca,'FontSize',fontSize)

### 2014-12-30 Report/project_report.tex
Redid null space trick plots

@@
 Figure \ref{fig:nullspace} shows the effect of attempting to use this error directly in the EKF.
-The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.
+The correlation between the measurement error and the state causes the filter estimates to drift further from the true values.
 
@@
 
-    \caption{Effect of nullspace projection on MSCKF performance.
-The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.}
+    \caption{Effect of nullspace projection on MSCKF performance in a synthetic 100-feature dataset.
+The correlation between the measurement error and the state causes the filter estimates to drift further from the true values.}
     \label{fig:nullspace}

### 2014-12-30 Matlab/plotting/nullspace_plots.m

@@ -1,83 +1,132 @@
 clear all; close all; clc;
-load('msckf_NST_on.mat');
-load('msckf_NST_off.mat');
+nst_on = load('msckf_NST_on.mat');
+nst_off = load('msckf_NST_off.mat');
 
-msckf_trans_err = p_C_G_err;
-msckf_rot_err = theta_CG_err;
-swf_trans_err = p_C_G_err_NSToff;
-swf_rot_err = theta_CG_err_NSToff;
+plotAbsErr = true;
 
-transLim = 0.4;
-rotLim = 0.4;
+msckf_trans_err = nst_on.msckf_trans_err;
+msckf_rot_err = nst_on.msckf_rot_err;
+msckf_err_sigma = nst_on.err_sigma;
+swf_trans_err = nst_off.msckf_trans_err;
+swf_rot_err = nst_off.msckf_rot_err;
+swf_err_sigma = nst_off.err_sigma;
+tPlot = nst_on.tPlot;
+
+transLim = [-0.4 0.4];
+rotLim = [-0.2 0.2];
 fontSize = 14;
+
+if plotAbsErr
+    msckf_trans_err = abs(msckf_trans_err);
+    msckf_rot_err = abs(msckf_rot_err);
+    swf_trans_err = abs(swf_trans_err);
+    swf_rot_err = abs(swf_rot_err);
+    transLim = transLim - transLim(1);
+    rotLim = rotLim - rotLim(1);
+end
 
 figure
 subplot(3,1,1)
-plot(tPlot, msckf_trans_err(1,:), '-b', 'LineWidth', 1.2)
 hold on
+plot(tPlot, msckf_trans_err(1,:), '-b', 'LineWidth', 1.2)
+% plot(tPlot, -3*msckf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
+% plot(tPlot, 3*msckf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
 plot(tPlot, swf_trans_err(1,:), '-r', 'LineWidth', 1.2)
+% plot(tPlot, -3*swf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
+% plot(tPlot, 3*swf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
+
 xlim([tPlot(1) tPlot(end)]);
-ylim([-transLim transLim])
[…]

### 2014-12-30 Matlab/swf/SlidingWindow.m
Commited to committing

@@ -5,9 +5,9 @@ clc
 clear
 close all
 addpath('utils')
-fileName = '40lessnoisy';
-load(['../datasets/dataset3_fresh_' fileName '.mat'])
-%load('../datasets/dataset3.mat');
+fileName = 'dataset3';
+%load(['../datasets/dataset3_fresh_' fileName '.mat'])
+load('../datasets/dataset3.mat');
 
 %Set number of landmarks
 numLandmarks = size(y_k_j,3);
@@ -30,7 +30,7 @@ imuPropagationOnly = false; %Test again dead-reckoning
 
 kStart = 1215;
 kEnd = 1715; 
-kappa = 25; %Sliding window size
+kappa = 5i0; %Sliding window size
 maxOptIter = 25;
 
 k1 = kStart;

### 2014-12-30 Report/project_report.tex
Commited to committing
I'm all about dat plot, bout dat plot, no label

--- 2014-12-30 Report/project_report.tex
@@
 \end{align}
-The feature position in camera frame $C_i$ can be expressed as
+We can express the feature position in camera frame $C_i$ in terms of its position in camera frame $C_1$:
 \begin{align}
@@
 
-However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the error in the state.
-Since the error in the feature positions \textit{is} correlated to the error in the state, $\Vec{r}^{(j)}$ cannot be used in the EKF directly.
-Figure \ref{fig:nullspace} shows the effect of attempting to use this error directly in the EKF.
-The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.
+However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the state.
+Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state.
+Figure \ref{fig:nullspace} shows the effect of attempting to use $\Vec{r}^{(j)}$ directly in the EKF.
+The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
 
@@
 
-    \caption{Effect of nullspace projection on MSCKF performance.
-The correlation between the measurement error and the state causes the filter to behave very poorly after a small number of time steps.}
+    \caption{Effect of correlation between measurement noise and filter state on MSCKF performance.
+The correlation between the measurement error and the state causes the filter estimates to drift further from the true values than with a modified, de-corellated measurement error.}
     \label{fig:nullspace}
@@
     \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
-    \caption{Vicon ground truth for sensor head motion (blue) and feature positions (red) in the ``Starry Night'' dataset.}
+    \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
     \label{fig:trajectory_groundtruth}
@@
 
+\subsection{Experiment 3: Effect of feature density}
+Finally, we investigated the effect of feature density on the performance of both the MSCKF and SWF.
+For this experiment, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
+We retained the original IMU data, but simulated camera measurements with lower measurement noise.
+Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
+    \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
+    \label{fig:trajectory_groundtruth_100lessnoisy}
+\end{figure}
+
 \begin{figure}
@@
     \caption{Comparison of cumulative Root Mean Squared Error (RMSE) of MSCKF, SWF, and IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
-The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (20, 40 or 60).}
+The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
     \label{fig:comp_3}

### 2014-12-30 Matlab/plotting/datasetPlot.m

@@ -1,0 +1,19 @@
+% load('../datasets/dataset3.mat');
+load('../datasets/dataset3_fresh_100lessnoisy.mat');
+
+fontSize = 14;
+markerSize = 50;
+lineWidth = 1.2;
+
+figure(1); clf; hold on;
+
+scatter3(rho_i_pj_i(1,:), rho_i_pj_i(2,:), rho_i_pj_i(3,:), markerSize, 'or', 'fill');
+plot3(r_i_vk_i(1,:), r_i_vk_i(2,:), r_i_vk_i(3,:), '-b', 'LineWidth', lineWidth);
+legend('Features', 'Sensor trajectory');
+
+set(gca, 'FontSize', fontSize);
+campos([-18.7401, -5.8668, 10.6855]);
+camtarget([1.5, 1.5, 1]);
+grid on; grid minor;
+axis equal;
+xlabel('x [m]'); ylabel('y [m]'); zlabel('z [m]');

### 2014-12-30 Matlab/plotting/nullspace_plots.m

@@ -37,7 +37,7 @@ plot(tPlot, swf_trans_err(1,:), '-r', 'LineWidth', 1.2)
 
 xlim([tPlot(1) tPlot(end)]);
 ylim(transLim)
-legend('With Nullspace Projection', 'Without Nullspace Projection', 'Location','northwest');
+legend('Noise Correlated to State', 'Noise De-correlated from State', 'Location','northwest');
 title(sprintf('Absolute Translational Error'))
 ylabel('\delta r_x [m]')
 set(gca,'FontSize',fontSize)
@@ -92,7 +92,7 @@ xlim([tPlot(1) tPlot(end)]);
 ylim(rotLim)
 title(sprintf('Absolute Rotational Error'))
 ylabel('\delta\theta_x')
-legend('With Nullspace Projection', 'Without Nullspace Projection', 'Location','northwest');
+legend('Noise Correlated to State', 'Noise De-correlated from State', 'Location','northwest');
 set(gca,'FontSize',fontSize)
 grid on; grid minor;
 box on;

### 2014-12-30 Report/project_report.tex
Experimental wordsmithing

@@
 
-In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.
+In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, that combines the strengths of both.
 We present an experimental comparison of the SWF and MSCKF using a dataset consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
@@
 \section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:MSCKF}
-The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
+The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
 The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
@@
 
+Figure \ref{fig:comp_1} shows absolute translational and rotational errors for pure IMU integration and various parameter settings for the MSCKF and SWF.
+For the MSCKF, we varied both the maximum number of observations before triggering an update and the minimum number of observations for a feature track to be used in an update.
+For the SWF, we varied the number of states in the sliding window.
+These parameters did not significantly affect the accuracy of either filter, however Figure \ref{fig:comp_1} shows some small gains in accuracy for larger feature track lengths and window sizes.
+
+On this interval, the SWF slightly outperforms the MSCKF in terms of translational accuracy in the $y$- and $z$-directions, and significantly outperforms the MSCKF in terms of rotational accuracy about the $x$- and $y$-axes.
+The MSCKF does not perform much better than pure IMU integration on this interval, likely due to the overall low number of feature tracks in this dataset.
+
 \begin{figure}
@@
 Moreover, the features that are visible do not remain in view of the camera for as long a time as in the first interval.
+
+Figure \ref{fig:comp_2} shows absolute translational and rotational errors for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
+Compared to Interval 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy in the $y$-direction.
+Neither filter performs substantially better than pure IMU integration on this interval, again likely due to the overall low number of feature tracks in this dataset.
 
@@
 \subsection{Experiment 3: Effect of feature density}
-Finally, we investigated the effect of feature density on the performance of both the MSCKF and SWF.
-For this experiment, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
-We retained the original IMU data, but simulated camera measurements with lower measurement noise.
+In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
+We constructed each dataset so that the larger maps contained the same features as the smaller maps, plus additional features to make up the difference.
+Although the camera measurements in these datasets are synthetic, the IMU data are unchanged from the original dataset.
 Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
+By generating longer feature tracks and increasing the number of features visible at each time step, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.
 
@@
 \end{figure}
+
+Figure \ref{fig:comp_3} compares translational and rotational root mean squared error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
+On all three datasets, the SWF outperforms the MSCKF by a wide margin, although the MSCKF's performance improves as more features are added.
 
@@
 
-    \caption{Comparison of cumulative Root Mean Squared Error (RMSE) of MSCKF, SWF, and IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
+    \caption{Comparison of root mean squared error (RMSE) of the MSCKF, SWF, and pure IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
 The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}

### 2014-12-30 Matlab/plotting/datasetPlot.m

@@ -7,9 +7,12 @@ lineWidth = 1.2;
 
 figure(1); clf; hold on;
 
+% scatter3(rho_i_pj_i(1,1:40), rho_i_pj_i(2,1:40), rho_i_pj_i(3,1:40), markerSize, 'om', 'fill');
+% scatter3(rho_i_pj_i(1,41:60), rho_i_pj_i(2,41:60), rho_i_pj_i(3,41:60), markerSize, 'og', 'fill');
+% scatter3(rho_i_pj_i(1,61:end), rho_i_pj_i(2,61:end), rho_i_pj_i(3,61:end), markerSize, 'or', 'fill');
 scatter3(rho_i_pj_i(1,:), rho_i_pj_i(2,:), rho_i_pj_i(3,:), markerSize, 'or', 'fill');
 plot3(r_i_vk_i(1,:), r_i_vk_i(2,:), r_i_vk_i(3,:), '-b', 'LineWidth', lineWidth);
-legend('Features', 'Sensor trajectory');
+legend('Features (100)', 'Features (60)', 'Features (40)', 'Sensor trajectory');
 
 set(gca, 'FontSize', fontSize);
 campos([-18.7401, -5.8668, 10.6855]);

### 2014-12-30 Matlab/plotting/reportPlots.m
1.2'd

@@ -507,14 +507,14 @@ fontSize = 14;
 
 figure
 set(gcf, 'Position', [100 100 600 200]);
-plot(t(kStart:kEnd), sqrt(sum(imu.msckf_trans_err.^2, 1)/3), '-k', 'LineWidth', 2)
+plot(t(kStart:kEnd), sqrt(sum(imu.msckf_trans_err.^2, 1)/3), '-k', 'LineWidth', 1.2)
 hold on
-plot(t(kStart:kEnd), sqrt(sum(msckf40.msckf_trans_err.^2, 1)/3), '-b', 'LineWidth', 2)
-plot(t(kStart:kEnd), sqrt(sum(msckf60.msckf_trans_err.^2, 1)/3), '--b', 'LineWidth', 2)
-plot(t(kStart:kEnd), sqrt(sum(msckf100.msckf_trans_err.^2, 1)/3), '-.b', 'LineWidth', 2)
-plot(t(kStart:kEnd), sqrt(sum(swf40.swf_trans_err.^2, 1)/3), '-g', 'LineWidth', 2)
-plot(t(kStart:kEnd), sqrt(sum(swf60.swf_trans_err.^2, 1)/3), '--g', 'LineWidth', 2)
-plot(t(kStart:kEnd), sqrt(sum(swf100.swf_trans_err.^2, 1)/3), '-.g', 'LineWidth', 2)
+plot(t(kStart:kEnd), sqrt(sum(msckf40.msckf_trans_err.^2, 1)/3), '-b', 'LineWidth', 1.2)
+plot(t(kStart:kEnd), sqrt(sum(msckf60.msckf_trans_err.^2, 1)/3), '--b', 'LineWidth', 1.2)
+plot(t(kStart:kEnd), sqrt(sum(msckf100.msckf_trans_err.^2, 1)/3), '-.b', 'LineWidth', 1.2)
+plot(t(kStart:kEnd), sqrt(sum(swf40.swf_trans_err.^2, 1)/3), '-g', 'LineWidth', 1.2)
+plot(t(kStart:kEnd), sqrt(sum(swf60.swf_trans_err.^2, 1)/3), '--g', 'LineWidth', 1.2)
+plot(t(kStart:kEnd), sqrt(sum(swf100.swf_trans_err.^2, 1)/3), '-.g', 'LineWidth', 1.2)
 
 
 xlim([t(kStart) t(kEnd) ]);
[…]

### 2014-12-30 Report/project_report.tex
Added some conclusiony words

@@
 Figure \ref{fig:comp_3} compares translational and rotational root mean squared error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
-On all three datasets, the SWF outperforms the MSCKF by a wide margin, although the MSCKF's performance improves as more features are added.
+With additional features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
+Note that the MSCKF's performance improves as the number of features increases while the SWF's performance is not significantly affected by increasing feature count.
+This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
 
@@
 
+Table [TABLETABLETABLE] provides summary statistics of each algorithm's performance on the synthetic datasets.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{CONCLUSIONS} \label{sec:conclusions}
-Things quickly got out of hand when Matlab achieved sentience. 
-We were forced to abandon our lab and have been surviving in the harsh environment of the MarsDome for weeks.
-If this ever gets published, please send help!
+Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
+In environments with more features, both algorithms provide substantial gains over pure IMU integration, although the SWF significantly outperformed the MSCKF on the datasets we tested.
+However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.
+It is unclear from these results which algorithm would perform best in the limit of infinite features.
+
+[Something something consistency NEES]
+
+It is worth noting that although the SWF achieved superior accuracy in most cases, the computational effort it required was significantly greater than for the MSCKF.
+The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
 

### 2014-12-30 Report/project_report.tex
Added rmse and nees table

@@
 
+\begin{table*}[t]
+  \centering
+    \caption{Comparison of root mean squared error (RMSE) and Normalized Estimation Error Squared (NEES) of the MSCKF and SWF for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.}
+  \begin{tabular}{*{7}{c}}
+    \hline
+Feature Count & MSCKF Position RMSE & MSCKF Rotation RMSE & MSCKF NEES & SWF Position RMSE & SWF Rotation RMSE & SWF NEES \\ \hline
+40            & 0.2672              & 0.1378              & 10.179     & 0.1750            & 0.0495            & 2280     \\
+60            & 0.2550              & 0.1247              & 12.03      & 0.1687            & 0.0377            & 2093     \\
+100           & 0.2304              & 0.0952              & 16.755     & 0.1755            & 0.0481            & 2013     \\ \hline
+  \end{tabular}
+\end{table*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-12-30 Matlab/plotting/calculateStats.m

@@ -1,9 +1,9 @@
-swf20 = load('swf_1215_1715_50_20lessnoisy.mat');
-swf40 = load('swf_1215_1715_50_40lessnoisy.mat');
-swf60 = load('swf_1215_1715_50_60lessnoisy.mat');
-msckf20 = load('msckf_1215_1715_min20_max100_20lessnoisy.mat');
+swf40 = load('swf_1215_1715_25_40lessnoisy.mat');
+swf60 = load('swf_1215_1715_25_60lessnoisy.mat');
+swf100 = load('swf_1215_1715_25_100lessnoisy.mat');
 msckf40 = load('msckf_1215_1715_min20_max100_40lessnoisy.mat');
 msckf60 = load('msckf_1215_1715_min20_max100_60lessnoisy.mat');
+msckf100 = load('msckf_1215_1715_min20_max100_100lessnoisy.mat');
 imu = load('imu_1215_1715.mat');
 
 
@@ -14,21 +14,79 @@ mean(sqrt(sum(imu.msckf_trans_err.^2, 1)/3))
 mean(sqrt(sum(imu.msckf_rot_err.^2, 1)/3))
 
 disp('MSCKF RMSE')
-mean(sqrt(sum(msckf20.msckf_trans_err.^2, 1)/3))
 mean(sqrt(sum(msckf40.msckf_trans_err.^2, 1)/3))
 mean(sqrt(sum(msckf60.msckf_trans_err.^2, 1)/3))
+mean(sqrt(sum(msckf100.msckf_trans_err.^2, 1)/3))
 disp('===============')
-mean(sqrt(sum(msckf20.msckf_rot_err.^2, 1)/3))
 mean(sqrt(sum(msckf40.msckf_rot_err.^2, 1)/3))
 mean(sqrt(sum(msckf60.msckf_rot_err.^2, 1)/3))
+mean(sqrt(sum(msckf100.msckf_rot_err.^2, 1)/3))
 
 disp('SWF RMSE')
-mean(sqrt(sum(swf20.swf_trans_err.^2, 1)/3))
 mean(sqrt(sum(swf40.swf_trans_err.^2, 1)/3))
 mean(sqrt(sum(swf60.swf_trans_err.^2, 1)/3))
+mean(sqrt(sum(swf100.swf_trans_err.^2, 1)/3))
 disp('===============')
-mean(sqrt(sum(swf20.swf_rot_err.^2, 1)/3))
[…]

### 2014-12-30 Matlab/swf/SlidingWindow.m
Added compute times

@@ -5,10 +5,10 @@ clc
 clear
 close all
 addpath('utils')
-fileName = 'dataset3';
-%load(['../datasets/dataset3_fresh_' fileName '.mat'])
-load('../datasets/dataset3.mat');
-
+fileName = '100lessnoisy';
+load(['../datasets/dataset3_fresh_' fileName '.mat'])
+%load('../datasets/dataset3.mat');
+tic
 %Set number of landmarks
 numLandmarks = size(y_k_j,3);
 
@@ -82,8 +82,7 @@ stateVecHistStruct = {};
 stateSigmaHistMat = [];
 
 
-for k1 = kStart:kEnd    
-tic    
+for k1 = kStart:kEnd        
 k2 = k1+kappa;
 
 %How many exteroceptive measurements do we have?
@@ -324,8 +323,10 @@ if ~all(stateVar > 0)
 end
 stateVecHistStruct{k1 - kStart + 1} = currentStateStruct{2};
 stateSigmaHistMat(:,k1 - kStart + 1) = sqrt(abs(stateVar(1:6)));
-toc
+
 end %End Sliding window
+
+toc
 
 sigma_x = (stateSigmaHistMat(1,:));
 sigma_y = (stateSigmaHistMat(2,:));
@@ -336,7 +337,7 @@ sigma_th3 = (stateSigmaHistMat(6,:));
 
 
 %% Plot error and variances
-addpath('/Users/valentinp/Research/MATLAB/export_fig'); %Use Oliver Woodford's awesome export_fig package to get trimmed PDFs
+%addpath('/Users/valentinp/Research/MATLAB/export_fig'); %Use Oliver Woodford's awesome export_fig package to get trimmed PDFs
 
 rotErrVxec = zeros(3, length(stateVecHistStruct));
 transErrVec = zeros(3, length(stateVecHistStruct));

### 2014-12-30 Report/project_report.tex
Added compute times
Updated table formatting

--- 2014-12-30 Report/project_report.tex
@@
 \title{\LARGE \bf
-The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper
+The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and Sliding Window Filter
 }
@@
 
-Figure \ref{fig:comp_3} compares translational and rotational root mean squared error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
+Figure \ref{fig:comp_3} compares translational and rotational Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
 With additional features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
@@
 
-    \caption{Comparison of root mean squared error (RMSE) of the MSCKF, SWF, and pure IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
+    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
 The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
@@
 
-Table [TABLETABLETABLE] provides summary statistics of each algorithm's performance on the synthetic datasets.
+Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the synthetic datasets.
+As expected from Figure \ref{fig:comp_3}, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
+It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window.
+The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so the SWF may be considered inconsistent in the absence of a known map.
+
+Another point in favour of the MSCKF is that the computational effort required was significantly smaller than for the SWF.
+The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
+
+\begin{table*}[t]
+  \centering
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of the MSCKF and SWF on synthetic datasets.}
+  \begin{tabular}{*{11}{c}}
+              && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF} \B \\ \cline{3-6} \cline{8-11}
+Feature Count && Pos. ARMSE & Rot. ARMSE & ANEES   & Compute Time && Pos. ARMSE & Rot. ARMSE & ANEES & Compute Time \T\B \\ \hline
+40            && 0.2672        & 0.1378        & 10.18  &              && 0.1750        & 0.0495        & 2280 &    \T \\
+60            && 0.2550        & 0.1247        & 12.03  &              && 0.1687        & 0.0377        & 2093 &    \\
+100           && 0.2304        & 0.0952        & 16.76  &              && 0.1755        & 0.0481        & 2013 &    \B \\ \hline
+  \end{tabular}
+  \label{tab:rmse_nees}
+\end{table*}
 
@@
 Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
-In environments with more features, both algorithms provide substantial gains over pure IMU integration, although the SWF significantly outperformed the MSCKF on the datasets we tested.
+In environments with more features, both algorithms provide substantial gains over pure IMU integration, although the SWF significantly outperformed the MSCKF in terms of accuracy on the datasets we tested.
 However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.
-It is unclear from these results which algorithm would perform best in the limit of infinite features.
+It is unclear from these results which algorithm would perform best as the number of features grows very large.
 
-[Something something consistency NEES]
-
-It is worth noting that although the SWF achieved superior accuracy in most cases, the computational effort it required was significantly greater than for the MSCKF.
-The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
+Although the SWF appears to be more accurate than the MSCKF,
 

### 2014-12-30 Report/project_report.tex

@@
 Feature Count && Pos. ARMSE    & Rot. ARMSE & ANEES     & Compute Time && Pos. ARMSE   & Rot. ARMSE     & ANEES & Compute Time \T\B \\ \hline
-40            && 0.2672        & 0.1378        & 10.18  &              && 0.1750        & 0.0495        & 2280 &    \T \\
-60            && 0.2550        & 0.1247        & 12.03  &              && 0.1687        & 0.0377        & 2093 &    \\
-100           && 0.2304        & 0.0952        & 16.76  &              && 0.1755        & 0.0481        & 2013 &    \B \\ \hline
+40            && 0.2672        & 0.1378        & 10.18  & 12.19 s      && 0.1750        & 0.0495        & 2280  & 114.3 s   \T \\
+60            && 0.2550        & 0.1247        & 12.03  & 14.64 s      && 0.1687        & 0.0377        & 2093  & 175.9 s   \\
+100           && 0.2304        & 0.0952        & 16.76  & 20.58 s      && 0.1755        & 0.0481        & 2013  & 245.3 s   \B \\ \hline
   \end{tabular}
@@
 Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
-In environments with more features, both algorithms provide substantial gains over pure IMU integration, although the SWF significantly outperformed the MSCKF in terms of accuracy on the datasets we tested.
+In environments with more features, both algorithms provide substantial gains over pure IMU integration, al though the SWF significantly outperformed the MSCKF in terms of accuracy on the datasets we tested.
 However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.

### 2014-12-30 Report/project_report.tex
added abstract start

--- 2014-12-30 Report/project_report.tex
@@
 \begin{abstract}
-
-Monkey Style Chinese Kung Fu vs. SWFfer Wet Jet
-
+Accurate and consistent state estimation is an important part of autonomous vehicles. The combination of visual and inertial sensors provides
 \end{abstract}
@@
 \end{table*}
+
+\begin{table}[t]
+  \centering
+  \caption{Compute times.}
+\begin{tabular}{*{3}{c}}
+\hline
+Map Size & MSCKF Run Time [s] & SWF 25 Run Time [s]   \\ \hline
+40       & 12.19          & 114.3            \\
+60       & 14.64          & 175.9            \\
+100      & 20.58          & 245.3           \\ \hline
+\end{tabular}
+\end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-12-30 Report/project_report.tex
Transposed table and added IMU only bit

--- 2014-12-30 Report/project_report.tex
@@
 
-\begin{table*}[t]
+% \begin{table*}[t]
+%   \centering
+%     \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of the MSCKF and SWF on synthetic datasets.}
+%   \begin{tabular}{*{11}{c}}
+%               && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF} \B \\ \cline{3-6} \cline{8-11}
+% Feature Count && Pos. ARMSE    & Rot. ARMSE & ANEES     & Compute Time && Pos. ARMSE   & Rot. ARMSE     & ANEES & Compute Time \T\B \\ \hline
+% 40            && 0.2672        & 0.1378        & 10.18  & 12.19 s      && 0.1750        & 0.0495        & 2280  & 114.3 s   \T \\
+% 60            && 0.2550        & 0.1247        & 12.03  & 14.64 s      && 0.1687        & 0.0377        & 2093  & 175.9 s   \\
+% 100           && 0.2304        & 0.0952        & 16.76  & 20.58 s      && 0.1755        & 0.0481        & 2013  & 245.3 s   \B \\ \hline
+%   \end{tabular}
+%   \label{tab:rmse_nees}
+% \end{table*}
+
+\begin{table}
     \centering
-    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of the MSCKF and SWF on synthetic datasets.}
-  \begin{tabular}{*{11}{c}}
-              && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF} \B \\ \cline{3-6} \cline{8-11}
-Feature Count && Pos. ARMSE & Rot. ARMSE & ANEES   & Compute Time && Pos. ARMSE & Rot. ARMSE & ANEES & Compute Time \T\B \\ \hline
-40            && 0.2672        & 0.1378        & 10.18  &              && 0.1750        & 0.0495        & 2280 &    \T \\
-60            && 0.2550        & 0.1247        & 12.03  &              && 0.1687        & 0.0377        & 2093 &    \\
-100           && 0.2304        & 0.0952        & 16.76  &              && 0.1755        & 0.0481        & 2013 &    \B \\ \hline
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets.}
+    \begin{tabular}{*{5}c} 
+                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
+                    &                   & 40        & 60        & 100       \T\B    \\ \hline
+        IMU Only    & Position ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
+                    & Rotation ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
+                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
+                    & Compute Time      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
+        MSCKF       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
+                    & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
+                    & ANEES             & 10.18     & 12.03     & 16.76             \\
+                    & Compute Time      & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
+        SWF         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
+                    & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
+                    & ANEES             & 2280      & 2093      & 2013              \\
+                    & Compute Time      & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
     \end{tabular}
     \label{tab:rmse_nees}
-\end{table*}
+\end{table}
 
@@
 
-Although the SWF appears to be more accurate than the MSCKF,
+Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF has the advantage of being less computationally intensive than the SWF, and of having better consistency properties.
+The MSCKF may be a better choice of algorithm in situations where computational resources are limited, or where a consistent estimate of state uncertainty is desirable.
 

### 2014-12-30 Report/project_report.tex
More abstract stuff

--- 2014-12-30 Report/project_report.tex
@@
 \begin{abstract}
-
-Monkey Style Chinese Kung Fu vs. SWFfer Wet Jet
-
+Accurate and consistent egomotion estimation is an important part of autonomous vehicles.
+The combination of visual and inertial sensors provides a relatively inexpensive, compact and complementary hardware suite that can be used on many ground and aerial vehicles.
+In this work, we compare two modern approaches to egomotion estimation: the Multi-State Costraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
+Each approach is motivated and outlined in detail with.
 \end{abstract}
@@
   \begin{tabular}{*{11}{c}}
-              && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF} \B \\ \cline{3-6} \cline{8-11}
+              && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF (25)} \B \\ \cline{3-6} \cline{8-11}
 Feature Count && Pos. ARMSE    & Rot. ARMSE & ANEES     & Compute Time && Pos. ARMSE   & Rot. ARMSE     & ANEES & Compute Time \T\B \\ \hline

### 2014-12-30 Matlab/msckf/MSCKF.m

@@ -14,8 +14,8 @@ tic
 % load('../datasets/dataset3_fresh_10noisy.mat')
 % load('../datasets/dataset3_fresh_10lessnoisy.mat')
 %load('../datasets/dataset3_fresh_20lessnoisy.mat')
-load('../datasets/dataset3_fresh_40lessnoisy.mat')
-%load('../datasets/dataset3_fresh_60lessnoisy.mat')
+%load('../datasets/dataset3_fresh_40lessnoisy.mat')
+load('../datasets/dataset3_fresh_60lessnoisy.mat')
 %load('../datasets/dataset3_fresh_80lessnoisy.mat')
 %load('../datasets/dataset3_fresh_100lessnoisy.mat')
 
@@ -46,7 +46,7 @@ noiseParams.Q_imu = 0.1*diag([w_var', 1e-4*ones(1,3), v_var', 1e-4*ones(1,3)]);
 noiseParams.initialIMUCovar = 1e-4 * eye(12);
     
 %MSCKF parameters
-msckfParams.minTrackLength = 20;     % Set to inf to dead-reckon only
+msckfParams.minTrackLength = Inf;     % Set to inf to dead-reckon only
 msckfParams.maxTrackLength = 100;     % Set to inf to wait for features to go out of view
 msckfParams.maxGNCost      = 10;     % Set to inf to allow any triangulation, no matter how bad
 msckfParams.minRCOND       = 0;

### 2014-12-30 Matlab/plotting/calculateStats.m

@@ -32,6 +32,14 @@ mean(sqrt(sum(swf60.swf_rot_err.^2, 1)/3))
 mean(sqrt(sum(swf100.swf_rot_err.^2, 1)/3))
 
 %Calculate Average NEES
+stateErr = [imu.msckf_trans_err; imu.msckf_rot_err];
+stateVar = imu.err_sigma.^2;
+avgNEES = 0;
+stepNum = size(stateErr, 2);
+for i = 1:stepNum
+    avgNEES = avgNEES + 1/stepNum*stateErr(:,i)'*inv(diag(stateVar(:,i)))*stateErr(:,i);
+end
+avgNEES
 
 stateErr = [msckf40.msckf_trans_err; msckf40.msckf_rot_err];
 stateVar = msckf40.err_sigma.^2;

### 2014-12-30 Report/project_report.tex
Changed justification in table

--- 2014-12-30 Report/project_report.tex
@@
     \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets.}
-    \begin{tabular}{*{5}c} 
+    \begin{tabular}{rlccc} 
                     &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}

### 2014-12-30 Report/project_report.tex

--- 2014-12-30 Report/project_report.tex
@@
 \begin{abstract}
-
-Monkey Style Chinese Kung Fu vs. SWFfer Wet Jet
-
+Accurate and consistent egomotion estimation is an important part of autonomous vehicles.
+The combination of visual and inertial sensors provides a relatively inexpensive, compact and complementary hardware suite that can be used on many ground and aerial vehicles.
+In this work, we compare two modern approaches to egomotion estimation: the Multi-State Costraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
+Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a monocular camera and then correct this motion with image observations of salient features.
+The SWF includes the feature positions in the state estimate itself, while the MSCKF estimates the feature positions in a separate optimization routine.
+Herein, we compare the performance of the MSCKF and SWF, in error bounds and consistency, on two different datasets with real and synthetic features.
+We analyze the effect of feature track length and feature density.
+In general, the SWF is found to be more accurate than the MSCKF.
+The MSCKF, however, is signficantly faster to compute, has good consistency, and its performance improves with more tracked features.
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figs/sensorhead}
+    \caption{The sensor head used in our experiments.
+The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
+In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
+    \label{fig:sensorhead}
+\end{figure}
+
 \section{Extended Kalman Filter (EKF)} \label{sec:EKF}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figs/sensorhead}
-    \caption{The sensor head used in our experiments.
-The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
-In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
-    \label{fig:sensorhead}
-\end{figure}
-
-\begin{figure}
-    \centering
     \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}

### 2014-12-30 Report/project_report.tex
Some grammar changes

@@
 \title{\LARGE \bf
-The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and Sliding Window Filter
+The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and the Sliding Window Filter
 }
@@
 The SWF includes the feature positions in the state estimate itself, while the MSCKF estimates the feature positions in a separate optimization routine.
-Herein, we compare the performance of the MSCKF and SWF, in error bounds and consistency, on two different datasets with real and synthetic features.
+Herein, we compare the performance of the MSCKF and SWF, in error bounds and consistency, on several datasets with real and synthetic features.
 We analyze the effect of feature track length and feature density.
@@
 However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
-The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is moving slowly.
+The addition of a camera is an excellent way to bound this cumulative drift error because the camera's signal-to-noise ratio is highest when the camera is moving slowly.
 On the other hand, cameras are not robust to motion blur induced by large accelerations.
-In these cases, the strength of the IMU's signal far exceeds its baseline noise and can be relied upon more heavily in estimating pose changes.
+In these cases, IMU data can be relied upon more heavily in estimating pose changes.
 
@@
 
-In this work, we compare and contrast three modern solutions, the Extended Kalman Filter (EKF), the Sliding Window Filter (SWF), and a hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, that combines the strengths of both.
-We present an experimental comparison of the SWF and MSCKF using a dataset consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
+In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
+The MSCKF is a hybrid approach that combines the computational efficiency of an Extended Kalman Filter (EKF) with the accuracy of the SWF.
+We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
+To motivate the MSCKF, we first outline a more commmon approach of the EKF.
 
@@
 
-Another limitation of EKF-SLAM is that it is forgetful.
+Another limitation of EKF-SLAM is that it is \textit{forgetful}.
 Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.

### 2014-12-31 Report/project_report.tex
SWF math fixes

@@
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
-The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.
+The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization which acts on a sliding window of $K$ poses and $N$ features: 
+\begin{equation}
+\delta\Vec{x} := \bbm \delta \Vec x_0^T & \hdots & \delta \Vec x_K^T & \delta \Vec f_1^T & \hdots & \delta \Vec f_N^T\ebm ^T.
+\end{equation}
+The optimization applies an update, $ \delta \Vec{x}^*$, by solving the linear system:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-\delta\Vec{x} := \bbm \delta \Vec x_0 & \hdots & \delta \Vec x_K & \delta \Vec f_1 & \hdots & \delta \Vec f_N\ebm ^T,
-\end{equation}
-where $N$ is the number of unique features observed in the window.
 The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
@@
 \mathbf{H} = \left[\begin{array}{cccccc}
- -\Vec H_{\Vec x, 1} & \Vec 1 & &  & &   \\ 
- & -\Vec G_{\Vec x, 1} &  &  &  &   -\Vec G_{\Vec f, 1} \\ \hline
- & -\Vec H_{\Vec x, 2} & \Vec 1 & & &    \\ 
- & & -\Vec G_{\Vec x, 2} &  & &  -\Vec G_{\Vec f, 2}   \\ \hline
- &  & \ddots & \ddots & & \vdots  \\  \hline
-  & & & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
- & & & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
+ -\Vec H_{\Vec x, 1} & \Vec 1 & & &   \\ 
+ & -\Vec G_{\Vec x, 1} &  &   -\Vec G_{\Vec f, 1} \\ \hline
+ & \ddots & \ddots & \vdots  \\  \hline
+ & -\Vec H_{\Vec x, K} & \Vec 1 & \\ 
+ & & -\Vec G_{\Vec x, K} &  -\Vec G_{\Vec f, K}   \\  
 \end{array}\right], 
@@
 
+The update is performed repeatedly, until convergence (in this work, we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.

### 2014-12-31 Report/project_report.tex
Fixed references
More additions. Added Gabe Sibley citation

@@
 
-In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
+In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF) \cite{Sibley:2010:JFR}, the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
 The MSCKF is a hybrid approach that combines the computational efficiency of an Extended Kalman Filter (EKF) with the accuracy of the SWF.
@@
 \begin{equation}
-\delta\Vec{x} := \bbm \delta \Vec x_0^T & \hdots & \delta \Vec x_K^T & \delta \Vec f_1^T & \hdots & \delta \Vec f_N^T\ebm ^T.
+\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & \Vec f_1^T & \hdots & \Vec f_N^T\ebm ^T.
 \end{equation}
@@
 
+The weight matrix $\mbf T$ is given by:
+\begin{equation}
+\mbf T := \bbm \mbf T_1 &  &  \\ 
+ & \ddots &  \\
+ & & \mbf T_k
+\ebm 
+\end{equation}
+where
+\begin{equation}
+\mbf{T}_k := \bbm \mbf H_{\mbf w, k} \mbf Q_k \mbf H_{\mbf w, k}^T & \mbf 0 \\
+\mbf 0 & \mbf G_{\mbf n, k} \mbf R_k \mbf G_{\mbf n, k}^T \ebm,
+\end{equation}
+and $\mbf R_k$, $\mbf Q_k$ are observation and motion model noise covariances, and $\mbf G_{\mbf n, k}$, $\mbf H_{\mbf w, k}$ are the observation and motion model Jacobians with respect to the noise.
+
 The update is performed repeatedly, until convergence (in this work, we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
@@
 At time $k$, the full state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ 7-dimensional past camera poses in which active feature tracks were visible:
-\begin{align}
-    \hat{\Vec{x}}_k &:= \bbm \hat{\Vec{x}}_{I,k}^T & \hat{\Vec{q}}_{C_1 G}^T & \hat{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \hat{\Vec{q}}_{C_N G}^T & \hat{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
-\end{align}
+\begin{align*}
+    \hat{\Vec{x}}_k &:= \bbm \hat{\Vec{x}}_{I,k}^T & \hat{\Vec{q}}_{C_1 G}^T & \hat{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \hat{\Vec{q}}_{C_N G}^T & \hat{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T.
+\end{align*}
 
 We can also define the MSCKF ``error state'' at time $k$:
-\begin{align}
+\begin{align*}
     \widetilde{\Vec{x}}_k &:= \bbm \widetilde{\Vec{x}}_{I,k}^T & \boldsymbol{\delta\theta}_{C_1}^T & \widetilde{\Vec{p}}_G^{C_1 G \hspace{0.3em} T} & \hdots & \boldsymbol{\delta\theta}_{C_N}^T & \widetilde{\Vec{p}}_G^{C_N G \hspace{0.3em} T} \ebm ^T
-\end{align}
+\end{align*}
 where
@@
 However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the state.
-Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state.
+Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state (note that the SWF does not have this problem because feature positions are part of the state itself, and optimized jointly with the poses).
 Figure \ref{fig:nullspace} shows the effect of attempting to use $\Vec{r}^{(j)}$ directly in the EKF.
@@
 The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
+This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system was available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
+For both approaches, the filter window was bootstrapped with one ground truth pose, and for SWF, feature positions were intialized through triangulation.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{project_bib.bib}
+\bibliography{project_bib}
 

### 2014-12-31 Report/project_report.tex
Added smartphone and planetary citation

@@
 
-In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF) \cite{Sibley:2010:JFR}, the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
+In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
 The MSCKF is a hybrid approach that combines the computational efficiency of an Extended Kalman Filter (EKF) with the accuracy of the SWF.
+The MSCKF has recently been used successfully in real-time, accurate position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
+
 We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
@@
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
-The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window.
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure using marginalization of old poses).
 The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so the SWF may be considered inconsistent in the absence of a known map.

### 2014-12-31 Report/project_report.tex
Edited abstract

--- 2014-12-31 Report/project_report.tex
@@
 \begin{abstract}
-Accurate and consistent egomotion estimation is an important part of autonomous vehicles.
-The combination of visual and inertial sensors provides a relatively inexpensive, compact and complementary hardware suite that can be used on many ground and aerial vehicles.
-In this work, we compare two modern approaches to egomotion estimation: the Multi-State Costraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
-Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a monocular camera and then correct this motion with image observations of salient features.
-The SWF includes the feature positions in the state estimate itself, while the MSCKF estimates the feature positions in a separate optimization routine.
-Herein, we compare the performance of the MSCKF and SWF, in error bounds and consistency, on several datasets with real and synthetic features.
-We analyze the effect of feature track length and feature density.
-In general, the SWF is found to be more accurate than the MSCKF.
-The MSCKF, however, is signficantly faster to compute, has good consistency, and its performance improves with more tracked features.
+Accurate and consistent egomotion estimation is an important part of autonomous navigation.
+For this task, the combination of visual and inertial sensors is an inexpensive, compact, and complementary hardware suite that can be used on many ground and aerial vehicles.
+In this work, we compare two modern approaches to egomotion estimation: the Multi-State Constraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
+Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a vehicle and then correct this motion with observations of salient features from a monocular camera.
+While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
+We present experimental comparisons of the performance of the MSCKF and SWF on several datasets with real and synthetic features.
+In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
+In general, our results show the SWF to be more accurate than the MSCKF.
+However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
 \end{abstract}

### 2014-12-31 Matlab/swf/SlidingWindow.m
Commit

@@ -5,7 +5,7 @@ clc
 clear
 close all
 addpath('utils')
-fileName = '100lessnoisy';
+fileName = '100noisy';
 load(['../datasets/dataset3_fresh_' fileName '.mat'])
 %load('../datasets/dataset3.mat');
 tic
@@ -49,7 +49,7 @@ initialStateStruct = {};
 % Extract noise values
 Q = diag([v_var; w_var]);
 if useMonoCamera
-    R = eye(2);
+    R = 5*eye(2);
     %R = diag(y_var(1:2));
 else
     R = diag(y_var);

### 2014-12-31 Report/project_report.tex

@@
     \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets.}
-    \begin{tabular}{rlccc} 
+    \begin{threeparttable}
+    \begin{tabular}{llccc} 
                     &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
@@
                     & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
-                    & Compute Time      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
+                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
         MSCKF       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
@@
                     & ANEES             & 10.18     & 12.03     & 16.76             \\
-                    & Compute Time      & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
+                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
         SWF         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
@@
                     & ANEES             & 2280      & 2093      & 2013              \\
-                    & Compute Time      & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
+                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
     \end{tabular}
+    \begin{tablenotes}
+        \item [\dag] Running MATLAB 2014b on a MacBook Pro with Intel i7 processor and 16 GB of RAM.
+    \end{tablenotes}
+    \end{threeparttable}
     \label{tab:rmse_nees}

### 2014-12-31 Report/project_report.tex
Added explanation of MSCKF sensitivity to feature density

@@
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
-For both approaches, the filter window was bootstrapped with one ground truth pose, and for SWF, feature positions were intialized through triangulation.
+
+We implemented and tested both algorithms on a MacBook Pro with an Intel i7 processor and 16 GB of RAM running MATLAB 2014b.
+For both algorithms, we bootstrapped the filter window with one ground truth pose.
+For the SWF, we initialized feature positions using two-view triangulation.
 
@@
 With additional features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
-Note that the MSCKF's performance improves as the number of features increases while the SWF's performance is not significantly affected by increasing feature count.
+
+Note that the MSCKF's performance improves as the number of features increases, while the SWF's performance is not significantly affected by increasing feature count.
 This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
+This may be due to the fact that the MSCKF updates its state whenever a feature goes out of view and does not associate tracks corresponding to the same feature.
+If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
+As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF estimates benefit from more high-quality motion constraints.
+
+Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
+However, [EXPLANATION ABOUT WHY IT DOESN'T IMPROVE MUCH WITH MORE FEATURES]
 

### 2014-12-31 Matlab/simulation/augmentDataset.m

@@ -12,12 +12,12 @@ calibParams.f_v = fv;
 calibParams.b = b;
 
 %Noise parameters
-y_var = [1,1,1,1];
+% y_var = [1,1,1,1];
 mu = zeros(4,1);
 sigma = diag(y_var);
 
 %Generate new landmarks
-newLmNum = 60;
+newLmNum = 100;
 
 if exist('newLmPos', 'var')
     oldLmNum = size(newLmPos,2);
@@ -67,4 +67,4 @@ for k = 1:length(t)
 end
 
 
-save(['../datasets/dataset3_fresh_',num2str(newLmNum),'lessnoisy.mat']);
+save(['../datasets/dataset3_fresh_',num2str(newLmNum),'noisy.mat']);

### 2014-12-31 Matlab/swf/SlidingWindow.m
Added some more details about the laptop

@@ -5,7 +5,7 @@ clc
 clear
 close all
 addpath('utils')
-fileName = '100noisy';
+fileName = '60noisy';
 load(['../datasets/dataset3_fresh_' fileName '.mat'])
 %load('../datasets/dataset3.mat');
 tic

### 2014-12-31 Report/project_report.tex
Added some more details about the laptop
Rearranged SWF jacobians for clarity

@@
 \begin{align}
+\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
+\\[0.5em]
 \Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
@@
 \\[0.5em]
-\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
-\\[0.5em]
+\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}} \\[0.5em]
 \Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
-\\[0.5em]
-\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
 \end{align}

### 2014-12-31 Report/project_report.tex
Edits

--- 2014-12-31 Report/project_report.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{INTRODUCTION} \label{sec:introduction}
+\section{Introduction} \label{sec:introduction}
 
@@
 Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
-Since an IMU directly measures accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
+Since an IMU directly measures linear accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
 However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
 The addition of a camera is an excellent way to bound this cumulative drift error because the camera's signal-to-noise ratio is highest when the camera is moving slowly.
-On the other hand, cameras are not robust to motion blur induced by large accelerations.
+On the other hand, cameras are not robust to motion blur induced by rapid motions.
 In these cases, IMU data can be relied upon more heavily in estimating pose changes.
@@
 
-In this work, we compare and contrast two modern solutions the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
-The MSCKF is a hybrid approach that combines the computational efficiency of an Extended Kalman Filter (EKF) with the accuracy of the SWF.
-The MSCKF has recently been used successfully in real-time, accurate position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
-
+In this work, we compare and contrast two modern solutions: the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
+The MSCKF is a hybrid approach that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
+The MSCKF has recently been used successfully in accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
 We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
-To motivate the MSCKF, we first outline a more commmon approach of the EKF.
+In order to provide context for the MSCKF and SWF, we first outline the EKF approach to the SLAM problem.
 
@@
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
-The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization which acts on a sliding window of $K$ poses and $N$ features: 
+The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses observing $M$ features: 
 \begin{equation}
-\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & \Vec f_1^T & \hdots & \Vec f_N^T\ebm ^T.
+\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & \Vec f_1^T & \hdots & \Vec f_M^T\ebm ^T.
 \end{equation}
@@
 \\[0.5em]
-\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
-\\[0.5em]
+\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j\in[1,M]$}} \\[0.5em]
 \Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
-\\[0.5em]
-\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j$}}
 \end{align}
@@
 \begin{equation}
-\mbf T := \bbm \mbf T_1 &  &  \\ 
- & \ddots &  \\
- & & \mbf T_k
-\ebm 
+\Vec{T} := \text{diag} \left\{ \Vec{T}_1,\hdots,\Vec{T}_K \right\}
 \end{equation}
@@
 
-The update is performed repeatedly, until convergence (in this work, we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
+The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:MSCKF}
+\section{Multi-State Constraint Kalman Filter (MSCKF)} \label{sec:MSCKF}
 The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
-The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
-This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.
+The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
+This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds some preset threshold.
 
@@
 \begin{align}
-    \boldsymbol{\delta}\Vec{q} &:= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T.
+    \boldsymbol{\delta}\Vec{q} &:= \hat{\Vec{q}}^{-1} \otimes \Vec{q} \simeq \bbm \frac{1}{2}\boldsymbol{\delta\theta}^T & 1 \ebm^T
 \end{align}
+where $\otimes$ denotes quaternion multiplication.
 
@@
 \end{align}
-where $\hat{\Vec{C}}_{IG,k}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG,k}$ and $\otimes$ denotes quaternion multiplication.
+where $\hat{\Vec{C}}_{IG,k}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG,k}$.
 
-Assuming the MSCKF state has already been augmented by $N$ camera poses, we add the $\left(N+1\right)^\text{th}$ camera pose to the state as follows:
+Assuming the MSCKF state has already been augmented by $N$ camera poses, we add the $\left(N+1\right)^\text{th}$ camera pose to the state according to
 \begin{align}
@@
 \end{align}
-
-We must also augment the MSCKF state covariance:
+and augment the MSCKF state covariance according to
 \begin{align}
@@
     \dot{\hat{\Vec{P}}}_{II}\left(t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) 
-                \hat{x}_{C_i}^{C_1 C_i} + \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) \Vec{F}^T \notag \\ 
+                 + \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) \Vec{F}^T \notag \\ 
                 & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
@@
 \begin{align}
-\hat{\Vec{z}}_i^{(j)} &= \bbm u'_i & v'_i \ebm ^T = \bbm (u_i - c_u)/f_u & (v_i - c_v)/f_v \ebm ^T
+\hat{\Vec{z}}_i^{(j)} &= \bbm u'_i & v'_i \ebm ^T = \bbm (u_i - c_u)/f_u & (v_i - c_v)/f_v \ebm ^T.
 \end{align}
-We initialize the optimization by estimating the position of feature $f_j$ in camera frame $C_1$ using a linear least squares method with measurements from the first two camera frames, $C_1$ and $C_2$:
+We initialize the optimization by estimating the position of feature $f_j$ in camera frame $C_1$ using a linear least-squares method with measurements from the first two camera frames, $C_1$ and $C_2$:
 \begin{align}
@@
 \end{align}
-with
+where
 \begin{align}
@@
 However, the EKF assumes that measurement errors are linear in the error in the state and have an additive zero-mean Gaussian noise component that is \textit{uncorrelated} to the state.
-Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state (note that the SWF does not have this problem because feature positions are part of the state itself, and optimized jointly with the poses).
+Since the term $\Vec{H}_{f}^{(j)} \widetilde{\Vec{p}}_G^{f_j G}$ \textit{is} correlated to the state, $\Vec{r}^{(j)}$ is not of the correct form for the EKF and must be modified to de-correlate it from the state.
 Figure \ref{fig:nullspace} shows the effect of attempting to use $\Vec{r}^{(j)}$ directly in the EKF.
 The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
+As an aside, note that the SWF does not have this problem because feature positions are part of the state itself and are jointly optimized with the poses.
 
@@
     \caption{Effect of correlation between measurement noise and filter state on MSCKF performance.
-The correlation between the measurement error and the state causes the filter estimates to drift further from the true values than with a modified, de-corellated measurement error.}
+The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement noise.}
     \label{fig:nullspace}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{EXPERIMENTS} \label{sec:experiments}
+\section{Experiments} \label{sec:experiments}
 We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
 The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system was available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
+This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{CONCLUSIONS} \label{sec:conclusions}
+\section{Conclusions} \label{sec:conclusions}
 Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
-In environments with more features, both algorithms provide substantial gains over pure IMU integration, although the SWF significantly outperformed the MSCKF in terms of accuracy on the datasets we tested.
+In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, although the SWF significantly outperformed the MSCKF on the datasets we tested.
 However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.
@@
 Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF has the advantage of being less computationally intensive than the SWF, and of having better consistency properties.
-The MSCKF may be a better choice of algorithm in situations where computational resources are limited, or where a consistent estimate of state uncertainty is desirable.
+The MSCKF may be a better choice of algorithm in applications where computational resources are limited, or where a consistent estimate of state uncertainty is desirable.
 

### 2014-12-31 Report/project_report.tex
More conclusionery

@@
 As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF estimates benefit from more high-quality motion constraints.
-
 Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
-However, [EXPLANATION ABOUT WHY IT DOESN'T IMPROVE MUCH WITH MORE FEATURES]
 
@@
 Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
+We stress that our inertial data was obtained from a high-quality IMU and has been sanitized to account for gravity, biases, and integration error.
+Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and the benefit of the SWF and MSCKF would have been more apparent.
+
 In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, although the SWF significantly outperformed the MSCKF on the datasets we tested.

### 2014-12-31 Report/project_report.tex
Some small changes

@@
 In this work, we compare two modern approaches to egomotion estimation: the Multi-State Constraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
-Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a vehicle and then correct this motion with observations of salient features from a monocular camera.
+Both filters use an Inertial Measurement Unit (IMU) to estimate the motion of a vehicle and then correct this estimate with observations of salient features from a monocular camera.
 While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
@@
 The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
-A complicating factor in the general form of this problem is the absence of a known map of features from which the camera can generate measurements.
+This problem is often complicated by the absence of a known map of features from which the camera can generate measurements.
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
@@
     \begin{tablenotes}
-        \item [\dag] Running MATLAB 2014b on a MacBook Pro with Intel i7 processor and 16 GB of RAM.
+        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
     \end{tablenotes}

### 2014-12-31 Report/project_report.tex
Some more small features

@@
 Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
-This behavior is due to the fact that the dimension of the state grows linearly with the number of features, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
+This behavior is due to the fact that the dimension of the state grows linearly with the number of features, and the computational cost of inverting the state covariance matrix (during the Kalman gain computation) is cubic in the dimension of the state.
 Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
@@
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
-The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure using marginalization of old poses).
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure of using marginalization of old poses which was not implemented in this work).
 The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so the SWF may be considered inconsistent in the absence of a known map.

### 2014-12-31 Report/project_report.tex
Minor edits

--- 2014-12-31 Report/project_report.tex
@@
 While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
-We present experimental comparisons of the performance of the MSCKF and SWF on several datasets with real and synthetic features.
+We present experimental comparisons of the MSCKF and SWF on several datasets with real and synthetic features.
 In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
@@
 On the other hand, cameras are not robust to motion blur induced by rapid motions.
-In these cases, IMU data can be relied upon more heavily in estimating pose changes.
+In these cases, IMU data can be relied upon more heavily in estimating egomotion.
 
@@
 The MSCKF is a hybrid approach that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
-The MSCKF has recently been used successfully in accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
+The MSCKF has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
 We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.

### 2014-12-31 Report/project_report.tex
Fixed nullspace plot legend, and other things

@@
 We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
-In order to accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.
+In order to accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.
 
@@
 \begin{align}
-    \hat{\Vec{x}}_k &\leftarrow \bbm \hat{\Vec{x}}_k^T & \hat{\Vec{q}}_{C_{N+1} G}^T & \hat{\Vec{p}}_G^{C_{N+1} G \hspace{0.3em} T} \ebm^T .
+    \hat{\Vec{x}}_k &\leftarrow \bbm \hat{\Vec{x}}_k^T & \hat{\Vec{q}}_{C_{N+1} G}^T & \hat{\Vec{p}}_G^{C_{N+1} G \hspace{0.3em} T} \ebm^T 
 \end{align}
@@
 \subsection{Feature Position Estimation}
-When a feature $f_j$ goes out of view of the camera, the MSCKF estimates its position $\hat{\Vec{p}}_G^{f_j G}$ using an inverse-depth least-squares Gauss-Newton optimization. 
+When a tracked feature $f_j$ is selected for a state update, the MSCKF estimates its position $\hat{\Vec{p}}_G^{f_j G}$ using an inverse-depth least-squares Gauss-Newton optimization. 
 The procedure takes as input $N$ camera poses and $N$ sets of ``ideal'' pixel measurements, where ``ideal'' means that the pixel measurements have been corrected for the camera intrinsics:
@@
 
-Next, we define a vector of three parameters:
+We can express the feature position in camera frame $C_i$ in terms of its position in camera frame $C_1$ as
 \begin{align}
-\hat{\Vec{y}} &= \bbm \alpha & \beta & \gamma \ebm^T = \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T
+\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i} .
 \end{align}
-We can express the feature position in camera frame $C_i$ in terms of its position in camera frame $C_1$:
+By forming the vector
 \begin{align}
-\hat{\Vec{p}}_{C_i}^{f_j C_i} = \hat{\Vec{C}}_{i1} \hat{\Vec{p}}_{C_1}^{f_j C_1}  + \hat{\Vec{p}}_{C_i}^{C_1 C_i}
+\hat{\Vec{y}} &:= \bbm \alpha & \beta & \gamma \ebm^T := \frac{1}{\hat{Z}^{(j)}_{C_1}} \bbm \hat{X}^{(j)}_{C_1} & \hat{Y}^{(j)}_{C_1} & 1 \ebm ^T ,
 \end{align}
-With our state parameters, we can use this equation to define three functions:
+we can define three functions:
 \begin{align}
@@
             C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
-        \ebm 
+        \ebm .
 \end{align}
@@
 \end{align}
-The least squares linear system then becomes
+and the linear least-squares system then becomes
 \begin{align}
@@
 with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
-
 The Jacobian elements are given by
@@
 \subsection{MSCKF State Correction Equations}
-Now that we have estimated the positions of any features that have gone out of view, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
+Now that we have estimated the positions of any features to be used in the state update, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
 We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from the $i^\text{th}$ camera pose $C_i$ in the window:
@@
 The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
-As an aside, note that the SWF does not have this problem because feature positions are part of the state itself and are jointly optimized with the poses.
+As an aside, note that the SWF does not have this problem because feature positions are part of the state itself and are jointly optimized along with the poses.
 
@@
 Figure \ref{fig:comp_3} compares translational and rotational Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
-With additional features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
+With more features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
 
@@
 If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
-As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF estimates benefit from more high-quality motion constraints.
+As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF benefits from more high-quality motion constraints.
 Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
@@
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
-The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure of using marginalization of old poses which was not implemented in this work).
-The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so the SWF may be considered inconsistent in the absence of a known map.
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure of using marginalization of old poses, which was not implemented in this work).
+The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.
 
-Another point in favour of the MSCKF is that the computational effort required was significantly smaller than for the SWF.
+Another point in favour of the MSCKF is that the computational effort required was an order of magnitude smaller than for the SWF.
 The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
@@
 We stress that our inertial data was obtained from a high-quality IMU and has been sanitized to account for gravity, biases, and integration error.
-Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and the benefit of the SWF and MSCKF would have been more apparent.
+Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent.
 
-In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, although the SWF significantly outperformed the MSCKF on the datasets we tested.
+In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, with the SWF significantly outperforming the MSCKF on the datasets we tested.
 However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.

### 2014-12-31 Matlab/plotting/nullspace_plots.m

@@ -4,12 +4,12 @@ nst_off = load('msckf_NST_off.mat');
 
 plotAbsErr = true;
 
-msckf_trans_err = nst_on.msckf_trans_err;
-msckf_rot_err = nst_on.msckf_rot_err;
-msckf_err_sigma = nst_on.err_sigma;
-swf_trans_err = nst_off.msckf_trans_err;
-swf_rot_err = nst_off.msckf_rot_err;
-swf_err_sigma = nst_off.err_sigma;
+msckf_trans_err = nst_off.msckf_trans_err;
+msckf_rot_err = nst_off.msckf_rot_err;
+msckf_err_sigma = nst_off.err_sigma;
+swf_trans_err = nst_on.msckf_trans_err;
+swf_rot_err = nst_on.msckf_rot_err;
+swf_err_sigma = nst_on.err_sigma;
 tPlot = nst_on.tPlot;
 
 transLim = [-0.4 0.4];
@@ -28,12 +28,12 @@ end
 figure
 subplot(3,1,1)
 hold on
-plot(tPlot, msckf_trans_err(1,:), '-b', 'LineWidth', 1.2)
-% plot(tPlot, -3*msckf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
-% plot(tPlot, 3*msckf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
-plot(tPlot, swf_trans_err(1,:), '-r', 'LineWidth', 1.2)
-% plot(tPlot, -3*swf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
-% plot(tPlot, 3*swf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
+plot(tPlot, msckf_trans_err(1,:), '-r', 'LineWidth', 1.2)
+% plot(tPlot, -3*msckf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
+% plot(tPlot, 3*msckf_err_sigma(4,:), '--r', 'LineWidth', 1.2)
+plot(tPlot, swf_trans_err(1,:), '-b', 'LineWidth', 1.2)
+% plot(tPlot, -3*swf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
+% plot(tPlot, 3*swf_err_sigma(4,:), '--b', 'LineWidth', 1.2)
 
[…]

### 2014-12-31 Report/project_report.tex
SWF Math
Goodbye mbf, hello Vec!

@@
 \begin{equation}
-\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & \Vec f_1^T & \hdots & \Vec f_M^T\ebm ^T.
+\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & {\Vec {p}_G^{f_1G}}^T & \hdots & {\Vec {p}_G^{f_MG}}^T \ebm ^T.
 \end{equation}
@@
 \Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
-    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\boldsymbol{\rho}_i^{f_j G} - \Vec r_G^{IG,k} ) )^\times  \ebm ^T
+    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec p_G^{IG,k} ) )^\times  \ebm ^T
 \\[0.5em]
@@
 \begin{equation}
-\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\boldsymbol{\rho}_G^{f_j G} - \Vec r_G^{IG,k} )   - \boldsymbol \rho_I^{CI} \right),
+\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec{p}_G^{IG,k} )   - \Vec p_I^{CI} \right),
 \end{equation}
@@
 \begin{equation}
-\mbf{T}_k := \bbm \mbf H_{\mbf w, k} \mbf Q_k \mbf H_{\mbf w, k}^T & \mbf 0 \\
-\mbf 0 & \mbf G_{\mbf n, k} \mbf R_k \mbf G_{\mbf n, k}^T \ebm,
+\Vec{T}_k := \bbm \Vec H_{\mbf w, k} \Vec Q_k \Vec H_{\Vec w, k}^T & \Vec 0 \\
+\Vec 0 & \Vec G_{\Vec n, k} \Vec R_k \Vec G_{\Vec n, k}^T \ebm,
 \end{equation}
-and $\mbf R_k$, $\mbf Q_k$ are observation and motion model noise covariances, and $\mbf G_{\mbf n, k}$, $\mbf H_{\mbf w, k}$ are the observation and motion model Jacobians with respect to the noise.
+and $\mbf R_k$, $\mbf Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
 

### 2014-12-31 Report/project_report.tex
mbf to Vec

@@
 \begin{equation}
-\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & {\Vec {p}_G^{f_1G}}^T & \hdots & {\Vec {p}_G^{f_MG}}^T \ebm ^T.
+\Vec{x} := \bbm \Vec x_0^T & \hdots & \Vec x_K^T & {\Vec {p}_G^{f_1G \hspace{0.3em} T}} & \hdots & {\Vec {p}_G^{f_MG \hspace{0.3em} T}} \ebm ^T.
 \end{equation}
@@
 \Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
-    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec p_G^{IG,k} ) )^\times  \ebm ^T
+    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec p_{G,k}^{IG} ) )^\times  \ebm ^T
 \\[0.5em]
@@
 \begin{equation}
-\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec{p}_G^{IG,k} )   - \Vec p_I^{CI} \right),
+\bar{\Vec p}_{C_k}^{f_j C_k} := \bbm x \\ y \\ z \ebm = \Vec C_{CI} \left( \Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec{p}_{G,k}^{IG} )   - \Vec p_I^{CI} \right),
 \end{equation}
@@
 
-The weight matrix $\mbf T$ is given by:
+The weight matrix $\Vec T$ is given by:
 \begin{equation}
@@
 \begin{equation}
-\Vec{T}_k := \bbm \Vec H_{\mbf w, k} \Vec Q_k \Vec H_{\Vec w, k}^T & \Vec 0 \\
+\Vec{T}_k := \bbm \Vec H_{\Vec w, k} \Vec Q_k \Vec H_{\Vec w, k}^T & \Vec 0 \\
 \Vec 0 & \Vec G_{\Vec n, k} \Vec R_k \Vec G_{\Vec n, k}^T \ebm,
 \end{equation}
-and $\mbf R_k$, $\mbf Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
+and $\Vec R_k$, $\Vec Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
 

### 2014-12-31 Report/project_report.tex
Minor fixes

@@
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
-The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes the proper procedure of using marginalization of old poses, which was not implemented in this work).
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
 The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.
@@
     \centering
-    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets.}
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $k\in[1215,1715]$.}
     \begin{threeparttable}
@@
                     & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
-        SWF         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
+        SWF (25 Poses)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
                     & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\

### 2014-12-31 Report/project_report.tex
Added comment about synthetic data noise
Ran out of brain-compute

@@
 We constructed each dataset so that the larger maps contained the same features as the smaller maps, plus additional features to make up the difference.
-Although the camera measurements in these datasets are synthetic, the IMU data are unchanged from the original dataset.
+In each dataset, the IMU data remained unchanged from the original dataset and we corrupted the synthetic camera measurements with zero-mean Gaussian noise.
 Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
@@
 Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF has the advantage of being less computationally intensive than the SWF, and of having better consistency properties.
-The MSCKF may be a better choice of algorithm in applications where computational resources are limited, or where a consistent estimate of state uncertainty is desirable.
+The MSCKF may be a better choice of algorithm in applications where computational resources are limited and the operational environment is feature-rich.
 

### 2014-12-31 Report/project_report.tex
Added note about notation and other little edits

@@
 
-In this work, we compare and contrast two modern solutions: the Sliding Window Filter (SWF), the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
-The MSCKF is a hybrid approach that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
+In this work, we compare and contrast two modern solutions: the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
+The MSCKF is a hybrid estimator that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
 The MSCKF has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
@@
 In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
-This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of features by recursively updating the state as new measurements become available.
+This technique, sometimes referred to as EKF-SLAM, attempts to track vehicle motion and create a globally consistent map of features by recursively updating the state as new measurements become available.
 %Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
@@
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
-The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses observing $M$ features: 
+The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses $\Vec x_{k\in[1,K]}$ observing $M$ features $f_{j\in[1,M]}$:
 \begin{equation}
@@
 \end{equation}
+In our notation, $\Vec p^{BA}_A$ is the vector from the origin of frame $\mathcal{F}_A$ to the origin of frame $\mathcal{F}_B$ expressed in $\mathcal{F}_A$, and $\Vec C_{BA}$ is the rotation matrix from $\mathcal{F}_A$ to $\mathcal{F}_B$.
+
 The optimization applies an update, $ \delta \Vec{x}^*$, by solving the linear system:
@@
 \end{equation}
-
 The weight matrix $\Vec T$ is given by:
@@
 and $\Vec R_k$, $\Vec Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
+The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
 
-The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
@@
 \subsection{MSCKF State Covariance Propagation}
-With reference to the partitions defined in (\ref{eq:covariance}), we compute the predicted camera-camera and IMU-camera state covariances as follows:
+With reference to the partitions defined in (\ref{eq:covariance}), we compute the predicted camera state covariances and IMU-camera cross-correlations as follows:
 \begin{align}
@@
 
-We obtain the predicted IMU-IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
+We obtain the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
 \begin{align} \label{eq:imucovar}
@@
 \end{align}
-and
+is the direction of the ray emanating from camera $C_i$ along which feature $f_j$ must lie, and
 \begin{align}
@@
 \end{align}
-we can define three functions:
+we can define three functions
 \begin{align}
@@
             C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
-        \ebm .
+        \ebm
 \end{align}
-The camera measurement error can then be written as
+and rewrite the camera measurement error as
 \begin{align}
-\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm
+\Vec e(\hat{\Vec{y}}) = \hat{\Vec{z}}_i^{(j)} - \frac{1}{h_3 (\hat{\Vec{y}}) } \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \ebm .
 \end{align}
-and the linear least-squares system then becomes
+The linear least-squares system then becomes
 \begin{align}
@@
 Now that we have estimated the positions of any features to be used in the state update, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
-We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from the $i^\text{th}$ camera pose $C_i$ in the window:
+We begin by forming the exteroceptive measurement error corresponding to an observation $\Vec{z}_i^{(j)}$ of feature $f_j$ from camera pose $C_i$:
 \begin{align} \label{eq:residual_def}
@@
 \end{align}
-where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $i$ with respect to the filter state and the position of the feature, respectively.
+where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $C_i$ with respect to the filter state and the position of the feature, respectively.
 These are given by
@@
 
-We implemented and tested both algorithms on a MacBook Pro with an Intel i7 processor and 16 GB of RAM running MATLAB 2014b.
+We implemented and tested both algorithms in MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
 For both algorithms, we bootstrapped the filter window with one ground truth pose.

### 2014-12-31 Report/project_report.tex
Made figures with subfigures span both columns

@@
 
-\begin{figure}
+\begin{figure*}
     \centering
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/NST_trans}
+        \includegraphics[width=0.45\textwidth]{figs/NST_trans}
         \label{fig:nullspace:trans_err}
     }
-
+    ~
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/NST_rot}
+        \includegraphics[width=0.45\textwidth]{figs/NST_rot}
         \label{fig:nullspace:rot_err}
@@
     \label{fig:nullspace}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth}
+    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth}
     \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figs/nVisible_500_1000}
+    \includegraphics[width=0.45\textwidth]{figs/nVisible_500_1000}
     \caption{Number of visible features in the interval $k\in[500,1000]$.
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
+        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
         \label{fig:comp_1:trans_err}
     }
-
+    ~
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
+        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
         \label{fig:comp_1:rot_err}
@@
     \label{fig:comp_1}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figs/nVisible_1215_1715}
+    \includegraphics[width=0.45\textwidth]{figs/nVisible_1215_1715}
     \caption{Number of visible features in the interval $k\in[1215,1715]$.
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
+        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
         \label{fig:comp_2:trans_err}
     }
-
+    ~
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
+        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
         \label{fig:comp_2:rot_err}
@@
     \label{fig:comp_2}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
+    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
     \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
         \label{fig:comp_3:trans_err}
     }
-
+    ~
     \subfigure {
-        \includegraphics[width=0.48\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
         \label{fig:comp_3:rot_err}
@@
     \label{fig:comp_3}
-\end{figure}
+\end{figure*}
 

### 2014-12-31 Report/project_report.tex
vspace to the rescue!
Renumbered figs
Fixed numbering
Final touches

@@
 \end{equation}
-The weight matrix $\Vec T$ is given by:
+The weight matrix $\Vec T$ is given by
 \begin{equation}
@@
 
+\setcounter{figure}{2}
 \begin{figure*}
@@
 
+\setcounter{figure}{1}
 \begin{figure}
@@
 
+\setcounter{figure}{3}
 \begin{figure}
@@
 
+\setcounter{figure}{5}
 \begin{figure*}
@@
 
+\setcounter{figure}{4}
 \begin{figure}
@@
 
+\setcounter{figure}{6}
 \begin{figure*}
@@
 
+\setcounter{figure}{8}
 \begin{figure}
@@
 Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
-
+\setcounter{figure}{7}
 \begin{figure*}
@@
 
-Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the synthetic datasets.
-As expected from Figure \ref{fig:comp_3}, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
+Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
+As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
@@
                     & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
-        MSCKF       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
+        MSCKF (20-100)       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
                     & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
@@
                     & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
-        SWF (25 Poses)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
+        SWF (25)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
                     & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
@@
     \label{tab:rmse_nees}
+    \vspace{-0.4cm}
 \end{table}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-0.1cm}
 \section{Conclusions} \label{sec:conclusions}

### 2015-02-28 Matlab/msckf/MSCKF.m
Lots of extra datasets and various modifications that I apparently forgot to commit before?

@@ -9,23 +9,40 @@ clear;
 close all;
 clc;
 addpath('utils');
-addpath('stereoCamProject.m');
 
 tic
 % load('../datasets/dataset3.mat')
 % load('../datasets/dataset3_fresh_10noisy.mat')
 % load('../datasets/dataset3_fresh_10lessnoisy.mat')
-%load('../datasets/dataset3_fresh_20lessnoisy.mat')
-%load('../datasets/dataset3_fresh_40lessnoisy.mat')
+% load('../datasets/dataset3_fresh_20lessnoisy.mat')
+% load('../datasets/dataset3_fresh_40lessnoisy.mat')
 % load('../datasets/dataset3_fresh_60lessnoisy.mat')
-%load('../datasets/dataset3_fresh_80lessnoisy.mat')
+% load('../datasets/dataset3_fresh_80lessnoisy.mat')
 % load('../datasets/dataset3_fresh_100lessnoisy.mat')
+% load('../datasets/dataset3_fresh2_500lessnoisy.mat')
 % load('../datasets/2015-01-28-16-19-56_2loops.mat')
 % load('../datasets/2015-01-30-16-08-22_overtable.mat')
-load('../datasets/2015-01-30-16-35-49.mat')
+% load('../datasets/2015-01-30-16-35-49.mat')
+% load('../datasets/2015-02-04-12-54-12_chairwall2.mat')
+% load('../datasets/2015-02-04-13-40-35_chairwallstudentroom2.mat')
+% load('../datasets/2015-02-04-12-53-41_chairwall.mat')
+% load('../datasets/2015-01-28-16-11-17_1loop_5hz.mat');
+% load('../datasets/2015-01-28-16-11-17_1loop_KLT.mat');
+load('../datasets/2015-01-28-16-19-56_2loops_KLT.mat');
+% load('../datasets/2015-02-04-16-24-21_checkerboardsquare.mat');
+% load('../datasets/2015-02-04-16-24-21_checkerboardsquare_5hz_checkerCorners.mat')
+% load('../datasets/2015-02-04-16-24-21_checkerboardsquare_planarMeas.mat')
+% load('../datasets/2011_09_26_drive_0005_sync.mat')
+% load('../datasets/2011_09_26_drive_0009_sync.mat')
+% load('../datasets/2011_09_26_drive_0095_sync.mat')
+% load('../datasets/2011_09_26_drive_0005_sync_KLT.mat')
+% load('../datasets/2011_09_26_drive_0009_sync_KLT.mat')
[…]

### 2015-02-28 Matlab/msckf/plot_traj.m

@@ -1,16 +1,31 @@
 kNum = length(prunedStates);
-p_C_G_est = zeros(3,kNum);
-q_CG_est = zeros(4,kNum);
 
-for k=1:kNum
-    p_C_G_est(:,k) = prunedStates{k}.p_C_G;
-    q_CG_est(:,k)  = prunedStates{k}.q_CG;
-end
+if kNum > 0
+    p_I_G_est = zeros(3,kNum);
+    p_I_G_imu = zeros(3, kNum);
+    kPlot = zeros(1,kNum);
 
-figure(1); clf; hold on;
-plot3(p_C_G_est(1,:),p_C_G_est(2,:),p_C_G_est(3,:),'-b');
-if ~isempty(map)
-    scatter3(map(1,:),map(2,:),map(3,:),'or');
+    for k=1:kNum
+        state_k = prunedStates{k}.state_k;
+
+        C_CG = quatToRotMat(prunedStates{k}.q_CG);
+        C_CI = quatToRotMat(camera.q_CI);
+        p_I_G_est(:,k) = prunedStates{k}.p_C_G - C_CG' * C_CI * camera.p_C_I;
+        p_I_G_imu(:,k) = msckfState_imuOnly{state_k}.imuState.p_I_G;
+
+        kPlot(k) = state_k;
+    end
+
+    figure(1); clf; hold on;
+    plot3(p_I_G_est(1,:),p_I_G_est(2,:),p_I_G_est(3,:),'-b');
+    plot3(p_I_G_imu(1,:),p_I_G_imu(2,:),p_I_G_imu(3,:),'-r');
+%     plot3(r_i_vk_i(1,kPlot),r_i_vk_i(2,kPlot),r_i_vk_i(3,kPlot),'-g');
+    % if ~isempty(map)
+    %     scatter3(map(1,:),map(2,:),map(3,:),'or');
+    % end
+    xlabel('x');ylabel('y');zlabel('z');
+%     legend('MSCKF','IMU integration','Ground Truth');
+    legend('MSCKF','IMU integration');
+    grid on;
+     drawnow;
 end
-xlabel('x');ylabel('y');zlabel('z');
-pause(0.001); drawnow;

### 2015-02-28 Matlab/simulation/augmentDataset.m

@@ -12,12 +12,12 @@ calibParams.f_v = fv;
 calibParams.b = b;
 
 %Noise parameters
-% y_var = [1,1,1,1];
+y_var = [1,1,1,1];
 mu = zeros(4,1);
 sigma = diag(y_var);
 
 %Generate new landmarks
-newLmNum = 100;
+newLmNum = 500;
 
 if exist('newLmPos', 'var')
     oldLmNum = size(newLmPos,2);
@@ -34,7 +34,7 @@ end
 
 xRange = [min(rho_i_pj_i(1,:)) - 5, max(rho_i_pj_i(1,:)) + 5];
 yRange = [min(rho_i_pj_i(2,:)) - 5, max(rho_i_pj_i(2,:)) + 5];
-zRange = [min(rho_i_pj_i(3,:)) - 5, max(rho_i_pj_i(3,:)) ];
+zRange = [min(rho_i_pj_i(3,:)) - 5, max(rho_i_pj_i(3,:)) + 5];
 
 newLmPosX = range(xRange)*rand(1,newLmNum - oldLmNum) + xRange(1);
 newLmPosY = range(yRange)*rand(1,newLmNum - oldLmNum) + yRange(1);
@@ -57,7 +57,9 @@ for k = 1:length(t)
         p_li_i = newLmPos(:,lm_i);
         p_lc_c = homo2cart(T_ci*cart2homo(p_li_i));
         [yMeas] = stereoCamProject(p_lc_c, calibParams);
-        if all(yMeas > 0) && all(yMeas([1,3]) <= 640) && all(yMeas([2,4]) <= 480)
+        if all(yMeas > 0) ...
+            && all(yMeas([1,3]) <= 640) && all(yMeas([2,4]) <= 480)...
+            && p_lc_c(3) > 0    % in view and in front of camera
                 noise = mvnrnd(mu, sigma)';
                 y_k_j(:,k, lm_i) = yMeas + noise;
         else
@@ -67,4 +69,4 @@ for k = 1:length(t)
 end
 
 
-save(['../datasets/dataset3_fresh_',num2str(newLmNum),'noisy.mat']);
+save(['../datasets/dataset3_fresh2_',num2str(newLmNum),'lessnoisy.mat']);

### 2015-02-28 Report/project_report.tex
Added JL as author

@@
 
-\author{Lee Clement$^{1}$ and Valentin Peretroukhin$^{1}$% <-this % stops a space
+\author{Lee Clement, Valentin Peretroukhin, and Jacob Lambert% <-this % stops a space
 %\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Institute for Aerospace Studies,
-        University of Toronto, Toronto, ON, Canada
-        {\tt \{lee.clement, valentin.peretroukhin\} @robotics.utias.utoronto.ca}
+\thanks{All authors are at the Institute for Aerospace Studies,
+        University of Toronto, Canada
+        {\tt \{lee.clement, valentin.peretroukhin, jacob.lambert\} @robotics.utias.utoronto.ca}
         }%

### 2015-02-28 Matlab/kitti_extraction/sanitizeKITTI_KLT.m
Added KITTI extraction

@@ -1,0 +1,264 @@
+%% Clean up and import
+clc;
+close all;
+addpath('utils');
+addpath('utils/devkit');
+
+
+dataBaseDir = '~/Desktop/KITTI/2011_09_26/2011_09_26_drive_0035_sync';
+dataCalibDir = '~/Desktop/KITTI/2011_09_26';
+
+%% Get ground truth and import data
+frameRange = 1:130;
+%Image data
+leftImageData = loadImageData([dataBaseDir '/image_00'], frameRange);
+rightImageData = loadImageData([dataBaseDir '/image_01'], frameRange);
+%IMU data
+[imuData, imuFrames] = loadImuData(dataBaseDir, leftImageData.timestamps);
+%Ground Truth
+T_wIMU_GT = getGroundTruth(dataBaseDir, imuFrames);
+
+skipFrames = 1;
+minObsNum = 3;
+
+%% Load calibration
+[T_camvelo_struct, P_rect_cam1] = loadCalibration(dataCalibDir);
+T_camvelo = T_camvelo_struct{1}; 
+T_veloimu = loadCalibrationRigid(fullfile(dataCalibDir,'calib_imu_to_velo.txt'));
+T_camimu = T_camvelo*T_veloimu;
+
+%Add camera ground truth
+
+T_wCam_GT = T_wIMU_GT;
+
+for i = 1:size(T_wIMU_GT, 3)
+    T_wCam_GT(:,:,i) = T_wIMU_GT(:,:,i)*inv(T_camimu);
+end
+
+%% Process images
+
+meanErrorHist = [];
+errorVecHist = [];
+bmHist = [];
+imuDataHist = [];
+
+seenFeatureVectors = [];
+seenFeatureStructs = {};
+trackedFeatureIdx = [];
+
+numFrames = size(leftImageData.rectImages, 3);
+detectNewPoints = false;
+for i=1:skipFrames:numFrames
+    fprintf('Processing image %d \n', i);
+    %Extract stereo images
+    
+    viLeftImage = uint8(leftImageData.rectImages(:,:,i));
+    viRightImage = uint8(rightImageData.rectImages(:,:,i));
+    
+    if i == 1 || detectNewPoints
+    detectNewPoints = false;     
+    %Detect strongest corners
+    leftPoints = detectSURFFeatures(viLeftImage);
+    rightPoints = detectSURFFeatures(viRightImage);
+    
+    leftPoints = leftPoints.selectStrongest(500);
[…]

### 2015-03-02 Report/project_report.tex
Updated kitti result

--- 2014-12-31 Report/project_report.tex
@@
 
-\author{Lee Clement$^{1}$ and Valentin Peretroukhin$^{1}$% <-this % stops a space
+\author{Lee Clement, Valentin Peretroukhin, Jacob Lambert, and Jonathan Kelly% <-this % stops a space
 %\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{$^{1}$Institute for Aerospace Studies,
-        University of Toronto, Toronto, ON, Canada
-        {\tt \{lee.clement, valentin.peretroukhin\} @robotics.utias.utoronto.ca}
+\thanks{All authors are at the Institute for Aerospace Studies,
+        University of Toronto, Canada
+        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\}@mail.utoronto.ca, jkelly@utias.utoronto.ca}
         }%

### 2015-03-02 Matlab/msckf/MSCKF.m

@@ -11,27 +11,33 @@ clc;
 addpath('utils');
 
 tic
-fileName = '2011_09_26_drive_0001_sync_KLT';
-% load('../datasets/dataset3.mat')
-% load('../datasets/dataset3_fresh_10noisy.mat')
-% load('../datasets/dataset3_fresh_10lessnoisy.mat')
-% load('../datasets/dataset3_fresh_20lessnoisy.mat')
-% load('../datasets/dataset3_fresh_40lessnoisy.mat')
-% load('../datasets/dataset3_fresh_60lessnoisy.mat')
-% load('../datasets/dataset3_fresh_80lessnoisy.mat')
-% load('../datasets/dataset3_fresh_100lessnoisy.mat')
-% load('../datasets/dataset3_fresh2_500lessnoisy.mat')
-% load('../datasets/2011_09_26_drive_0009_sync_KLT.mat')
-% load('../datasets/2011_09_26_drive_0035_sync_KLT.mat');
-% load('../datasets/2011_09_30_drive_0027_sync_KLT.mat');
-% load('../datasets/2011_10_03_drive_0042_sync_KLT.mat');
-% load('../datasets/2011_09_30_drive_0020_sync_KLT.mat');
-load(sprintf('../datasets/%s.mat',fileName));
+dataDir = '../datasets';
+
+% fileName = 'dataset3';
+% fileName = 'dataset3_fresh_10noisy';
+% fileName = 'dataset3_fresh_10lessnoisy';
+% fileName = 'dataset3_fresh_20lessnoisy';
+% fileName = 'dataset3_fresh_40lessnoisy';
+% fileName = 'dataset3_fresh_60lessnoisy';
+% fileName = 'dataset3_fresh_80lessnoisy';
+% fileName = 'dataset3_fresh_100lessnoisy';
+% fileName = 'dataset3_fresh_500lessnoisy';
+% fileName = '2011_09_26_drive_0035_sync_KLT';
+% fileName = '2011_09_26_drive_0005_sync_KLT';
+% fileName = '2011_09_30_drive_0020_sync_KLT';
+% fileName = '2011_09_26_drive_0027_sync_KLT';
+
+% Good KITTI runs
+% fileName = '2011_09_26_drive_0001_sync_KLT'; kStart = 2; kEnd = 98;
+% fileName = '2011_09_26_drive_0051_sync_KLT'; kStart = 2; kEnd = 114;
+fileName = '2011_09_26_drive_0095_sync_KLT'; kStart = 2; kEnd = 139;
+
[…]

### 2015-03-02 Matlab/msckf/propagateMsckfStateAndCovar.m

@@ -20,6 +20,7 @@ function msckfState_prop = propagateMsckfStateAndCovar(msckfState, measurements_
 
     msckfState_prop.imuCovar = Phi * msckfState.imuCovar * Phi' ...
                                 + G * Q_imu * G' * measurements_k.dT; % Leutenegger 2013
+                            
                                     
     % Camera-Camera Covariance
     msckfState_prop.camCovar = msckfState.camCovar;

### 2015-03-02 Matlab/kitti_extraction/sanitizeKITTI_KLT_clean.m
Added 0036

@@ -6,11 +6,11 @@ addpath('utils');
 addpath('utils/devkit');
 
 
-dataBaseDir = '/Users/valentinp/Desktop/KITTI/2011_09_26/2011_09_26_drive_0095_sync';
+dataBaseDir = '/Users/valentinp/Desktop/KITTI/2011_09_26/2011_09_26_drive_0036_sync';
 dataCalibDir = '/Users/valentinp/Desktop/KITTI/2011_09_26';
 
 %% Get ground truth and import data
-frameRange = 1:150;
+frameRange = 1:250;
 %Image data
 leftImageData = loadImageData([dataBaseDir '/image_00'], frameRange);
 rightImageData = loadImageData([dataBaseDir '/image_01'], frameRange);

### 2015-03-02 Matlab/swf/SlidingWindow.m

@@ -7,7 +7,7 @@ close all
 addpath('utils')
 %fileName = '100noisy';
 %load(['../datasets/dataset3_fresh_' fileName '.mat'])
-fileName = '2011_09_26_drive_0051_sync_KLT.mat';
+fileName = '2011_09_26_drive_0095_sync_KLT.mat';
 load(['../datasets/' fileName]);
 
 tic
@@ -29,14 +29,14 @@ T_cv = [C_c_v -C_c_v*rho_v_c_v; 0 0 0 1];
 v_var = 0.05*ones(3,1);
 w_var = 0.05*ones(3,1);
 
-y_var = 5^2*ones(2,1);                 % pixel coord var
+y_var = 2^2*ones(2,1);                 % pixel coord var
 noiseParams.u_var_prime = y_var(1)/fu^2;
 noiseParams.v_var_prime = y_var(2)/fv^2;
 
 %Set up sliding window
 LMLambda = 1e-5;
 lineLambda = 0.25;
-JcostThresh = 5e-1;
+JcostThresh = 5e-2;
 useMonoCamera = true; %If true, only left camera will be used
 
 kappa = 10; %Sliding window size

### 2015-03-02 Report/project_report.tex
Committing Jacob's report changes

@@
 \usepackage{threeparttable}
+\usepackage{cite}
 
@@
         University of Toronto, Canada
-        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\}@mail.utoronto.ca, jkelly@utias.utoronto.ca}
+        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\} @mail.utias.utoronto.ca, jkelly@utias.utoronto.ca}
         }%
@@
 In this work, we compare and contrast two modern solutions: the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
-The MSCKF is a hybrid estimator that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
-The MSCKF has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
-We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
-In order to provide context for the MSCKF and SWF, we first outline the EKF approach to the SLAM problem.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
+    \vspace{0.25cm}
     \centering
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Related Work} \label{sec:relatedwork}
+
+Visual-inertial navigation systems (VINS) have been applied broadly in robotics \cite{Mourikis:2007:ICRA, Sibley:2010:JFR, Huster:2002:OC, Kottas:2013:IROS, Kim:2007:RSS, Li:2013coa,  You:2001:VR}. 
+As such, there is a considerable body of work on the subject and a wide range of optimal filtering algorithms for the Camera-IMU sensor pair. 
+These techniques are often characterized as either \textit{loosely coupled}, when image and IMU measurements are processed individually to produce pose estimates before being fused into a single estimate, or \textit{tightly coupled}, when all information is processed together in a single estimation paradigm.
+The decoupling of inertial and visual measurements in the former case necessarily leads to inexpensive computations, but at the cost of information: processing the IMU measurements separate from the filtering framework makes optimal estimation of biases impossible\cite{Li:2013coa}.
+The filters presented in this work fall in the latter category; tightly coupled methods are preferred when the goal is accurate, consistent visual-inertial navigation.
+
+In both cases, the optimal filtering algorithm used is typically based on the Extended Kalman filter (EKF)(\cite{Huster:2002:OC, You:2001:VR, Kottas:2013:RSS, Kim:2007:RSS, Mourikis:2007:ICRA}), though methods exist that employ unscented Kalman filters \cite{Ebcin:2007:ION}, particle filters \cite{Fox:1999:JAIR, Pupilli:2005:BMVC}, or batch optimization methods\cite{Triggs:2000va}, \cite{Strelow:2004:IJR}.
+The EKF-SLAM is an efficient recursive algorithm for small-scale, online tasks, but maintaining the entire map as part of the state is a weakness for navigation over long distances.
+Indeed, the computational complexity of such algorithm scales far too poorly with the number of features to be applied naively to this problem.
+Additionally, EKF-based methods for visual-inertial navigation are fundamentally inconsistent, that is, the state uncertainties are underestimated.
+It was shown in \cite{Li:2013coa} that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, severely reducing the reliability of such estimators.
+
+The shortcomings of the EKF motivate the use of filters that operate on a subset of the problem.
+Algorithms operating in constant time with respect to map size are particularly desirable.
+The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
+It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features.
+It also takes advantage of \textit{delayed linearization} in attempt to minimize linearization inaccuracies.
+The SWF analyzed in this work is similar as a both a recursive and batch approach, but unlike the VSDF it incorporates a dynamic model and applies batch optimization.
+The SWF is a constant time algorithm as it actively marginalizes states and landmarks to maintain a state vector remains of constant size.
+Batch optimization of a window of states is suboptimal but still leads to exceptional accuracy for its cost, making it applicable on delicate, large scale operations such as planetary landing\cite{Sibley:2010:JFR}.
+The MSCKF is also a sliding window approach, taking advantage of delayed linearization and exploiting the computational efficiency of an EKF.
+It has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}.
+
+In this work, we present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
+In order to provide context for the MSCKF and SWF, we first outline the EKF approach to the SLAM problem.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Extended Kalman Filter (EKF)} \label{sec:EKF}

### 2015-03-02 Matlab/kitti_extraction/sanitizeKITTI_KLT_clean.m
Re-sanitized 0036

@@ -21,7 +21,7 @@ T_wIMU_GT = getGroundTruth(dataBaseDir, imuFrames);
 
 skipFrames = 1;
 minObsNum = 3;
-ransacMaxDistance = 5;
+ransacMaxDistance = 2;
 
 %% Load calibration
 [T_camvelo_struct, P_rect_cam1] = loadCalibration(dataCalibDir);

### 2015-03-02 Matlab/swf/SlidingWindow.m

@@ -7,7 +7,7 @@ close all
 addpath('utils')
 %fileName = '100noisy';
 %load(['../datasets/dataset3_fresh_' fileName '.mat'])
-fileName = '2011_09_26_drive_0095_sync_KLT.mat';
+fileName = '2011_09_26_drive_0036_sync_KLT.mat';
 load(['../datasets/' fileName]);
 
 tic
@@ -36,7 +36,7 @@ noiseParams.v_var_prime = y_var(2)/fv^2;
 %Set up sliding window
 LMLambda = 1e-5;
 lineLambda = 0.25;
-JcostThresh = 5e-2;
+JcostThresh = 1e-2;
 useMonoCamera = true; %If true, only left camera will be used
 
 kappa = 10; %Sliding window size

### 2015-03-02 Matlab/msckf/MSCKF.m
Updated KITTI results with PSD enforcement fix

@@ -31,7 +31,8 @@ dataDir = '../datasets';
 % fileName = '2011_09_26_drive_0001_sync_KLT'; kStart = 2; kEnd = 98;
 % fileName = '2011_09_26_drive_0051_sync_KLT'; kStart = 2; kEnd = 114;
 % fileName = '2011_09_26_drive_0095_sync_KLT'; kStart = 2; kEnd = 139;
-fileName = '2011_09_26_drive_0036_sync_KLT'; kStart = 2; kEnd = 239;
+% fileName = '2011_09_26_drive_0036_sync_KLT'; kStart = 2; kEnd = 239;
+fileName = '2011_09_30_drive_0020_sync_KLT'; kStart = 2; kEnd = 300;
 
 
 load(sprintf('%s/%s.mat',dataDir,fileName));

### 2015-03-02 Matlab/kitti_extraction/sanitizeKITTI_KLT_clean.m
Added full 1000 frame 30-20

@@ -6,11 +6,11 @@ addpath('utils');
 addpath('utils/devkit');
 
 
-dataBaseDir = '/Users/valentinp/Desktop/KITTI/2011_09_26/2011_09_26_drive_0036_sync';
-dataCalibDir = '/Users/valentinp/Desktop/KITTI/2011_09_26';
+dataBaseDir = '/Users/valentinp/Desktop/KITTI/2011_09_30/2011_09_30_drive_0020_sync';
+dataCalibDir = '/Users/valentinp/Desktop/KITTI/2011_09_30';
 
 %% Get ground truth and import data
-frameRange = 1:250;
+frameRange = 1:1000;
 %Image data
 leftImageData = loadImageData([dataBaseDir '/image_00'], frameRange);
 rightImageData = loadImageData([dataBaseDir '/image_01'], frameRange);
@@ -21,7 +21,7 @@ T_wIMU_GT = getGroundTruth(dataBaseDir, imuFrames);
 
 skipFrames = 1;
 minObsNum = 3;
-ransacMaxDistance = 2;
+ransacMaxDistance = 5;
 
 %% Load calibration
 [T_camvelo_struct, P_rect_cam1] = loadCalibration(dataCalibDir);

### 2015-03-02 Matlab/swf/SlidingWindow.m

@@ -7,7 +7,7 @@ close all
 addpath('utils')
 %fileName = '100noisy';
 %load(['../datasets/dataset3_fresh_' fileName '.mat'])
-fileName = '2011_09_26_drive_0036_sync_KLT.mat';
+fileName = '2011_09_30_drive_0020_sync_KLT.mat';
 load(['../datasets/' fileName]);
 
 tic
@@ -26,15 +26,15 @@ vehicleCamTransform.rho_cv_v = rho_v_c_v;
 T_cv = [C_c_v -C_c_v*rho_v_c_v; 0 0 0 1];
 
 %Set up the noise parameters for MSCKF calcGNPosEst
-v_var = 0.05*ones(3,1);
-w_var = 0.05*ones(3,1);
+v_var = 0.1*ones(3,1);
+w_var = 0.1*ones(3,1);
 
-y_var = 2^2*ones(2,1);                 % pixel coord var
+y_var = 5^2*ones(2,1);                 % pixel coord var
 noiseParams.u_var_prime = y_var(1)/fu^2;
 noiseParams.v_var_prime = y_var(2)/fv^2;
 
 %Set up sliding window
-LMLambda = 1e-5;
+LMLambda = 1e-4;
 lineLambda = 0.25;
 JcostThresh = 1e-2;
 useMonoCamera = true; %If true, only left camera will be used
@@ -43,7 +43,7 @@ kappa = 10; %Sliding window size
 maxOptIter = 5;
 
 kStart = 1;
-kEnd = size(y_k_j,2) - kappa - 1 
+kEnd = 300%size(y_k_j,2) - kappa - 1 
 
 
 k1 = kStart;

### 2015-03-02 Report/project_report.tex

@@
         University of Toronto, Canada
-        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\} @mail.utias.utoronto.ca, jkelly@utias.utoronto.ca}
+        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\} @mail.utoronto.ca, jkelly@utias.utoronto.ca}
         }%

### 2015-03-02 Matlab/swf/SlidingWindow.m
added jacob stuff

@@ -34,9 +34,9 @@ noiseParams.u_var_prime = y_var(1)/fu^2;
 noiseParams.v_var_prime = y_var(2)/fv^2;
 
 %Set up sliding window
-LMLambda = 1e-5;
+LMLambda = 1e-4;
 lineLambda = 0.25;
-JcostThresh = 1e-2;
+JcostThresh = 0.1e-2;
 useMonoCamera = true; %If true, only left camera will be used
 
 kappa = 10; %Sliding window size

### 2015-03-02 Matlab/KITTI Trials/plotKITTI.m
KITTI plotting script and plots

@@ -1,0 +1,53 @@
+%% Load data
+clear; close all; clc;
+
+fileName = '2011_09_26_drive_0001';
+% fileName = '2011_09_26_drive_0051';
+% fileName = '2011_09_26_drive_0095';
+% fileName = '2011_09_26_drive_0036';
+
+fileSuffix = '_sync_KLT';
+
+swfData = load(sprintf('SWF_%s%s',fileName,fileSuffix));
+msckfData = load(sprintf('msckf_%s%s',fileName,fileSuffix));
+
+%% Compute RMSE
+swf_trans_rmse = sqrt(mean(swfData.transErrVec.^2,1));
+swf_rot_rmse = sqrt(mean(swfData.rotErrVec.^2,1));
+
+msckf_trans_rmse = sqrt(mean(msckfData.msckf_trans_err.^2,1));
+msckf_rot_rmse = sqrt(mean(msckfData.msckf_rot_err.^2,1));
+
+%% Plot stuff
+figure(1); clf;
+fontSize = 18;
+lineWidth = 2;
+pos = [200,200,640,480];
+xLim = [1, min([size(msckf_rot_rmse,2),size(swf_rot_rmse,2)])];
+
+% Translational RMSE
+subplot(2,1,1);
+plot(swf_trans_rmse,'LineWidth',lineWidth); hold on;
+plot(msckf_trans_rmse,'LineWidth',lineWidth);
+xlim(xLim);
+title(fileName,'Interpreter','none');
+ylabel('Translation RMSE');
+legend('SWF','MSCKF','Location','NorthWest');
+grid minor;
+set(gca,'FontSize',fontSize);
+set(gcf,'Position',pos);
+
+% Rotational RMSE
+subplot(2,1,2);
+plot(swf_rot_rmse,'LineWidth',lineWidth); hold on;
+plot(msckf_rot_rmse,'LineWidth',lineWidth);
+xlim(xLim);
+xlabel('Timestep'); ylabel('Rotational RMSE');
+legend('SWF','MSCKF','Location','NorthWest');
+grid minor;
+set(gca,'FontSize',fontSize);
+set(gcf,'Position',pos);
+
+%% Export figure
+figFileName = ['RMSE_',fileName,'.pdf'];
+export_fig(gcf, figFileName, '-transparent');

### 2015-03-03 Report/project_report.tex
added jacob stuff
Modified figures and made space for KITTI discussion

--- 2015-03-02 Report/project_report.tex
@@
 %\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{All authors are at the Institute for Aerospace Studies,
-        University of Toronto, Canada
-        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\}@mail.utoronto.ca, jkelly@utias.utoronto.ca}
+\thanks{All authors are at the Institute for Aerospace Studies, University of Toronto, Canada
+        {\tt \{lee.clement, v.peretroukhin, jacob.lambert\} @mail.utoronto.ca, jkelly@utias.utoronto.ca}
         }%
@@
 While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
-We present experimental comparisons of the MSCKF and SWF on several datasets with real and synthetic features.
-In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
-In general, our results show the SWF to be more accurate than the MSCKF.
-However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
+We present experimental comparisons of the MSCKF and SWF on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset.
+In particular, we compare the accuracy of the two filters and analyze the effect of feature track length and feature density on the performance of each.
+In general, our results show the SWF to be at least as accurate the MSCKF, however the MSCKF is significantly more efficient and improves in accuracy as more features are tracked. 
 \end{abstract}
@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
+In this work we compare and contrast two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.
 
-In this work, we compare and contrast two modern solutions: the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}.
-The MSCKF is a hybrid estimator that combines the accuracy of the SWF with the computational efficiency of an Extended Kalman Filter (EKF).
-The MSCKF has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}, while the SWF has been applied to planetary landing \cite{Sibley:2010:JFR}.
-We present an experimental comparison of the SWF and MSCKF using different datasets consisting of IMU and camera data with accurate ground truth for both sensor motion and landmark positions.
-In order to provide context for the MSCKF and SWF, we first outline the EKF approach to the SLAM problem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
+    \vspace{0.25cm}
     \centering
@@
     \caption{The sensor head used in our experiments.
-The IMU measures translational and rotational velocities, while the stereo camera measures the positions of point features.
+An IMU measures translational and rotational velocities, while a stereo camera measures the positions of point features.
 In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
@@
 
-\section{Extended Kalman Filter (EKF)} \label{sec:EKF}
-In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
-This technique, sometimes referred to as EKF-SLAM, attempts to track vehicle motion and create a globally consistent map of features by recursively updating the state as new measurements become available.
-%Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
-%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
-This behavior is due to the fact that the dimension of the state grows linearly with the number of features, and the computational cost of inverting the state covariance matrix (during the Kalman gain computation) is cubic in the dimension of the state.
-Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
+\section{Related Work} \label{sec:relatedwork}
+
+Visual-inertial navigation systems (VINS) have been applied broadly in robotics \cite{Mourikis:2007:ICRA, Sibley:2010:JFR, Huster:2002:OC, Kottas:2013:IROS, Kim:2007:RSS, Li:2013coa,  You:2001:VR}, and there is a considerable body of work covering a wide range of estimation algorithms for the camera-IMU sensor pair.
+These techniques are often characterized as either \textit{loosely coupled}, when image and IMU measurements are processed individually before being fused into a single estimate, or \textit{tightly coupled}, when all information is processed together.
+The decoupling of inertial and visual measurements in loosely coupled systems leads to inexpensive computations, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
+In this work, we consider tightly coupled algorithms, which are preferable for accurate and consistent visual-inertial navigation.
+
+For both tightly coupled and loosely coupled VINS, a popular choice of estimator is the Extended Kalman filter (EKF) or one of its variants \cite{Huster:2002:OC, You:2001:VR, Kottas:2013:RSS, Kim:2007:RSS, Mourikis:2007:ICRA}, though methods also exist that employ unscented Kalman filters \cite{Ebcin:2007:ION}, particle filters \cite{Fox:1999:JAIR, Pupilli:2005:BMVC}, or batch optimization methods\cite{Triggs:2000va,Strelow:2004:IJR}.
+EKF-SLAM is an efficient recursive algorithm for small-scale, online tasks, but maintaining the entire map as part of the state is a weakness for navigation over long distances.
+Indeed, the computational complexity of such algorithm scales far too poorly with the number of features to be applied naively to this problem.
+Additionally, EKF-based VINS are fundamentally inconsistent, that is, the state uncertainties are underestimated.
+Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, severely reducing the reliability of such estimators.
 
@@
 Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.
-By locking in past poses, the EKF-SLAM formulation condemns itself to sub-optimally estimating both vehicle motion and feature positions.
+By locking in past poses, EKF-SLAM sub-optimally estimates both vehicle motion and feature positions.
+
+The shortcomings of the EKF motivate the use of filters that operate on a subset of the problem.
+Algorithms operating in constant time with respect to map size are particularly desirable.
+The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
+It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features.
+It also takes advantage of \textit{delayed linearization} in attempt to minimize linearization errors.
+The SWF analyzed in this work is similar as a both a recursive and batch approach, but unlike the VSDF it incorporates a motion model and applies batch optimization.
+The SWF operates in constant time by limiting the optimization problem to a set number of states at all times.
+Batch optimization over a window of states is suboptimal but still leads to exceptional accuracy for its cost, making it useful for delicate, large scale operations such as planetary landing\cite{Sibley:2010:JFR}.
+The MSCKF \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF in the sense that it maintains a variable window of poses and applies batch updates using all observations of each landmark.
+It has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Extended Kalman Filter (EKF)} \label{sec:EKF}
+% In order to provide context for the MSCKF and SWF, we first outline the EKF approach to the SLAM problem.
+% In the Extended Kalman Filter (EKF) solution, vehicle poses and feature positions are simultaneously estimated at each time step by augmenting the filter state with feature positions.
+% This technique, sometimes referred to as EKF-SLAM, attempts to track vehicle motion and create a globally consistent map of features by recursively updating the state as new measurements become available.
+% %Features are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
+% %In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.
+
+% Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
+% This behavior is due to the fact that the dimension of the state grows linearly with the number of features, and the computational cost of inverting the state covariance matrix (during the Kalman gain computation) is cubic in the dimension of the state.
+% Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.
+
+% Another limitation of EKF-SLAM is that it is \textit{forgetful}.
+% Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.
+% By locking in past poses, the EKF-SLAM formulation condemns itself to sub-optimally estimating both vehicle motion and feature positions.
 
@@
 \begin{align}
-    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
-    \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \\
-    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
-    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
+    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} 
+        & \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \notag \\
+    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} 
+        & \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
 \end{align}
@@
 we can define three functions
+% \begin{align}
+%     \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
+%     & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
+%     & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
+%             C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
+%             C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
+%         \ebm
+% \end{align}
 \begin{align}
     \bbm h_1 (\hat{\Vec{y}}) \\ h_2 (\hat{\Vec{y}}) \\ h_3 (\hat{\Vec{y}}) \ebm 
-    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i} \notag \\
-    & = \bbm C_{11} \alpha + C_{12} \beta + C_{13} + \gamma \hat{x}_{C_i}^{C_1 C_i} \\[0.5em]
-            C_{21} \alpha + C_{22} \beta + C_{23} + \gamma \hat{y}_{C_i}^{C_1 C_i} \\[0.5em]
-            C_{31} \alpha + C_{32} \beta + C_{33} + \gamma \hat{z}_{C_i}^{C_1 C_i} 
-        \ebm
+    & = \hat{\Vec{C}}_{i1} \bbm \alpha \\ \beta \\ 1 \ebm + \gamma  \hat{\Vec{p}}_{C_i}^{C_1 C_i}
 \end{align}
@@
 with $\Vec{R}_i^{(j)} = \text{diag}\left\{\sigma^2_{u'}, \sigma^2_{v'}\right\}$.
-The Jacobian elements are given by
-\begin{align}
-\frac{\partial \Vec e}{\partial \alpha} &= 
-    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_1 \\[0.5em]
-                        -\frac{\partial h_2}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_2 \ebm
-\\[0.5em]
-\frac{\partial \Vec e}{\partial \beta} &= 
-    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_1 \\[0.5em] 
-                        -\frac{\partial h_2}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_2 \ebm
-\\[0.5em]
-\frac{\partial \Vec e}{\partial \gamma} &= 
-    \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_1 \\[0.5em] 
-                        -\frac{\partial h_2}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_2 \ebm
-\end{align}
-where
-\begin{align}
-\frac{\partial h_1}{\partial \alpha} &= C_{11}, & \frac{\partial h_1}{\partial \beta} &= C_{12}, & \frac{\partial h_1}{\partial \gamma} &= \hat{x}_{C_i}^{C_1 C_i} \notag \\
-\frac{\partial h_2}{\partial \alpha} &= C_{21}, & \frac{\partial h_2}{\partial \beta} &= C_{22}, & \frac{\partial h_2}{\partial \gamma} &= \hat{y}_{C_i}^{C_1 C_i} \\
-\frac{\partial h_3}{\partial \alpha} &= C_{31}, & \frac{\partial h_3}{\partial \beta} &= C_{32}, & \frac{\partial h_3}{\partial \gamma} &= \hat{z}_{C_i}^{C_1 C_i} \notag .
-\end{align}
+% The Jacobian elements are given by
+% \begin{align}
+% \frac{\partial \Vec e}{\partial \alpha} &= 
+%     \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_1 \\[0.5em]
+%                         -\frac{\partial h_2}{\partial \alpha} h_3 + \frac{\partial h_3}{\partial \alpha} h_2 \ebm
+% \\[0.5em]
+% \frac{\partial \Vec e}{\partial \beta} &= 
+%     \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_1 \\[0.5em] 
+%                         -\frac{\partial h_2}{\partial \beta} h_3 + \frac{\partial h_3}{\partial \beta} h_2 \ebm
+% \\[0.5em]
+% \frac{\partial \Vec e}{\partial \gamma} &= 
+%     \frac{1}{h_3^2} \bbm -\frac{\partial h_1}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_1 \\[0.5em] 
+%                         -\frac{\partial h_2}{\partial \gamma} h_3 + \frac{\partial h_3}{\partial \gamma} h_2 \ebm
+% \end{align}
+% where
+% \begin{align}
+% \frac{\partial h_1}{\partial \alpha} &= C_{11}, & \frac{\partial h_1}{\partial \beta} &= C_{12}, & \frac{\partial h_1}{\partial \gamma} &= \hat{x}_{C_i}^{C_1 C_i} \notag \\
+% \frac{\partial h_2}{\partial \alpha} &= C_{21}, & \frac{\partial h_2}{\partial \beta} &= C_{22}, & \frac{\partial h_2}{\partial \gamma} &= \hat{y}_{C_i}^{C_1 C_i} \\
+% \frac{\partial h_3}{\partial \alpha} &= C_{31}, & \frac{\partial h_3}{\partial \beta} &= C_{32}, & \frac{\partial h_3}{\partial \gamma} &= \hat{z}_{C_i}^{C_1 C_i} \notag .
+% \end{align}
 
@@
 \begin{align}
-    \hat{\Vec{p}}_{C_i}^{f_j C_i} &= \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T = \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
+    \hat{\Vec{p}}_{C_i}^{f_j C_i} &= \bbm \hat{X}^{(j)}_{C_i} & \hat{Y}^{(j)}_{C_i} & \hat{Z}^{(j)}_{C_i} \ebm ^T \notag \\
+                                 &= \hat{\Vec{C}}_{C_i G} \left( \hat{\Vec{p}}_G^{f_j G} - \hat{\Vec{p}}_G^{C_i G} \right).
 \end{align}
@@
 
-\setcounter{figure}{2}
-\begin{figure*}
+% \setcounter{figure}{2}
+\begin{figure}
     \centering
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/NST_trans}
-        \label{fig:nullspace:trans_err}
-    }
-    ~
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/NST_rot}
-        \label{fig:nullspace:rot_err}
-    }
-
+    \includegraphics[width=0.45\textwidth]{figs/NST_RMSE}    
     \caption{Effect of correlation between measurement noise and filter state on MSCKF performance.
@@
     \label{fig:nullspace}
-\end{figure*}
+\end{figure}
 
@@
 \section{Experiments} \label{sec:experiments}
-We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
-The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions (Figure \ref{fig:trajectory_groundtruth}).
+We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as several runs from the KITTI Dataset \cite{Geiger:2013:IJRR}.
+Both datasets include data streams from a stereo camera and IMU.
 Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
@@
 
-\setcounter{figure}{1}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth}
-    \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
-    \label{fig:trajectory_groundtruth}
-\end{figure}
+% \setcounter{figure}{1}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth}
+%     \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
+%     \label{fig:trajectory_groundtruth}
+% \end{figure}
 
-\subsection{Experiment 1: Many visible features}
+\subsection{Starry Night Dataset}
+The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
+This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
+
+\subsubsection{Several visible features}
 In the first experiment, we compared the SWF and MSCKF for various parameter settings on an interval with many visible features.
-Figure \ref{fig:nVisible_int1} shows the number of visible features in the first interval.
+% Figure \ref{fig:nVisible_int1} shows the number of visible features in the first interval.
 At least three features are visible in $64\%$ of this interval, and the number of visible features is frequently in excess of ten.
 
-\setcounter{figure}{3}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/nVisible_500_1000}
-    \caption{Number of visible features in the interval $k\in[500,1000]$.
+% \setcounter{figure}{3}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.45\textwidth]{figs/nVisible_500_1000}
+%     \caption{Number of visible features in the interval $k\in[500,1000]$.
 Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
-    \label{fig:nVisible_int1}
-\end{figure}
+%     \label{fig:nVisible_int1}
+% \end{figure}
 
@@
 
-\setcounter{figure}{5}
+% \setcounter{figure}{5}
 \begin{figure*}
     \centering
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
-        \label{fig:comp_1:trans_err}
+    \subfigure[At least three features are visible in $64\%$ of the interval $k\in(500,1000)$. The number of visible features is frequently exceeds ten.] {
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
+        \label{fig:comp_1}
     }
     ~
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
-        \label{fig:comp_1:rot_err}
+    \subfigure[At least three features are visible in only $56\%$ of the interval $k\in(1215,1715)$. The number of visible features rarely exceeds ten.] {
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
+        \label{fig:comp_2}
     }
-
-    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[500,1000]$.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
-    \label{fig:comp_1}
+    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset.
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
 \end{figure*}
 
-\subsection{Experiment 2: Fewer visible features}
+\subsubsection{Fewer visible features}
 In the second experiment, we compared the SWF and MSCKF for various parameter settings on an interval with few visible features.
-Figure \ref{fig:nVisible_int2} shows the number of visible features in the second interval.
+% Figure \ref{fig:nVisible_int2} shows the number of visible features in the second interval.
 At least three features are visible in only $56\%$ of this interval, and the number of visible features rarely exceeds ten.
@@
 
-\setcounter{figure}{4}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/nVisible_1215_1715}
-    \caption{Number of visible features in the interval $k\in[1215,1715]$.
+% \setcounter{figure}{4}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.45\textwidth]{figs/nVisible_1215_1715}
+%     \caption{Number of visible features in the interval $k\in[1215,1715]$.
 Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
-    \label{fig:nVisible_int2}
-\end{figure}
+%     \label{fig:nVisible_int2}
+% \end{figure}
 
-\setcounter{figure}{6}
-\begin{figure*}
-    \centering
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
-        \label{fig:comp_2:trans_err}
-    }
-    ~
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
-        \label{fig:comp_2:rot_err}
-    }
+% \setcounter{figure}{6}
 
-    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on the dataset interval $k\in[1215,1715]$.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the number next to ``SWF'' refers to the number of states in the sliding window.}
-    \label{fig:comp_2}
-\end{figure*}
-
-\subsection{Experiment 3: Effect of feature density}
+\subsubsection{Effect of feature density}
 In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
@@
 In each dataset, the IMU data remained unchanged from the original dataset and we corrupted the synthetic camera measurements with zero-mean Gaussian noise.
-Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
+% Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
 By generating longer feature tracks and increasing the number of features visible at each time step, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.
 
-\setcounter{figure}{8}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
-    \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
-    \label{fig:trajectory_groundtruth_100lessnoisy}
-\end{figure}
+% \setcounter{figure}{8}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
+%     \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
+%     \label{fig:trajectory_groundtruth_100lessnoisy}
+% \end{figure}
 
@@
 Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
-\setcounter{figure}{7}
-\begin{figure*}
+% \setcounter{figure}{7}
+\begin{figure}
     \centering
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
-        \label{fig:comp_3:trans_err}
-    }
-    ~
-    \subfigure {
-        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
-        \label{fig:comp_3:rot_err}
-    }
+    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}
 
@@
     \label{fig:comp_3}
-\end{figure*}
+\end{figure}
 
@@
 The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
-
-% \begin{table*}[t]
-%   \centering
-%     \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of the MSCKF and SWF on synthetic datasets.}
-%   \begin{tabular}{*{11}{c}}
-%               && \multicolumn{4}{c}{MSCKF} & & \multicolumn{4}{c}{SWF} \B \\ \cline{3-6} \cline{8-11}
-% Feature Count && Pos. ARMSE    & Rot. ARMSE & ANEES     & Compute Time && Pos. ARMSE   & Rot. ARMSE     & ANEES & Compute Time \T\B \\ \hline
-% 40            && 0.2672        & 0.1378        & 10.18  & 12.19 s      && 0.1750        & 0.0495        & 2280  & 114.3 s   \T \\
-% 60            && 0.2550        & 0.1247        & 12.03  & 14.64 s      && 0.1687        & 0.0377        & 2093  & 175.9 s   \\
-% 100           && 0.2304        & 0.0952        & 16.76  & 20.58 s      && 0.1755        & 0.0481        & 2013  & 245.3 s   \B \\ \hline
-%   \end{tabular}
-%   \label{tab:rmse_nees}
-% \end{table*}
 
@@
     \label{tab:rmse_nees}
-    \vspace{-0.4cm}
+    % \vspace{-0.4cm}
 \end{table}
 
+\subsection{KITTI Dataset}
+\begin{figure*}
+    \centering
+    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0001} traverse.] {
+        \centering
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0001}
+    }
+    ~
+    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0036} traverse.] {
+        \centering
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0036}
+    }
+
+    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0051} traverse.] {
+        \centering
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0051}
+    }
+    ~
+    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0095} traverse.] {
+        \centering
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0095}
+    }
+    \caption{RMSE over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling 534 m of urban driving.}
+    \label{fig:kitti}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{-0.1cm}
+% \vspace{-0.1cm}
 \section{Conclusions} \label{sec:conclusions}

### 2015-03-03 Matlab/KITTI Trials/plotKITTI.m

@@ -1,16 +1,21 @@
 %% Load data
 clear; close all; clc;
 
-fileName = '2011_09_26_drive_0001';
+% fileName = '2011_09_26_drive_0001';
 % fileName = '2011_09_26_drive_0051';
 % fileName = '2011_09_26_drive_0095';
-% fileName = '2011_09_26_drive_0036';
+fileName = '2011_09_26_drive_0036';
 
 fileSuffix = '_sync_KLT';
 
 swfData = load(sprintf('SWF_%s%s',fileName,fileSuffix));
 msckfData = load(sprintf('msckf_%s%s',fileName,fileSuffix));
 
+dists = [0,cumsum(sqrt(sum(diff(msckfData.p_I_G_GT,1,2).^2,1)))];
+swfData.transErrVec = swfData.transErrVec(:,1:size(dists,2));
+swfData.rotErrVec = swfData.rotErrVec(:,1:size(dists,2));
+
+dists(end)
 %% Compute RMSE
 swf_trans_rmse = sqrt(mean(swfData.transErrVec.^2,1));
 swf_rot_rmse = sqrt(mean(swfData.rotErrVec.^2,1));
@@ -23,16 +28,16 @@ imu_rot_rmse = sqrt(mean(msckfData.imu_rot_err.^2,1));
 
 %% Plot stuff
 figure(1); clf;
-fontSize = 16;
+fontSize = 14;
 lineWidth = 2;
 pos = [200,200,640,480];
-xLim = [1, min([size(msckf_rot_rmse,2),size(swf_rot_rmse,2)])];
+xLim = [dists(1), dists(end)];
 
 % Translational RMSE
 subplot(2,1,1);
-plot(imu_trans_rmse,'-k','LineWidth',lineWidth); hold on;
-plot(swf_trans_rmse,'-r','LineWidth',lineWidth); 
-plot(msckf_trans_rmse,'-b','LineWidth',lineWidth);
+plot(dists,imu_trans_rmse,'-k','LineWidth',lineWidth); hold on;
+plot(dists,swf_trans_rmse,'-r','LineWidth',lineWidth); 
+plot(dists,msckf_trans_rmse,'-b','LineWidth',lineWidth);
 xlim(xLim);
 title(fileName,'Interpreter','none');
 ylabel('Trans. RMSE (m)');
@@ -43,11 +48,11 @@ set(gcf,'Position',pos);
 
 % Rotational RMSE
[…]

### 2015-03-03 Matlab/plotting/featureDensity_RMSE_plots.m

@@ -37,11 +37,12 @@ imu_rot_rmse = sqrt(mean(imu.msckf_rot_err.^2,1));
 
 %% Plot Stuff
 figure(1); clf;
-fontSize = 16;
+fontSize = 14;
 lineWidth = 2;
 pos = [200,200,640,480];
 xLim = [kStart,kEnd];
 k = kStart:kEnd;
+xticks = linspace(kStart,kEnd,11);
 
 set(gcf, 'Position', pos);
 
@@ -55,6 +56,7 @@ plot(k,swf40_trans_rmse, '-g', 'LineWidth', lineWidth);
 plot(k,swf60_trans_rmse, '--g', 'LineWidth', lineWidth);
 plot(k,swf100_trans_rmse, '-.g', 'LineWidth', lineWidth);
 xlim(xLim);
+set(gca,'XTick',xticks);
 legend('IMU Only','MSCKF 40', 'MSCKF 60', 'MSCKF 100', 'SWF 40','SWF 60','SWF 100', 'Location', 'northwest');
 title('Feature Density Comparison');
 ylabel('Trans. RMSE (m)')
@@ -71,6 +73,7 @@ plot(k,swf40_rot_rmse, '-g', 'LineWidth', lineWidth);
 plot(k,swf60_rot_rmse, '--g', 'LineWidth', lineWidth);
 plot(k,swf100_rot_rmse, '-.g', 'LineWidth', lineWidth);
 xlim(xLim);
+set(gca,'XTick',xticks);
 legend('IMU Only','MSCKF 40', 'MSCKF 60', 'MSCKF 100', 'SWF 40','SWF 60','SWF 100', 'Location', 'northwest');
 ylabel('Rot. RMSE (Axis-Angle)');
 xlabel('Timestep');

### 2015-03-03 Matlab/plotting/nullspace_RMSE_plots.m

@@ -4,6 +4,8 @@ clear all; close all; clc;
 nst_on = load('msckf_NST_on.mat');
 nst_off = load('msckf_NST_off.mat');
 
+kStart = 500; kEnd = 1000;
+
 %% Compute RMSE
 nst_off_trans_rmse = sqrt(mean(nst_off.msckf_trans_err.^2,1));
 nst_off_rot_rmse = sqrt(mean(nst_off.msckf_rot_err.^2,1));
@@ -13,19 +15,22 @@ nst_on_rot_rmse = sqrt(mean(nst_on.msckf_rot_err.^2,1));
 
 %% Plot stuff
 figure(1); clf;
-fontSize = 16;
+fontSize = 14;
 lineWidth = 2;
 pos = [200,200,640,480];
-xLim = [1, min([size(nst_on_rot_rmse,2),size(nst_off_rot_rmse,2)])];
+xLim = [kStart, kEnd];
+k = kStart:kEnd;
+xticks = linspace(kStart,kEnd,11);
 
 transLim = [0 0.6];
 rotLim = [0 0.25];
 
 % Translational RMSE
 subplot(2,1,1);
-plot(nst_off_trans_rmse, '-r','LineWidth',lineWidth); hold on;
-plot(nst_on_trans_rmse, '-b','LineWidth',lineWidth);
+plot(k,nst_off_trans_rmse, '-r','LineWidth',lineWidth); hold on;
+plot(k,nst_on_trans_rmse, '-b','LineWidth',lineWidth);
 xlim(xLim);
+set(gca,'XTick',xticks);
 % ylim(transLim)
 legend('Noise Correlated to State', 'Noise De-correlated from State', ...
     'Location','northwest');
@@ -37,9 +42,10 @@ box on;
 
 % Rotational RMSE
 subplot(2,1,2);
-plot(nst_off_rot_rmse,'-r','LineWidth',lineWidth); hold on;
-plot(nst_on_rot_rmse,'-b','LineWidth',lineWidth);
+plot(k,nst_off_rot_rmse,'-r','LineWidth',lineWidth); hold on;
+plot(k,nst_on_rot_rmse,'-b','LineWidth',lineWidth);
 xlim(xLim);
+set(gca,'XTick',xticks);
 % ylim(rotLim)
 legend('Noise Correlated to State', 'Noise De-correlated from State', ...
     'Location','northwest');

### 2015-03-03 Matlab/plotting/windowSize_RMSE_plots.m

@@ -2,24 +2,24 @@
 clear; close all; clc;
 
 % 500-1000
-% swf10 = load('swf_500_1000_10_dataset3.mat');
-% swf50 = load('swf_500_1000_50_dataset3.mat');
-% swf100 = load('swf_500_1000_100_dataset3.mat');
-% msckf5 = load('msckf_500_1000_min5_maxInf');
-% msckf10 = load('msckf_500_1000_min10_max50');
-% msckf20 = load('msckf_500_1000_min20_max100');
-% imu = load('imu_500_1000.mat');
-% kStart = 500; kEnd = 1000;
+swf10 = load('swf_500_1000_10_dataset3.mat');
+swf50 = load('swf_500_1000_50_dataset3.mat');
+swf100 = load('swf_500_1000_100_dataset3.mat');
+msckf5 = load('msckf_500_1000_min5_maxInf');
+msckf10 = load('msckf_500_1000_min10_max50');
+msckf20 = load('msckf_500_1000_min20_max100');
+imu = load('imu_500_1000.mat');
+kStart = 500; kEnd = 1000;
 
 % 1215-1715
-swf10 = load('swf_1215_1715_10_dataset3.mat');
-swf50 = load('swf_1215_1715_50_dataset3.mat');
-swf100 = load('swf_1215_1715_100_dataset3.mat');
-msckf5 = load('msckf_1215_1715_min5_maxInf');
-msckf10 = load('msckf_1215_1715_min10_max50');
-msckf20 = load('msckf_1215_1715_min20_max100');
-imu = load('imu_1215_1715.mat');
-kStart = 1215; kEnd = 1715;
+% swf10 = load('swf_1215_1715_10_dataset3.mat');
+% swf50 = load('swf_1215_1715_50_dataset3.mat');
+% swf100 = load('swf_1215_1715_100_dataset3.mat');
+% msckf5 = load('msckf_1215_1715_min5_maxInf');
+% msckf10 = load('msckf_1215_1715_min10_max50');
+% msckf20 = load('msckf_1215_1715_min20_max100');
+% imu = load('imu_1215_1715.mat');
+% kStart = 1215; kEnd = 1715;
 
 %% Compute RMSE
 swf10_trans_rmse = sqrt(mean(swf10.swf_trans_err.^2,1));
[…]

### 2015-03-03 Matlab/KITTI Trials/ANEES.m
First final draft.

@@ -1,6 +1,6 @@
 function avgNEES = ANEES(trans_err, rot_err, err_sigma)
 
-    stateErr = [trans_err; rot_err];
+    stateErr = [rot_err;trans_err];
     stateVar = err_sigma.^2;
     avgNEES = 0;
     stepNum = size(stateErr, 2);

### 2015-03-03 Matlab/KITTI Trials/plotKITTI.m

@@ -1,14 +1,14 @@
 %% Load data
 clear; close all; clc;
 
-fileName = '2011_09_26_drive_0001';
+% fileName = '2011_09_26_drive_0001';
+fileName = '2011_09_26_drive_0036';
 % fileName = '2011_09_26_drive_0051';
 % fileName = '2011_09_26_drive_0095';
-% fileName = '2011_09_26_drive_0036';
 
 fileSuffix = '_sync_KLT';
 
-swfData = load(sprintf('SWF_%s%s',fileName,fileSuffix));
+swfData = load(sprintf('SWF_RERUN_%s%s',fileName,fileSuffix));
 msckfData = load(sprintf('msckf_%s%s',fileName,fileSuffix));
 
 dists = [0,cumsum(sqrt(sum(diff(msckfData.p_I_G_GT,1,2).^2,1)))];
@@ -27,9 +27,16 @@ imu_trans_rmse = sqrt(mean(msckfData.imu_trans_err.^2,1));
 imu_rot_rmse = sqrt(mean(msckfData.imu_rot_err.^2,1));
 
 %% Compute ANEES
+msckf_err_sigma = msckfData.err_sigma;
 msckf_anees = ANEES(msckfData.msckf_trans_err, msckfData.msckf_rot_err,...
-    msckfData.err_sigma);
-swf_anees = 0;
+    msckf_err_sigma);
+
+swf_err_sigma = [swfData.sigma_th1;swfData.sigma_th2;swfData.sigma_th3; ...
+                    swfData.sigma_x;swfData.sigma_y;swfData.sigma_y];
+swf_anees = ANEES(swfData.transErrVec, swfData.rotErrVec, swf_err_sigma);
+
+imu_err_sigma = msckfData.err_sigma_imu;
+imu_anees = ANEES(msckfData.imu_trans_err, msckfData.imu_rot_err, imu_err_sigma);
 
 %% Plot stuff
 figure(1); clf;
@@ -70,6 +77,8 @@ export_fig(gcf, figFileName, '-nocrop','-transparent');
 
 %% Stats
 fprintf('Total distance: %f \n', dists(end));
+fprintf('IMU:\n\t trans armse %f \n\t rot armse %f \n\t anees %f\n',...
+    mean(imu_trans_rmse),mean(imu_rot_rmse),imu_anees);
 fprintf('MSCKF:\n\t trans armse %f \n\t rot armse %f \n\t anees %f\n',...
[…]

### 2015-03-03 Matlab/msckf/MSCKF.m

@@ -30,9 +30,9 @@ dataDir = '../datasets';
 
 % Good KITTI runs
 % fileName = '2011_09_26_drive_0001_sync_KLT'; kStart = 2; kEnd = 98;
+fileName = '2011_09_26_drive_0036_sync_KLT'; kStart = 2; kEnd = 239;
 % fileName = '2011_09_26_drive_0051_sync_KLT'; kStart = 2; kEnd = 114;
 % fileName = '2011_09_26_drive_0095_sync_KLT'; kStart = 2; kEnd = 139;
-fileName = '2011_09_26_drive_0036_sync_KLT'; kStart = 2; kEnd = 239;
 
 
 load(sprintf('%s/%s.mat',dataDir,fileName));
@@ -73,9 +73,9 @@ bv_var_init = 1e-6 * ones(1,3);        % init vel bias var
 noiseParams.initialIMUCovar = diag([q_var_init, bg_var_init, bv_var_init, p_var_init]);
    
 % MSCKF parameters
-msckfParams.minTrackLength = 3;        % Set to inf to dead-reckon only
+msckfParams.minTrackLength = 5;        % Set to inf to dead-reckon only
 msckfParams.maxTrackLength = Inf;      % Set to inf to wait for features to go out of view
-msckfParams.maxGNCostNorm  = 1e-3;     % Set to inf to allow any triangulation, no matter how bad
+msckfParams.maxGNCostNorm  = 1e-2;     % Set to inf to allow any triangulation, no matter how bad
 msckfParams.minRCOND       = 1e-12;
 msckfParams.doNullSpaceTrick = true;
 msckfParams.doQRdecomp = true;
@@ -374,6 +374,7 @@ p_C_G_GT = NaN(3, kNum);
 theta_CG_err = NaN(3,kNum);
 theta_CG_err_imu = NaN(3,kNum);
 err_sigma = NaN(6,kNum); % cam state is ordered as [rot, trans]
+err_sigma_imu = NaN(6,kNum);
 % 
 tPlot = NaN(1, kNum);
 % 
@@ -389,6 +390,8 @@ for k = 1:kNum
                         * ( C_c_v * axisAngleToRotMat(theta_vk_i(:,kStart+k-1)) )' );
       
     err_sigma(:,k) = prunedStates{k}.sigma;
+    imusig = sqrt(diag(msckfState_imuOnly{state_k}.imuCovar));
[…]

### 2015-03-03 Report/project_report.tex
Whoops, forgot to add the report

@@
 While the SWF estimates feature positions as part of the filter state itself, the MSCKF optimizes feature positions in a separate procedure without including them in the filter state.
-We present experimental comparisons of the MSCKF and SWF on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset.
-In particular, we compare the accuracy of the two filters and analyze the effect of feature track length and feature density on the performance of each.
-In general, our results show the SWF to be at least as accurate the MSCKF, however the MSCKF is significantly more efficient and improves in accuracy as more features are tracked. 
+We present experimental characterizations and comparisons of the MSCKF and SWF on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset.
+In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
+In general, our results show the SWF to be more accurate than the MSCKF.
+However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
 \end{abstract}
@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
-In this work we compare and contrast two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.
+In this work we characterize, compare, and contrast the performance of two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Sliding Window Filter (SWF)} \label{sec:SWF}
+\section{Sliding Window Filter} \label{sec:SWF}
 In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
@@
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
-By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
+By varying the extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
 
-However, the hard cut-off of the SWF may result in only some measurements of a particular feature contributing to the optimization.
-As a result, the filter may not maximally constrain some vehicle poses, and hence localization may be less accurate than we could expect from the full batch solution.
+However, the hard cut-off of the SWF may result in only a subset of measurements of each feature contributing to the optimization.
+As a result, the filter may not maximally constrain some poses, and hence localization may be less accurate than we could expect from the full batch solution.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Multi-State Constraint Kalman Filter (MSCKF)} \label{sec:MSCKF}
+\section{Multi-State Constraint Kalman Filter} \label{sec:MSCKF}
 The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
@@
 
-\subsection{MSCKF State Correction Equations}
+\subsection{MSCKF State Correction}
 Now that we have estimated the positions of any features to be used in the state update, we can apply the corresponding motion constraints to the window of poses from which each feature was observed.
@@
 \end{align}
-where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $C_i$ with respect to the filter state and the position of the feature, respectively.
-These are given by
+where $\Vec{H}_{\Vec{x}, i}^{(j)}$ and $\Vec{H}_{f, i}^{(j)}$ are the Jacobians of the measurement of feature $f_j$ from camera pose $C_i$ with respect to the filter state and the position of the feature, respectively:
 \begin{align}
@@
 The correlation between the measurement noise and the state causes the filter estimates to drift further from the true values than with a modified, de-correlated measurement error.
-As an aside, note that the SWF does not have this problem because feature positions are part of the state itself and are jointly optimized along with the poses.
+As an aside, we note that the SWF does not have this problem because feature positions are jointly optimized along with the poses.
 
@@
 \section{Experiments} \label{sec:experiments}
-We compared the performance of the MSCKF and SWF algorithms on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as several runs from the KITTI Dataset \cite{Geiger:2013:IJRR}.
+We compared the performance of the MSCKF and SWF on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as 530 m of urban driving from the KITTI Dataset \cite{Geiger:2013:IJRR}.
 Both datasets include data streams from a stereo camera and IMU.
-Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
+Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using images from the left camera only.
 
@@
 The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-This dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
+This dataset is well-suited to evaluating SLAM algorithms since ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
+We conducted three experiments on this dataset to compare the effect of feature visibility and window size on each algorithm.
+We discuss each of these in turn.
+
+\begin{figure*}
+    \centering
+    \subfigure[At least three features are visible in $64\%$ of the interval $(500,1000)$. The number of visible features is frequently exceeds ten.] {
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
+        \label{fig:comp_1}
+    }
+    ~
+    \subfigure[At least three features are visible in only $56\%$ of the interval $(1215,1715)$. The number of visible features rarely exceeds ten.] {
+        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
+        \label{fig:comp_2}
+    }
+    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset.
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
+\end{figure*}
 
@@
 
-Figure \ref{fig:comp_1} shows absolute translational and rotational errors for pure IMU integration and various parameter settings for the MSCKF and SWF.
-For the MSCKF, we varied both the maximum number of observations before triggering an update and the minimum number of observations for a feature track to be used in an update.
-For the SWF, we varied the number of states in the sliding window.
+Figure \ref{fig:comp_1} shows translational and rotational root mean squared errors (RMSE) for pure IMU integration and various parameter settings for the MSCKF and SWF.
+For the MSCKF, we varied the maximum number of observations before triggering an update and the minimum number of observations for a feature track to be used in an update.
+For the SWF, we varied the number of states in the window.
 These parameters did not significantly affect the accuracy of either filter, however Figure \ref{fig:comp_1} shows some small gains in accuracy for larger feature track lengths and window sizes.
 
-On this interval, the SWF slightly outperforms the MSCKF in terms of translational accuracy in the $y$- and $z$-directions, and significantly outperforms the MSCKF in terms of rotational accuracy about the $x$- and $y$-axes.
+On this interval, the SWF outperforms the MSCKF in terms of both translational and rotational RMSE.
 The MSCKF does not perform much better than pure IMU integration on this interval, likely due to the overall low number of feature tracks in this dataset.
-
-% \setcounter{figure}{5}
-\begin{figure*}
-    \centering
-    \subfigure[At least three features are visible in $64\%$ of the interval $k\in(500,1000)$. The number of visible features is frequently exceeds ten.] {
-        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-500-1000}
-        \label{fig:comp_1}
-    }
-    ~
-    \subfigure[At least three features are visible in only $56\%$ of the interval $k\in(1215,1715)$. The number of visible features rarely exceeds ten.] {
-        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-WindowSize-1215-1715}
-        \label{fig:comp_2}
-    }
-    \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
-\end{figure*}
 
@@
 
-Figure \ref{fig:comp_2} shows absolute translational and rotational errors for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
+Figure \ref{fig:comp_2} shows translational and rotational RMSE for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
 Compared to Interval 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy in the $y$-direction.
@@
 \subsubsection{Effect of feature density}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}
+
+    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$.
+The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
+    \label{fig:comp_3}
+    \vspace{-0.4cm}
+\end{figure}
+
 In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
 We constructed each dataset so that the larger maps contained the same features as the smaller maps, plus additional features to make up the difference.
-In each dataset, the IMU data remained unchanged from the original dataset and we corrupted the synthetic camera measurements with zero-mean Gaussian noise.
+In each dataset, we retained the IMU data from the original dataset and corrupted the synthetic camera measurements with zero-mean Gaussian noise.
 % Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
@@
 
-Figure \ref{fig:comp_3} compares translational and rotational Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
+Figure \ref{fig:comp_3} compares translational and rotational RMSE of the MSCKF, SWF, and pure IMU integration on three synthetic datasets with 40, 60, and 100 features.
 With more features, both algorithms consistently outperform pure IMU integration, and the SWF outperforms the MSCKF by a wide margin on all three datasets.
@@
 Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
-% \setcounter{figure}{7}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}
 
-    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three different synthetic maps generated based on the IMU data in the dataset interval $k\in[1215,1715]$.
-The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
-    \label{fig:comp_3}
-\end{figure}
-
-Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
-As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
+Table \ref{tab:rmse_nees} summarizes each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
+As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation and translation.
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
@@
     \centering
-    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $k\in[1215,1715]$.}
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $(1215,1715)$.}
     \begin{threeparttable}
@@
                     &                   & 40        & 60        & 100       \T\B    \\ \hline
-        IMU Only    & Position ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
-                    & Rotation ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
+        IMU Only    & Trans. ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
+                    & Rot. ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                     & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                     & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
-        MSCKF (20-100)       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
-                    & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
+        MSCKF       & Trans. ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
+        (20-100)            & Rot. ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                     & ANEES             & 10.18     & 12.03     & 16.76             \\
                     & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
-        SWF (25)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
-                    & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
+        SWF         & Trans. ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
+         (25)       & Rot. ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                     & ANEES             & 2280      & 2093      & 2013              \\
@@
 \subsection{KITTI Dataset}
+In addition to the ``Starry Night'' datasets, we also compared the performance of the MSCKF and SWF over several traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totalling $530$ m of urban driving.
+To accommodate our alternative state parametrization, we used the pre-integrated linear velocity measurements provided in the dataset instead of the raw linear accelerations from the IMU.
+Figure \ref{fig:kitti} shows the RMSE of pure IMU integration, the MSCKF, and the SWF over each of the four traverses we tested, representative images from which are shown in Figure \ref{fig:composite-kitti}.
+Similarly to the ``Starry Night'' results, both algorithms outperformed pure IMU integration, and the SWF slightly outperformed MSCKF in terms of rotational and translational RMSE.
+As expected, the MSCKF produced estimates that were more consistent than those of the SWF.
+
 \begin{figure*}
     \centering
-    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0001} traverse.] {
+    \subfigure[RMSE on KITTI traverse \texttt{0001}.] {
         \centering
@@
     ~
-    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0036} traverse.] {
+    \subfigure[RMSE on KITTI traverse \texttt{0036}.] {
         \centering
@@
 
-    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0051} traverse.] {
+    \subfigure[RMSE on KITTI traverse \texttt{0051}.] {
         \centering
@@
     ~
-    \subfigure[RMSE on the \texttt{2011\_09\_26\_drive\_0095} traverse.] {
+    \subfigure[RMSE on KITTI traverse \texttt{0095}.] {
         \centering
@@
     }
-    \caption{RMSE over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling 534 m of urban driving.}
+    \caption{Translational and rotational RMSE over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling 534 m of urban driving.}
     \label{fig:kitti}
 \end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.9\textwidth]{figs/kitti/composite-kitti}
+    \caption{Sample frames from the four KITTI traverses shown in Figure \ref{fig:kitti}.}
+    \label{fig:composite-kitti}
+\end{figure*}
+
+\begin{table}
+    \centering
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE) and Average Normalized Estimation Error Squared (ANEES) IMU integration, MSCKF, and SWF on KITTI traverses.}
+    \begin{tabular}{llcccc} 
+                    &                   & \multicolumn{4}{c}{KITTI Traverse} \B      \\ \cline{3-6}
+                    &                   & \texttt{0001} & \texttt{0036} & \texttt{0051} & \texttt{0095}    \T\B    \\ \hline
+        IMU Only    & Trans. ARMSE      & 0.7197    & 0.5131    & 0.7834    & 1.039     \T      \\
+                    & Rot. ARMSE        & 0.0025    & 0.0113    & 0.0034    & 0.0046       \\
+                    & ANEES             & 0.1630    & 0.0092    & 0.1170    & 0.6254         \B      \\ \hline
+        MSCKF       & Trans. ARMSE      & 0.3492    & 0.4401    & 0.7530    & 0.8170  \T      \\
+        (5-Inf)     & Rot. ARMSE        & 0.0111    & 0.0096    & 0.0091    & 0.0345          \\
+                    & ANEES             & 5.103     & 1.826     & 2.031     & 14.98          \B      \\ \hline
+        SWF         & Trans. ARMSE      & 0.3372    & 0.3778    & 0.5832    & 0.7196  \T      \\
+        (10)        & Rot. ARMSE        & 0.0113    & 0.0112    & 0.0111    & 0.0154       \\
+                    & ANEES             & 358.3     & 703.2     & 1124      & 3767    \B      \\ \hline
+    \end{tabular}
+    \label{tab:kitti_results}
+    % \vspace{-0.4cm}
+\end{table}
 
@@
 Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
-We stress that our inertial data was obtained from a high-quality IMU and has been sanitized to account for gravity, biases, and integration error.
+We stress that our inertial data was obtained from high-quality IMUs and has been sanitized to account for gravity, biases, and integration error.
 Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent.
 
-In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, with the SWF significantly outperforming the MSCKF on the datasets we tested.
+In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, with the SWF slightly outperforming the MSCKF on the datasets we tested.
 However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.
@@
 
-Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF has the advantage of being less computationally intensive than the SWF, and of having better consistency properties.
+Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF is less computationally intensive than the SWF and of has better consistency properties.
 The MSCKF may be a better choice of algorithm in applications where computational resources are limited and the operational environment is feature-rich.

### 2015-03-03 Report/project_report.tex
Edited related work, and SWF intro

@@
 The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
-It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features.
-It also takes advantage of \textit{delayed linearization} in attempt to minimize linearization errors.
-The SWF analyzed in this work is similar as a both a recursive and batch approach, but unlike the VSDF it incorporates a motion model and applies batch optimization.
+It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features, taking advantage of \textit{delayed linearization} in attempt to minimize linearization errors.
+The SWF analyzed in this work is similar in that it is both a recursive and batch approach, but unlike the VSDF, it incorporates a motion model from interoceptive data and applies batch optimization.
 The SWF operates in constant time by limiting the optimization problem to a set number of states at all times.
@@
 The MSCKF \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF in the sense that it maintains a variable window of poses and applies batch updates using all observations of each landmark.
-It has recently been used to achieve accurate, online position tracking onboard a smartphone \cite{Li:2013:ICRA}.
+It is also notable for its computationally efficiency, achieving accurate, real-time position tracking onboard a consumer smartphone \cite{Li:2013:ICRA}.
 
@@
 \section{Sliding Window Filter} \label{sec:SWF}
-In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and feature positions.
-The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) optimization over a sliding window of $K$ poses $\Vec x_{k\in[1,K]}$ observing $M$ features $f_{j\in[1,M]}$:
+The aim of the Sliding Window Filter (SWF) is to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and observed landmarks.
+The optimization problem in the SWF is typically solved as a non-linear least squares problem using Gauss-Newton (GN) or Levenberg-Marquardt (LM) optimization over a sliding window of $K$ poses $\Vec x_{k\in[1,K]}$ observing $M$ features $f_{j\in[1,M]}$:
 \begin{equation}
@@
 \begin{equation}
-(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}})
+\label{eqn:gnOpt}
+(\Vec H^T \Vec T ^{-1} \Vec H ) \delta \Vec{x}^*  = - \Vec{H}^T\Vec{T}^{-1}\Vec{e}(\bar{\Vec{x}}).
 \end{equation}
@@
 \begin{align}
-\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T
+\Vec G_{\Vec x, k} &= \bbm \Vec G_{\Vec x, k}^1 & \hdots & \Vec G_{\Vec x, k}^M \ebm ^T,
 \\[0.5em]
 \Vec G_{\Vec{x}, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}}
-    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec p_{G,k}^{IG} ) )^\times  \ebm ^T
+    \bbm -\Vec C_{CG} \Vec C_{IG,k} \\[0.5em] \Vec C_{CG} (\Vec C_{IG,k} (\Vec{p}_G^{f_j G} - \Vec p_{G,k}^{IG} ) )^\times  \ebm ^T,
 \\[0.5em]
-\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j\in[1,M]$}} \\[0.5em]
-\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k}
+\Vec G_{\Vec f, k} &= \underbrace{\bbm \Vec G_{\Vec f, k}^1 & & \\ & & \Vec G_{\Vec f, k}^2 & \\ & & \vdots &  \\ & & & \Vec G_{\Vec f, k}^M \ebm}_{\text{$j$th block column position given by feature ID $j\in[1,M]$}}, \\[0.5em]
+\Vec G_{\Vec f, k}^j &= \pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} \Vec C_{CI} \Vec C_{IG,k},
 \end{align}
@@
 \begin{equation}
-\Vec{T} := \text{diag} \left\{ \Vec{T}_1,\hdots,\Vec{T}_K \right\}
+\Vec{T} := \text{diag} \left\{ \Vec{T}_1,\hdots,\Vec{T}_K \right\},
 \end{equation}
@@
 and $\Vec R_k$, $\Vec Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise. 
-The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).
-
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
@@
 \subsection{MSCKF State Parametrization}
-We evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
+We evaluated the MSCKF with datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
 In order to accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.
@@
 
-We can also define the MSCKF ``error state'' at time $k$:
+We can also define the MSCKF  \textit{error state} at time $k$:
 \begin{align*}
@@
 \end{align}
-The linear least-squares system then becomes
+The least-squares system then becomes
 \begin{align}

### 2015-03-03 Report/project_report.tex
Removed rotational rmse from KITTI

@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0001}
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0001}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0036}
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0036}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0051}
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0051}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_2011_09_26_drive_0095}
+        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0095}
     }

### 2015-03-03 Matlab/KITTI Trials/plotKITTI.m

@@ -1,8 +1,8 @@
 %% Load data
 clear; close all; clc;
 
-% fileName = '2011_09_26_drive_0001';
-fileName = '2011_09_26_drive_0036';
+ fileName = '2011_09_26_drive_0001';
+%fileName = '2011_09_26_drive_0036';
 % fileName = '2011_09_26_drive_0051';
 % fileName = '2011_09_26_drive_0095';
 
@@ -42,13 +42,13 @@ imu_anees = ANEES(msckfData.imu_trans_err, msckfData.imu_rot_err, imu_err_sigma)
 figure(1); clf;
 fontSize = 14;
 lineWidth = 2;
-pos = [200,200,640,400];
+pos = [200,200,640,200];
 xLim = [dists(1), dists(end)];
 
 set(gcf,'Position',pos);
 
 % Translational RMSE
-subplot(2,1,1);
+%subplot(2,1,1);
 plot(dists,imu_trans_rmse,'-k','LineWidth',lineWidth); hold on;
 plot(dists,swf_trans_rmse,'-r','LineWidth',lineWidth); 
 plot(dists,msckf_trans_rmse,'-b','LineWidth',lineWidth);
@@ -61,18 +61,18 @@ set(gca,'FontSize',fontSize);
 set(gcf,'Position',pos);
 
 % Rotational RMSE
-subplot(2,1,2);
-plot(dists,imu_rot_rmse,'-k','LineWidth',lineWidth); hold on;
-plot(dists,swf_rot_rmse,'-r','LineWidth',lineWidth); 
-plot(dists,msckf_rot_rmse,'-b','LineWidth',lineWidth);
-xlim(xLim);
-xlabel('Distance Travelled (m)'); ylabel('Rot. RMSE (Axis-Angle)');
-legend('IMU Only','SWF','MSCKF','Location','NorthWest');
-grid minor; box on;
-set(gca,'FontSize',fontSize);
+% subplot(2,1,2);
+% plot(dists,imu_rot_rmse,'-k','LineWidth',lineWidth); hold on;
+% plot(dists,swf_rot_rmse,'-r','LineWidth',lineWidth); 
+% plot(dists,msckf_rot_rmse,'-b','LineWidth',lineWidth);
+% xlim(xLim);
+% xlabel('Distance Travelled (m)'); ylabel('Rot. RMSE (Axis-Angle)');
+% legend('IMU Only','SWF','MSCKF','Location','NorthWest');
[…]

### 2015-03-03 Report/project_report.tex
Updated conclusion

@@
 
-We implemented and tested both algorithms in MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
+We implemented and tested both algorithms in MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3 GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
 For both algorithms, we bootstrapped the filter window with one ground truth pose.
@@
 
-\subsection{Starry Night Dataset}
+\subsection{``Starry Night'' Dataset}
 The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
@@
     \begin{tablenotes}
-        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
+        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3 GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
     \end{tablenotes}
@@
     \centering
-    \subfigure[RMSE on KITTI traverse \texttt{0001}.] {
+    \subfigure[Translational RMSE on KITTI traverse \texttt{0001}.] {
         \centering
@@
     ~
-    \subfigure[RMSE on KITTI traverse \texttt{0036}.] {
+    \subfigure[Translational RMSE on KITTI traverse \texttt{0036}.] {
         \centering
@@
 
-    \subfigure[RMSE on KITTI traverse \texttt{0051}.] {
+    \subfigure[Translational RMSE on KITTI traverse \texttt{0051}.] {
         \centering
@@
     ~
-    \subfigure[RMSE on KITTI traverse \texttt{0095}.] {
+    \subfigure[Translational RMSE on KITTI traverse \texttt{0095}.] {
         \centering
@@
     }
-    \caption{Translational and rotational RMSE over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling 534 m of urban driving.}
+    \caption{Translational RMSE over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling 534 m of urban driving.}
     \label{fig:kitti}
@@
     \centering
-    \caption{Comparison of Average Root Mean Squared Error (ARMSE) and Average Normalized Estimation Error Squared (ANEES) IMU integration, MSCKF, and SWF on KITTI traverses.}
+    \caption{Comparison of Average Root Mean Squared Error (ARMSE) and Average Normalized Estimation Error Squared (ANEES) of IMU integration, MSCKF, and SWF on KITTI traverses.}
     \begin{tabular}{llcccc} 
@@
 \section{Conclusions} \label{sec:conclusions}
-Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
-We stress that our inertial data was obtained from high-quality IMUs and has been sanitized to account for gravity, biases, and integration error.
-Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent.
-
-In environments with more features, both algorithms provided substantial gains in accuracy over pure IMU integration, with the SWF slightly outperforming the MSCKF on the datasets we tested.
-However, the accuracy of the MSCKF improves as more features are added, while the SWF is less sensitive to feature quantity.
+Our experimental results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
+On the datasets we tested, the SWF slightly outperformed the MSCKF, however, the accuracy of the MSCKF improved as more features were added while the SWF was less sensitive to feature quantity.
 It is unclear from these results which algorithm would perform best as the number of features grows very large.
 
-Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF is less computationally intensive than the SWF and of has better consistency properties.
+In relatively featureless environments, neither filter performed substantially better than pure IMU integration.
+We stress that our inertial data was obtained from high-quality IMUs and has been sanitized to account for gravity, biases, and integration error.
+Using a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent in these cases.
+
+Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF is less computationally intensive than the SWF and has better consistency properties in its most basic form.
 The MSCKF may be a better choice of algorithm in applications where computational resources are limited and the operational environment is feature-rich.

### 2015-03-03 Report/project_report.tex
Couple more conclusion edits

@@
 Our experimental results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
-On the datasets we tested, the SWF slightly outperformed the MSCKF, however, the accuracy of the MSCKF improved as more features were added while the SWF was less sensitive to feature quantity.
-It is unclear from these results which algorithm would perform best as the number of features grows very large.
+On the datasets we tested, the SWF slightly outperformed the MSCKF, however, the MSCKF improved in accuracy with additional features while the SWF was less sensitive to feature quantity.
+It is unclear which algorithm would perform best as the number of features grows very large.
 

### 2015-03-03 Report/project_report.tex
Added KITTI surf,klt,msac

--- 2015-03-03 Report/project_report.tex
@@
 \subsection{Feature Position Estimation}
+\label{sec:featPosinvDepth}
 When a tracked feature $f_j$ is selected for a state update, the MSCKF estimates its position $\hat{\Vec{p}}_G^{f_j G}$ using an inverse-depth least-squares Gauss-Newton optimization. 
@@
 To accommodate our alternative state parametrization, we used the pre-integrated linear velocity measurements provided in the dataset instead of the raw linear accelerations from the IMU. 
+We extracted point features from the left camera in the stereo pair using SURF (Speeded Up Robust Features \cite{Bay:2008:CVIU}) and tracked them temporally using KLT (Kanade-Lucas-Tomasi \cite{Tomasi:91}).
+We employed an MSAC (M-estimator SAmple Consensus \cite{TorrZisserman:1998}) procedure with a 2-point similarity transform to reject outliers.
+To initialize 3D points, we used the same inverse-depth optimization used by MSCKF (see section \ref{sec:featPosinvDepth}).
+
 Figure \ref{fig:kitti} shows the RMSE of pure IMU integration, the MSCKF, and the SWF over each of the four traverses we tested, representative images from which are shown in Figure \ref{fig:composite-kitti}.

### 2015-03-03 Matlab/swf/SlidingWindow_ReRun.m

@@ -4,8 +4,7 @@
 clc
 clear
 close all
-load('../KITTI Trials/SWF_2011_09_26_drive_0095_sync_KLT.mat');
-
+load('../KITTI Trials/SWF_2011_09_26_drive_0036_sync_KLT.mat');
 k1 = kStart;
 k2 = k1+kappa;
 K = k2 - k1;  %There are K + 1 total states, since x0 is the k1th state

### 2015-03-03 Report/project_report.tex
Added acknowledgement

@@
 
-The shortcomings of the EKF motivate the use of filters that operate on a subset of the problem.
+T he shortcomings of the EKF motivate the use of filters that operate on a subset of the problem.
 Algorithms operating in constant time with respect to map size are particularly desirable.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section*{ACKNOWLEDGMENT}
+\section*{ACKNOWLEDGMENT}
+We would like to thank Professor Timothy D. Barfoot for the use of the \textit{Starry Night Dataset} that was prepared for the graduate state estimation course, AER1513, at the University of Toronto.
 %\vspace{8pt}

### 2015-03-03 Report/project_report.tex
Added leutenegger reference and comparison

--- 2015-03-03 Report/project_report.tex
@@
 Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
+
 In this work we characterize, compare, and contrast the performance of two modern solutions to the visual-inertial SLAM problem, the Sliding Window Filter (SWF) and the Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA}, on data from a moving hand-held sensor rig, as well as several traverses from the KITTI dataset \cite{Geiger:2013:IJRR}.
+The most similar work to ours is that of Leutenegger et al.
+\cite{Leutenegger:2015:IJRR}, which compares the accuracy of the MSCKF to a keyframe-based SWF on datasets consisting of relatively planar motion through urban and indoor environments.
+In contrast to \cite{Leutenegger:2015:IJRR}, our SWF optimizes over a constant number of timesteps rather than keyframes.
+We also conduct a more extensive characterization of the sensitivity of the MSCKF to certain parameters and compare both algorithms using data from a hand-held sensor rig that mimics the more arbitrary motion of a micro aerial vehicle (MAV).
 
@@
     \includegraphics[width=0.45\textwidth]{figs/sensorhead}
-    \caption{The sensor head used in our experiments.
-An IMU measures translational and rotational velocities, while a stereo camera measures the positions of point features.
-In our experiments, we artificially blinded the stereo camera by using measurements from the left camera only.}
+    \caption{The hand-held sensor head used in our experiments with the ``Starry Night'' dataset.
+An IMU reports translational and rotational velocities, while a stereo camera observes point features.
+We artificially blinded the stereo camera by using measurements from the left camera only.}
     \label{fig:sensorhead}
@@
 
-\section{Related Work} \label{sec:relatedwork}
+\section{Background} \label{sec:relatedwork}
 
@@
 The MSCKF \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF in the sense that it maintains a variable window of poses and applies batch updates using all observations of each landmark.
-It is also notable for its computationally efficiency, achieving accurate, real-time position tracking onboard a consumer smartphone \cite{Li:2013:ICRA}.
+It is also notable for its computationally efficiency, achieving accurate, real-time position tracking onboard a smartphone \cite{Li:2013:ICRA}.
 
@@
     \label{fig:comp_3}
-    \vspace{-0.4cm}
+    % \vspace{-0.4cm}
 \end{figure}

### 2015-03-03 Report/project_report.tex
Minor stuff

--- 2015-03-03 Report/project_report.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section*{ACKNOWLEDGMENT}
+\section*{ACKNOWLEDGMENT}
+\noindent
+\small{Many thanks to Prof. Timothy D. Barfoot for the use of the ``Starry Night'' dataset that was prepared for the state estimation course, AER1513, taught at the University of Toronto.}
 %\vspace{8pt}

### 2015-03-03 Report/project_report.tex
Acknowledge

@@
 \noindent
-\small{Many thanks to Prof. Timothy D. Barfoot for the use of the ``Starry Night'' dataset that was prepared for the state estimation course, AER1513, taught at the University of Toronto.}
+\small{We would like to thank Prof. Timothy D. Barfoot for the use of the ``Starry Night'' dataset prepared for the graduate state estimation course, AER1513, taught at the University of Toronto.}
 %\vspace{8pt}

### 2015-03-04 Report/project_report.tex
Final version unless anyone has edits

@@
     \caption{The hand-held sensor head used in our experiments with the ``Starry Night'' dataset.
-An IMU reports translational and rotational velocities, while a stereo camera observes point features.
+The IMU reports translational and rotational velocities, while the stereo camera observes point features.
 We artificially blinded the stereo camera by using measurements from the left camera only.}
@@
 Visual-inertial navigation systems (VINS) have been applied broadly in robotics \cite{Mourikis:2007:ICRA, Sibley:2010:JFR, Huster:2002:OC, Kottas:2013:IROS, Kim:2007:RSS, Li:2013coa,  You:2001:VR}, and there is a considerable body of work covering a wide range of estimation algorithms for the camera-IMU sensor pair.
-These techniques are often characterized as either \textit{loosely coupled}, when image and IMU measurements are processed individually before being fused into a single estimate, or \textit{tightly coupled}, when all information is processed together.
+These techniques are often characterized as either \textit{loosely coupled} or \textit{tightly coupled}.
+In loosely coupled systems, image and IMU measurements are processed individually before being fused into a single estimate, while tightly coupled systems process all information together.  
 The decoupling of inertial and visual measurements in loosely coupled systems leads to inexpensive computations, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
@@
 Indeed, the computational complexity of such algorithm scales far too poorly with the number of features to be applied naively to this problem.
-Additionally, EKF-based VINS are fundamentally inconsistent, that is, the state uncertainties are underestimated.
-Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, severely reducing the reliability of such estimators.
 
-Another limitation of EKF-SLAM is that it is \textit{forgetful}.
-Because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.
-By locking in past poses, EKF-SLAM sub-optimally estimates both vehicle motion and feature positions.
+Furthermore, EKF-based VINS are fundamentally inconsistent, that is, the state uncertainties are underestimated.
+Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, which reduces the reliability of such estimators.
 
-T he shortcomings of the EKF motivate the use of filters that operate on a subset of the problem.
-Algorithms operating in constant time with respect to map size are particularly desirable.
-The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
+Finally, EKF-SLAM is \textit{forgetful}:
+because the filter state includes only the most recent vehicle pose, a given update step can never modify past poses even if later feature measurements ought to constrain them.
+By locking in past poses, EKF-SLAM condemns itself to sub-optimally estimating both vehicle motion and feature positions.
+
+The shortcomings of the EKF motivate the use of filters that operate on a subset of the problem. Algorithms operating in constant time with respect to map size are particularly desirable.
+% The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
 It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features, taking advantage of \textit{delayed linearization} in attempt to minimize linearization errors.
-The SWF analyzed in this work is similar in that it is both a recursive and batch approach, but unlike the VSDF, it incorporates a motion model from interoceptive data and applies batch optimization.
-The SWF operates in constant time by limiting the optimization problem to a set number of states at all times.
-Batch optimization over a window of states is suboptimal but still leads to exceptional accuracy for its cost, making it useful for delicate, large scale operations such as planetary landing\cite{Sibley:2010:JFR}.
+The SWF analyzed in this work is a modern example of such algorithms, operating in constant time by limiting the optimization problem to a set number of states.
+It applies batch optimization over a set window of states, which is suboptimal but still leads to exceptional accuracy for its cost.
+This makes the SWF useful for delicate, large scale operations such as planetary landing\cite{Sibley:2010:JFR}. 
 The MSCKF \cite{Mourikis:2006:TechReport,Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF in the sense that it maintains a variable window of poses and applies batch updates using all observations of each landmark.
 It is also notable for its computationally efficiency, achieving accurate, real-time position tracking onboard a smartphone \cite{Li:2013:ICRA}.
+In Sections \ref{sec:SWF} and \ref{sec:MSCKF}, we discuss both the SWF and MSCKF in detail.
 
@@
 \end{equation}
-In our notation, $\Vec p^{BA}_A$ is the vector from the origin of frame $\mathcal{F}_A$ to the origin of frame $\mathcal{F}_B$ expressed in $\mathcal{F}_A$, and $\Vec C_{BA}$ is the rotation matrix from $\mathcal{F}_A$ to $\mathcal{F}_B$.
+In our notation, $\Vec p^{BA}_A$ is the vector from the origin of frame $\mathcal{F}_A$ to the origin of frame $\mathcal{F}_B$ expressed in $\mathcal{F}_A$, $\Vec C_{BA}$ is the rotation matrix from $\mathcal{F}_A$ to $\mathcal{F}_B$, and $\Vec{1}$ is the identity matrix.
 
@@
 An important advantage of the SWF is that its computational cost depends on the number of features in the current window rather than the number of features in the entire map.
-By varying the extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
+By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.
 
@@
 \subsection{MSCKF State Parametrization}
-We evaluated the MSCKF with datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
-In order to accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.
+We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
+To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.
 
@@
                             \left(\hat{\Vec{C}}_{I G, k}^T \Vec{p}_{I,k}^{CI}\right)^\times & \Vec{0}_{3\times6} & \Vec{1}_3 & \Vec{0}_{3\times6N}
-                \ebm .
+                \ebm
 \end{align}
+with $\Vec 1_m$ denoting the $m$-dimensional identity matrix and $\Vec 0_{m\times n}$ an $m\times n$ matrix of zeros.
 
@@
         & \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \notag \\
-    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} 
-        & \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
+    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T \hat{\Vec{v}}
+        & \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} 
 \end{align}
@@
 with the initial condition $\hat{\Vec{P}}_{II}\left(t_k\right) = \hat{\Vec{P}}_{II,k}$. 
+To improve numerical stability, we found it useful to enforce the positive semi-definiteness of $\hat{\Vec{P}}^-_{II,k+1}$ by making the replacements
+\begin{align*}
+    P_{ii} \leftarrow |P_{ii}| && \text{and} && P_{ij}, P_{ji} \leftarrow \dfrac{1}{2}\left(P_{ij} + P_{ji}\right), i\neq j,
+\end{align*}
+where $P_{ij}$ denotes the $i^\text{th}$ row and $j^\text{th}$ column of $\hat{\Vec{P}}^-_{II,k+1}$.
 
@@
 \section{Experiments} \label{sec:experiments}
-We compared the performance of the MSCKF and SWF on two intervals in the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS), as well as 530 m of urban driving from the KITTI Dataset \cite{Geiger:2013:IJRR}.
+We compared the performance of the MSCKF and SWF on two intervals in the ``Starry Night'' dataset recorded at the University of Toronto Institute for Aerospace Studies (UTIAS), as well as 534 m of urban driving from the KITTI Dataset \cite{Geiger:2013:IJRR}.
 Both datasets include data streams from a stereo camera and IMU.
@@
 We implemented and tested both algorithms in MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3 GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
-For both algorithms, we bootstrapped the filter window with one ground truth pose.
-For the SWF, we initialized feature positions using two-view triangulation.
+We initialized both filters with the first ground truth pose, and obtained initial estimates of 3D feature positions using the inverse-depth batch optimization of Section \ref{sec:featPosinvDepth}.
 
@@
 % \end{figure}
-
-\subsection{``Starry Night'' Dataset}
-The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
-This dataset is well-suited to evaluating SLAM algorithms since ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
-We conducted three experiments on this dataset to compare the effect of feature visibility and window size on each algorithm.
-We discuss each of these in turn.
 
@@
     \caption{Comparison of MSCKF, SWF, and IMU integration for multiple parameter settings on two intervals of the ``Starry Night'' dataset.
-The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refers to the number of states in the sliding window.}
+The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refer to the number of states in the sliding window.}
 \end{figure*}
+
+\subsection{``Starry Night'' Dataset}
+The ``Starry Night'' dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 features while moving along an arbitrary 3D path.
+This dataset is well-suited to evaluating SLAM algorithms since ground truth from a Vicon motion capture system is available for both the sensor head motion and the feature positions.
+We conducted three experiments on this dataset to compare the effect of feature visibility and window size on each algorithm.
+We discuss each of these in turn.
 
@@
 Figure \ref{fig:comp_2} shows translational and rotational RMSE for pure IMU integration and the same MSCKF and SWF parameter settings as in Experiment 1.
-Compared to Interval 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy in the $y$-direction.
+Compared to Experiment 1, the difference in performance between the MSCKF and SWF is less pronounced, with the exception of the SWF's clearly superior translational accuracy.
 Neither filter performs substantially better than pure IMU integration on this interval, again likely due to the overall low number of feature tracks in this dataset.
@@
 
-\subsubsection{Effect of feature density}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}
-
-    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$.
-The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
-    \label{fig:comp_3}
-    % \vspace{-0.4cm}
-\end{figure}
-
+\subsubsection{Effect of feature density in synthetic maps}
 In order to investigate the effect of feature density on the performance of the MSCKF and SWF, we modified the ``Starry Night'' dataset by creating synthetic feature distributions with larger spatial extents and more features than the original dataset.
@@
 % Figure \ref{fig:trajectory_groundtruth_100lessnoisy} shows the feature distribution for the 100-feature synthetic dataset.
-By generating longer feature tracks and increasing the number of features visible at each time step, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.
+By generating longer feature tracks and increasing the number of visible features, these synthetic datasets allowed for a clearer comparison between the MSCKF and SWF.
 
@@
 Table \ref{tab:rmse_nees} summarizes each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
-As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation and translation.
+As noted above, the SWF outperforms the MSCKF in terms of average RMSE on both rotation and translation.
 It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
-The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
-The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.
+The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window.
+The authors of \cite{Sibley:2010:JFR} describe how this can be remedied by marginalizing old poses, but we did not implement this technique because our intent was to directly compare the two algorithms in their simplest forms.
+% The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.
 
@@
 The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Noisy}
+
+    \caption{Comparison of Root Mean Squared Error (RMSE) of the MSCKF, SWF, and pure IMU integration for three synthetic maps generated using the IMU data in the interval $(1215,1715)$.
+The numbers next to ``MSCKF'' and ``SWF'' refer to the number of features in the dataset (40, 60 or 100).}
+    \label{fig:comp_3}
+    % \vspace{-0.4cm}
+\end{figure}
 
@@
 \subsection{KITTI Dataset}
-In addition to the ``Starry Night'' datasets, we also compared the performance of the MSCKF and SWF over several traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totalling $530$ m of urban driving.
+In addition to the ``Starry Night'' datasets, we also compared the performance of the MSCKF and SWF over four traverses from the KITTI dataset \cite{Geiger:2013:IJRR} totaling $534$ m of urban driving.
+We selected these particular traverses because they contained few moving objects, which we found to be common failure cases for both filters.
 To accommodate our alternative state parametrization, we used the pre-integrated linear velocity measurements provided in the dataset instead of the raw linear accelerations from the IMU. 
-We extracted point features from the left camera in the stereo pair using SURF (Speeded Up Robust Features \cite{Bay:2008:CVIU}) and tracked them temporally using KLT (Kanade-Lucas-Tomasi \cite{Tomasi:91}).
-We employed an MSAC (M-estimator SAmple Consensus \cite{TorrZisserman:1998}) procedure with a 2-point similarity transform to reject outliers.
-To initialize 3D points, we used the same inverse-depth optimization used by MSCKF (see section \ref{sec:featPosinvDepth}).
+We extracted point features from the left camera in the stereo pair using Speeded Up Robust Features (SURF) \cite{Bay:2008:CVIU} and tracked them temporally using Kanade-Lucas-Tomasi (KLT) tracking \cite{Tomasi:91}.
+We rejected outliers using an M-estimator Sample Consensus (MSAC) \cite{TorrZisserman:1998} procedure with a 2-point similarity transform.
+% Like the ``Starry Night'' datasets, we initialized 3D points in both filters using the inverse-depth batch optimization of Section \ref{sec:featPosinvDepth}.
 
 Figure \ref{fig:kitti} shows the RMSE of pure IMU integration, the MSCKF, and the SWF over each of the four traverses we tested, representative images from which are shown in Figure \ref{fig:composite-kitti}.
-Similarly to the ``Starry Night'' results, both algorithms outperformed pure IMU integration, and the SWF slightly outperformed MSCKF in terms of rotational and translational RMSE.
-As expected, the MSCKF produced estimates that were more consistent than those of the SWF.
+Table \ref{tab:kitti_results} summarizes these results and reports NEES values for each traverse.
+Similarly to the ``Starry Night'' results, both algorithms outperformed pure IMU integration, and the SWF slightly outperformed the MSCKF in terms of translational RMSE.
+We did not consider rotational RMSE for the KITTI traverses because ground truth was obtained from GPS and did not provide reliable estimates of the entire 6DOF vehicle pose.
+As expected, the reported ANEES values show that the MSCKF produced estimates that were more consistent than those of the SWF.
 
@@
     \label{fig:composite-kitti}
+    % \vspace{-0.4cm}
 \end{figure*}
@@
         IMU Only    & Trans. ARMSE      & 0.7197    & 0.5131    & 0.7834    & 1.039     \T      \\
-                    & Rot. ARMSE        & 0.0025    & 0.0113    & 0.0034    & 0.0046       \\
+                    % & Rot. ARMSE        & 0.0025    & 0.0113    & 0.0034    & 0.0046       \\
                     & ANEES             & 0.1630    & 0.0092    & 0.1170    & 0.6254         \B      \\ \hline
         MSCKF       & Trans. ARMSE      & 0.3492    & 0.4401    & 0.7530    & 0.8170  \T      \\
-        (5-Inf)     & Rot. ARMSE        & 0.0111    & 0.0096    & 0.0091    & 0.0345          \\
-                    & ANEES             & 5.103     & 1.826     & 2.031     & 14.98          \B      \\ \hline
+        % (5-Inf)     & Rot. ARMSE        & 0.0111    & 0.0096    & 0.0091    & 0.0345          \\
+        (5-Inf)     & ANEES             & 5.103     & 1.826     & 2.031     & 14.98          \B      \\ \hline
         SWF         & Trans. ARMSE      & 0.3372    & 0.3778    & 0.5832    & 0.7196  \T      \\
-        (10)        & Rot. ARMSE        & 0.0113    & 0.0112    & 0.0111    & 0.0154       \\
-                    & ANEES             & 358.3     & 703.2     & 1124      & 3767    \B      \\ \hline
+        % (10)        & Rot. ARMSE        & 0.0113    & 0.0112    & 0.0111    & 0.0154       \\
+        (10)        & ANEES             & 358.3     & 703.2     & 1124      & 3767    \B      \\ \hline
     \end{tabular}
@@
 \section{Conclusions} \label{sec:conclusions}
-Our experimental results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
-On the datasets we tested, the SWF slightly outperformed the MSCKF, however, the MSCKF improved in accuracy with additional features while the SWF was less sensitive to feature quantity.
-It is unclear which algorithm would perform best as the number of features grows very large.
+Our results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
+On the datasets we tested, the SWF slightly outperformed the MSCKF, but the MSCKF improved in accuracy with additional features while the SWF was less sensitive to feature quantity.
+It is unclear from these results which algorithm would perform best as the number of features grows very large.
 
@@
 Although the SWF appears to produce more accurate pose estimates than the MSCKF in many cases, the MSCKF is less computationally intensive than the SWF and has better consistency properties in its most basic form.
-The MSCKF may be a better choice of algorithm in applications where computational resources are limited and the operational environment is feature-rich.
+However, in our experiments we found the MSCKF to be much more sensitive to tuning parameters than the SWF, sometimes diverging wildly with small changes in parameters.
+We conclude that the MSCKF may be a better choice of algorithm when computational resources are limited and the operational environment is feature-rich, but that the SWF may be preferable in situations where robustness is paramount.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{ACKNOWLEDGMENT}
+\section*{Acknowledgment}
 \noindent

### 2015-03-04 Report/project_report.tex
Added note about tuning parameters to abstract
Fixed missing x label on KITTI plots

@@
 In particular, we compare the accuracy and consistency of the two filters, and analyze the effect of feature track length and feature density on the performance of each filter.
-In general, our results show the SWF to be more accurate than the MSCKF.
+In general, our results show the SWF to be more accurate and less sensitive to tuning parameters than the MSCKF.
 However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0001}
+        \includegraphics[width=0.47\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0001}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0036}
+        \includegraphics[width=0.47\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0036}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0051}
+        \includegraphics[width=0.47\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0051}
     }
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0095}
+        \includegraphics[width=0.47\textwidth]{figs/kitti/RMSE_trans_2011_09_26_drive_0095}
     }

### 2015-03-04 Matlab/KITTI Trials/plotKITTI.m

@@ -1,10 +1,10 @@
 %% Load data
 clear; close all; clc;
 
-%fileName = '2011_09_26_drive_0001';
-fileName = '2011_09_26_drive_0036';
+% fileName = '2011_09_26_drive_0001';
+% fileName = '2011_09_26_drive_0036';
 % fileName = '2011_09_26_drive_0051';
-% fileName = '2011_09_26_drive_0095';
+fileName = '2011_09_26_drive_0095';
 
 
 fileSuffix = '_sync_KLT';
@@ -56,6 +56,7 @@ plot(dists,msckf_trans_rmse,'-b','LineWidth',lineWidth);
 xlim(xLim);
 title(fileName,'Interpreter','none');
 ylabel('Trans. RMSE (m)');
+xlabel('Distance Travelled (m)');
 legend('IMU Only','SWF','MSCKF','Location','NorthWest');
 grid minor; box on;
 set(gca,'FontSize',fontSize);

### 2015-03-05 Report/project_report.tex
Added citation for loosely coupled

@@
 In loosely coupled systems, image and IMU measurements are processed individually before being fused into a single estimate, while tightly coupled systems process all information together.  
-The decoupling of inertial and visual measurements in loosely coupled systems leads to inexpensive computations, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
+The decoupling of inertial and visual measurements in loosely coupled systems limits computational complexity \cite{Leutenegger:2015:IJRR}, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
 In this work, we consider tightly coupled algorithms, which are preferable for accurate and consistent visual-inertial navigation.

### 2015-03-05 Report/project_report.tex
Changed covariance propagation equations to what we actually do in the code

--- 2015-03-04 Report/project_report.tex
@@
 These techniques are often characterized as either \textit{loosely coupled} or \textit{tightly coupled}.
-In loosely coupled systems, image and IMU measurements are processed individually before being fused into a single estimate, while tightly coupled systems process all information together.  
+In loosely coupled systems, image and IMU measurements are processed separately before being fused into a single estimate, while tightly coupled systems process all information together.  
 The decoupling of inertial and visual measurements in loosely coupled systems leads to inexpensive computations, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
@@
 For both tightly coupled and loosely coupled VINS, a popular choice of estimator is the Extended Kalman filter (EKF) or one of its variants \cite{Huster:2002:OC, You:2001:VR, Kottas:2013:RSS, Kim:2007:RSS, Mourikis:2007:ICRA}, though methods also exist that employ unscented Kalman filters \cite{Ebcin:2007:ION}, particle filters \cite{Fox:1999:JAIR, Pupilli:2005:BMVC}, or batch optimization methods\cite{Triggs:2000va,Strelow:2004:IJR}.
+
 EKF-SLAM is an efficient recursive algorithm for small-scale, online tasks, but maintaining the entire map as part of the state is a weakness for navigation over long distances.
-Indeed, the computational complexity of such algorithm scales far too poorly with the number of features to be applied naively to this problem.
+Indeed, the computational complexity of EKF-SLAM scales far too poorly with the number of features to be applied naively to this problem.
 
-Furthermore, EKF-based VINS are fundamentally inconsistent, that is, the state uncertainties are underestimated.
+Furthermore, EKF-based VINS are inconsistent, that is, the state uncertainties are underestimated.
 Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, which reduces the reliability of such estimators.
@@
 It applied a Kalman filter over a sliding window of states to achieve linear complexity with the number of features, taking advantage of \textit{delayed linearization} in attempt to minimize linearization errors.
-The SWF analyzed in this work is a modern example of such algorithms, operating in constant time by limiting the optimization problem to a set number of states.
-It applies batch optimization over a set window of states, which is suboptimal but still leads to exceptional accuracy for its cost.
+The SWF analyzed in this work is a modern example of such algorithms, operating in constant time by performing a batch optimization over a set number of states, which is suboptimal but still leads to exceptional accuracy for its cost.
 This makes the SWF useful for delicate, large scale operations such as planetary landing\cite{Sibley:2010:JFR}. 
@@
 \end{equation}
-The matrix $\Vec{H}$ is given by (showing non-zero blocks only)
+Showing non-zero blocks only, the matrix $\Vec{H}$ is given by
 \begin{equation}
@@
 However, the hard cut-off of the SWF may result in only a subset of measurements of each feature contributing to the optimization.
-As a result, the filter may not maximally constrain some poses, and hence localization may be less accurate than we could expect from the full batch solution.
+As a result, the filter may not maximally constrain some poses, and hence localization may be less accurate than one would expect from the full batch solution.
 
@@
 \end{align}
-with $\Vec 1_m$ denoting the $m$-dimensional identity matrix and $\Vec 0_{m\times n}$ an $m\times n$ matrix of zeros.
+with $\Vec 1_m$ denoting the $m$-dimensional identity matrix and $\Vec 0_{n\times p}$ an $n\times p$ matrix of zeros.
 
@@
     \hat{\Vec{P}}^-_{CC,k+1} &= \hat{\Vec{P}}_{CC,k} \\
-    \hat{\Vec{P}}^-_{IC,k+1} &= \boldsymbol{\Phi}\left(t_k + T, t_k \right) \hat{\Vec{P}}_{IC,k}
+    \hat{\Vec{P}}^-_{IC,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{IC,k}
 \end{align}
-where $T$ is the IMU sampling period, and the state transition matrix $\boldsymbol{\Phi}\left(t_k + T, t_k \right)$ is obtained by integrating
+where $\Delta T$ is the IMU sampling period, and the state transition matrix $\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right)$ is obtained from
 \begin{align} \label{eq:statetransition}
-    \dot{\boldsymbol{\Phi}}\left(t_k + \tau, t_k \right) &= \Vec{F} \boldsymbol{\Phi}\left(t_k + \tau, t_k \right), \tau\in\left[0,T\right]
+	\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) &= \Vec{1}_{12} + \Vec{F} \Delta T.
 \end{align}
-with the initial condition $\boldsymbol{\Phi}\left(t_k, t_k \right) = \Vec{1}_{12}$.
 
-We obtain the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ by integrating
-\begin{align} \label{eq:imucovar}
-    \dot{\hat{\Vec{P}}}_{II}\left(t_k  + \tau \right) &= \Vec{F} \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) 
-                 + \hat{\Vec{P}}_{II}\left(t_k  + \tau \right) \Vec{F}^T \notag \\ 
-                & + \Vec{G} \Vec{Q}_I \Vec{G}^T, \tau\in\left[0,T\right]
+We obtain the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ according to
+\begin{align}
+	\hat{\Vec{P}}^-_{II,k+1} &= \boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) \hat{\Vec{P}}_{II,k} \boldsymbol{\Phi}^T\left(t_k + \Delta T, t_k \right) \notag \\ &+ \Vec{G} \Vec{Q}_I \Vec{G}^T \Delta T.
 \end{align}
-with the initial condition $\hat{\Vec{P}}_{II}\left(t_k\right) = \hat{\Vec{P}}_{II,k}$. 
 To improve numerical stability, we found it useful to enforce the positive semi-definiteness of $\hat{\Vec{P}}^-_{II,k+1}$ by making the replacements

### 2015-03-05 Report/project_report.tex

@@
 These techniques are often characterized as either \textit{loosely coupled} or \textit{tightly coupled}.
-In loosely coupled systems, image and IMU measurements are processed separately before being fused into a single estimate, while tightly coupled systems process all information together.  
-The decoupling of inertial and visual measurements in loosely coupled systems leads to inexpensive computations, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
+In loosely coupled systems, image and IMU measurements are processed individually before being fused into a single estimate, while tightly coupled systems process all information together.  
+The decoupling of inertial and visual measurements in loosely coupled systems limits computational complexity \cite{Leutenegger:2015:IJRR}, but at the cost of information: processing camera and IMU measurements separately makes optimal estimation of biases impossible \cite{Li:2013coa}.
 In this work, we consider tightly coupled algorithms, which are preferable for accurate and consistent visual-inertial navigation.

### 2015-03-05 Report/project_report.tex
Added Leutenegger RSS paper and words about MSCKF

@@
 The key idea of the MSCKF is to maintain a variable window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of features that are visible across the entire window.
-This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds some preset threshold.
+This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
+Except where otherwise noted (i.e., state parametrization and integration method), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
 
@@
 We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
-To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport} and \cite{Mourikis:2007:ICRA}.
+To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
 
@@
 \end{align}
-where $\Delta T$ is the IMU sampling period, and the state transition matrix $\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right)$ is obtained from
+where $\Delta T$ is the IMU sampling period.
+
+The state transition matrix $\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right)$ and the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ are computed according to \cite{Leutenegger:2013:RSS}:
 \begin{align} \label{eq:statetransition}
-	\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) &= \Vec{1}_{12} + \Vec{F} \Delta T.
+	\boldsymbol{\Phi}\left(t_k + \Delta T, t_k \right) &= \Vec{1}_{12} + \Vec{F} \Delta T
 \end{align}
-
-We obtain the predicted IMU state covariance $\hat{\Vec{P}}^-_{II,k+1}$ according to
 \begin{align}
@@
 The numbers next to ``MSCKF'' in the legend refer to the minimum and maximum feature track lengths before an EKF update is triggered, while the numbers next to ``SWF'' refer to the number of states in the sliding window.}
+
+    \vspace{-0.2cm}
 \end{figure*}
@@
     \label{fig:comp_3}
-    % \vspace{-0.4cm}
+    \vspace{-0.2cm}
 \end{figure}
@@
     \label{tab:rmse_nees}
-    % \vspace{-0.4cm}
+    \vspace{-0.2cm}
 \end{table}
@@
     \label{fig:composite-kitti}
-    % \vspace{-0.4cm}
+    \vspace{-0.2cm}
 \end{figure*}
@@
     \label{tab:kitti_results}
-    % \vspace{-0.4cm}
+    \vspace{-0.2cm}
 \end{table}

### 2015-03-05 Report/project_report.tex
Spacing fix

@@
 This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
-Except where otherwise noted (i.e., state parametrization and integration method), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
+Except where otherwise noted (i.e., \ref{sec:state_parametrization} and \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
 
-\subsection{MSCKF State Parametrization}
+\subsection{MSCKF State Parametrization} \label{sec:state_parametrization}
 We evaluated the MSCKF on datasets in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations (see Section \ref{sec:experiments}).
-To accommodate this alternative sensor configuration, the mathematical framework described in this section differs slightly from \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
-
-In our implementation, we parametrize the IMU state at time $k$ as the 13-dimensional vector
+To accommodate this alternative sensor configuration, we parametrize the IMU state at time $k$ as the 13-dimensional vector
 \begin{align}
@@
 
-\subsection{IMU State Estimate Propagation}
+\subsection{IMU State Estimate Propagation} \label{sec:imu_prop}
 The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by a continuous-time motion model:

### 2015-03-05 Report/project_report.tex
Another spacing fix

@@
 
-For both tightly coupled and loosely coupled VINS, a popular choice of estimator is the Extended Kalman filter (EKF) or one of its variants \cite{Huster:2002:OC, You:2001:VR, Kottas:2013:RSS, Kim:2007:RSS, Mourikis:2007:ICRA}, though methods also exist that employ unscented Kalman filters \cite{Ebcin:2007:ION}, particle filters \cite{Fox:1999:JAIR, Pupilli:2005:BMVC}, or batch optimization methods\cite{Triggs:2000va,Strelow:2004:IJR}.
+For both tightly coupled and loosely coupled VINS, a popular estimator is the Extended Kalman filter (EKF) or one of its variants \cite{Huster:2002:OC, You:2001:VR, Kottas:2013:RSS, Kim:2007:RSS, Mourikis:2007:ICRA}, though methods also exist that employ unscented Kalman filters \cite{Ebcin:2007:ION}, particle filters \cite{Fox:1999:JAIR, Pupilli:2005:BMVC}, or batch optimization methods\cite{Triggs:2000va,Strelow:2004:IJR}.
 
@@
 Furthermore, EKF-based VINS are inconsistent, that is, the state uncertainties are underestimated.
-Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, which reduces the reliability of such estimators.
+Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, reducing the reliability of such estimators.
 
@@
 This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
-Except where otherwise noted (i.e., \ref{sec:state_parametrization} and \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
+Except where otherwise noted (i.e., state parametrization in \ref{sec:state_parametrization} and integration method in \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
 

### 2015-03-05 Report/project_report.tex
updated conclusion

@@
 To accommodate our alternative state parametrization, we used the pre-integrated linear velocity measurements provided in the dataset instead of the raw linear accelerations from the IMU. 
-We extracted point features from the left camera in the stereo pair using Speeded Up Robust Features (SURF) \cite{Bay:2008:CVIU} and tracked them temporally using Kanade-Lucas-Tomasi (KLT) tracking \cite{Tomasi:91}.
+We extracted between 50 and 100 salient point features from the left camera in the stereo pair using Speeded Up Robust Features (SURF) \cite{Bay:2008:CVIU} and tracked them temporally using Kanade-Lucas-Tomasi (KLT) tracking \cite{Tomasi:91}.
 We rejected outliers using an M-estimator Sample Consensus (MSAC) \cite{TorrZisserman:1998} procedure with a 2-point similarity transform.
@@
 \section{Conclusions} \label{sec:conclusions}
-Our results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
+% Our results show that both the MSCKF and SWF provide substantial gains in accuracy over pure IMU integration on typical datasets with many visible features.
 On the datasets we tested, the SWF slightly outperformed the MSCKF, but the MSCKF improved in accuracy with additional features while the SWF was less sensitive to feature quantity.
-It is unclear from these results which algorithm would perform best as the number of features grows very large.
+In relatively featureless environments, neither filter performed substantially better than pure IMU integration.
 
-In relatively featureless environments, neither filter performed substantially better than pure IMU integration.
 We stress that our inertial data was obtained from high-quality IMUs and has been sanitized to account for gravity, biases, and integration error.
@@
 We conclude that the MSCKF may be a better choice of algorithm when computational resources are limited and the operational environment is feature-rich, but that the SWF may be preferable in situations where robustness is paramount.
+
+In future work we would like to investigate the sensitivity of these algorithms to variations in other parameters such as scene geometry, frame rate, field of view, and camera type (e.g., stereo and omnidirectional cameras).
 

### 2015-04-09 Report/project_report.tex
Off to Halifax!

@@
 
-\author{Lee Clement, Valentin Peretroukhin, Jacob Lambert, and Jonathan Kelly% <-this % stops a space
+\author{Lee E Clement, Valentin Peretroukhin, Jacob Lambert, and Jonathan Kelly% <-this % stops a space
 %\thanks{*This work was not supported by any organization}% <-this % stops a space
-\thanks{All authors are at the Institute for Aerospace Studies, University of Toronto, Canada
+\thanks{Lee E Clement and Valentin Peretroukhin jointly assert first authorship. All authors are at the Institute for Aerospace Studies, University of Toronto, Canada
         {\tt \{lee.clement, v.peretroukhin, jacob.lambert\} @mail.utoronto.ca, jkelly@utias.utoronto.ca}
@@
 \begin{abstract}
-Accurate and consistent egomotion estimation is an important part of autonomous navigation.
-For this task, the combination of visual and inertial sensors is an inexpensive, compact, and complementary hardware suite that can be used on many ground and aerial vehicles.
+Accurate and consistent egomotion estimation is a critical component of autonomous navigation.
+For this task, the combination of visual and inertial sensors is an inexpensive, compact, and complementary hardware suite that can be used on many types of vehicles.
 In this work, we compare two modern approaches to egomotion estimation: the Multi-State Constraint Kalman Filter (MSCKF) and the Sliding Window Filter (SWF).
@@
 In general, our results show the SWF to be more accurate and less sensitive to tuning parameters than the MSCKF.
-However, the MSCKF is significantly more computationally efficient, has good consistency properties, and improves in accuracy as more features are tracked. 
+However, the MSCKF is computationally cheaper, has good consistency properties, and improves in accuracy as more features are tracked. 
 \end{abstract}
@@
 Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
-Since an IMU directly measures linear accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
+Since an IMU measures linear accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
 However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
@@
 On the other hand, cameras are not robust to motion blur induced by rapid motions.
-In these cases, IMU data can be relied upon more heavily in estimating egomotion.
+In these cases, IMU data can be relied upon more heavily when estimating egomotion.
 
@@
 This problem is often complicated by the absence of a known map of features from which the camera can generate measurements.
-Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
+Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on mapping may vary from algorithm to algorithm.
 
@@
 In contrast to \cite{Leutenegger:2015:IJRR}, our SWF optimizes over a constant number of timesteps rather than keyframes.
-We also conduct a more extensive characterization of the sensitivity of the MSCKF to certain parameters and compare both algorithms using data from a hand-held sensor rig that mimics the more arbitrary motion of a micro aerial vehicle (MAV).
+We also conduct a more extensive characterization of the sensitivity of the MSCKF to certain tuning parameters and compare both algorithms using data from a hand-held sensor rig that mimics the more arbitrary motion of a micro aerial vehicle (MAV).
 
@@
 The IMU reports translational and rotational velocities, while the stereo camera observes point features.
-We artificially blinded the stereo camera by using measurements from the left camera only.}
+Since we are comparing monocular algorithms, we used measurements from the left camera of the stereo pair only.}
     \label{fig:sensorhead}
+    \vspace{-0.25cm}
 \end{figure}
@@
 
-EKF-SLAM is an efficient recursive algorithm for small-scale, online tasks, but maintaining the entire map as part of the state is a weakness for navigation over long distances.
+EKF-SLAM is an efficient recursive algorithm for small-scale, online tasks, but maintaining the entire map as part of the state is computationally expensive when navigating over long distances.
 Indeed, the computational complexity of EKF-SLAM scales far too poorly with the number of features to be applied naively to this problem.
@@
 Furthermore, EKF-based VINS are inconsistent, that is, the state uncertainties are underestimated.
-Li and Mourikis \cite{Li:2013coa} showed that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system, reducing the reliability of such estimators.
+Prior work has shown that this is due to the fact that the Jacobians in the linearized model of a VINS have different observability properties than the actual nonlinear system \cite{Huang:2010:IJRR,Li:2013coa}.
 
@@
 
-The shortcomings of the EKF motivate the use of filters that operate on a subset of the problem. Algorithms operating in constant time with respect to map size are particularly desirable.
+The shortcomings of the EKF motivate the use of filters that operate on a larger subset of the problem. Algorithms operating in constant time with respect to map size are particularly desirable.
 % The Variable State Dimension Filter (VSDF)\cite{McLauchlan:1993:SPIE} is an early example of such algorithms.
@@
 This update step typically occurs when a tracked feature goes out of view of the camera, but it may also be triggered if the length of a feature track exceeds a preset threshold.
-Except where otherwise noted (i.e., state parametrization in \ref{sec:state_parametrization} and integration method in \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
+Except where otherwise noted (i.e., state parametrization in Section \ref{sec:state_parametrization} and integration method in Section \ref{sec:imu_prop}), we implemented the MSCKF as presented in \cite{Mourikis:2006:TechReport, Mourikis:2007:ICRA}.
 
@@
 \end{align}
-where $\hat{\Vec{P}}_{II,k}$ is the $12\times12$ covariance matrix of the current IMU state, $\hat{\Vec{P}}_{CC,k}$ is the $6N\times6N$ covariance matrix of the camera poses, and $\hat{\Vec{P}}_{IC,k}$ is the $12\times6N$ cross-correlation between the current IMU state and the past camera poses.
+where $\hat{\Vec{P}}_{II,k}$ is the $12\times12$ covariance matrix of the current IMU state, $\hat{\Vec{P}}_{CC,k}$ is the $6N\times6N$ covariance matrix of the camera poses, and $\hat{\Vec{P}}_{IC,k}$ is the $12\times6N$ cross-covariance between the current IMU state and the past camera poses.
 
@@
     \label{fig:nullspace}
+    \vspace{-0.25cm}
 \end{figure}
@@
 
-    \vspace{-0.2cm}
+    \vspace{-0.25cm}
 \end{figure*}
@@
 Another point in favour of the MSCKF is that the computational effort required was an order of magnitude smaller than for the SWF.
-The MSCKF may therefore be better suited to vehicles with limited computational resources, particularly if they are operating in feature-rich environments.
+The MSCKF may therefore be better suited to platforms with limited computational resources, particularly if they are operating in feature-rich environments.
 
@@
     \label{fig:comp_3}
-    \vspace{-0.2cm}
+    \vspace{-0.25cm}
 \end{figure}
@@
     \label{tab:rmse_nees}
-    \vspace{-0.2cm}
+    \vspace{-0.5cm}
 \end{table}
@@
     \label{fig:kitti}
+
 \end{figure*}
@@
     \label{fig:composite-kitti}
-    \vspace{-0.2cm}
+    \vspace{-0.5cm}
 \end{figure*}
@@
     \label{tab:kitti_results}
-    \vspace{-0.2cm}
+    \vspace{-0.25cm}
 \end{table}
